\section{User Guide}
\label{use}

This part of the manual documents the user interface of the \biblatex package. The user guide covers everything you need to know in order to use \biblatex with the default styles that come with this package. You should read the user guide first in any case. If you want to write your own citation and\slash or bibliography styles, continue with the author guide afterwards.

\subsection{Package Options}
\label{use:opt}

All package options are given in \keyval notation. The value \texttt{true} is omissible with all boolean keys. For example, giving \opt{sortcites} without a value is equivalent to \kvopt{sortcites}{true}.

\subsubsection{Load-time Options}
\label{use:opt:ldt}

The following options must be used as \biblatex is loaded, \ie in the optional argument to \cmd{usepackage}.

\begin{optionlist}

\optitem[biber]{backend}{\opt{bibtex}, \opt{bibtex8}, \opt{biber}}

Specifies the database backend. The following backends are supported:

\begin{valuelist}

\item[biber] \biber, the default backend of \biblatex, supports Ascii, 8-bit encodings, \utf, on-the-fly reencoding, locale"=specific sorting, and many other features. Locale"=specific sorting, case"=sensitive sorting, and upper\slash lowercase precedence are controlled by the options \opt{sortlocale}, \opt{sortcase}, and \opt{sortupper}, respectively.

\item[bibtex] Legacy \bibtex. Traditional \bibtex supports Ascii encoding only. Sorting is always case"=insensitive.

\item[bibtex8] \bin{bibtex8}, the 8-bit implementation of \bibtex, supports Ascii and 8-bit encodings such as Latin~1.


\end{valuelist}

See \secref{use:bibtex} for details of using \bibtex as a backend.

\valitem[numeric]{style}{file}

Loads the bibliography style \prm{file}\path{.bbx} and the citation style \prm{file}\path{.cbx}. See \secref{use:xbx} for an overview of the standard styles.

\valitem[numeric]{bibstyle}{file}

Loads the bibliography style \prm{file}\path{.bbx}. See \secref{use:xbx:bbx} for an overview of the standard bibliography styles.

\valitem[numeric]{citestyle}{file}

Loads the citation style \prm{file}\path{.cbx}. See \secref{use:xbx:cbx} for an overview of the standard citation styles.

\boolitem[false]{natbib}

Loads compatibility module which provides aliases for the citation commands of the \sty{natbib} package. See \secref{use:cit:nat} for details.

\boolitem[false]{mcite}

Loads a citation module which provides \sty{mcite}\slash\sty{mciteplus}-like citation commands. See \secref{use:cit:mct} for details.

\optitem[auto]{casechanger}{\opt{auto}, \opt{latex2e}, \opt{expl3}}

This option selects the implementation of \biblatex's case changing functions, most prominently \cmd{MakeSentenceCase*}. \opt{expl3} selects the new implementation based on the \latex3 module \sty{l3text}. Note that the \sty{l3text} module assumes \utf input and that your \sty{expl3} version should be new enough (at least version 2020-04-06). \opt{latex2e} selects the original implementation, which has tricky brace protection behaviour and some shortcomings when dealing with non-ASCII characters. The default \opt{auto} selects the case changing code based on the available \sty{expl3} version and detected document encoding (\opt{expl3} is selected if \sty{expl3} is at least version 2020-04-06 and the document encoding is detected as \utf).

\end{optionlist}

\subsubsection{Preamble Options}
\label{use:opt:pre}

\paragraph{General}
\label{use:opt:pre:gen}

The following options may be used in the optional argument to \cmd{usepackage} as well as in the configuration file and the document preamble. The default value listed to the right is the package default. Note that bibliography and citation styles may modify the default setting at load time, see \secref{use:xbx} for details.

\begin{optionlist}

\optitem[nty]{sorting}{\opt{nty}, \opt{nyt}, \opt{nyvt}, \opt{anyt}, \opt{anyvt}, \opt{ynt}, \opt{ydnt}, \opt{none}, \opt{debug}, \prm{name}}

The sorting order of the bibliography. Unless stated otherwise, the entries are sorted in ascending order. The following choices are available by default:

\begin{valuelist}
\item[nty] Sort by name, title, year.
\item[nyt] Sort by name, year, title.
\item[nyvt] Sort by name, year, volume, title.
\item[anyt] Sort by alphabetic label, name, year, title.
\item[anyvt] Sort by alphabetic label, name, year, volume, title.
\item[ynt] Sort by year, name, title.
\item[ydnt] Sort by year (descending), name, title.
\item[none] Do not sort at all. All entries are processed in citation order.
\item[debug] Sort by entry key. This is intended for debugging only.
\item[\prm{name}] Use \prm{name}, as defined with \cmd{DeclareSortingTemplate} (\secref{aut:ctm:srt})
\end{valuelist}

Using any of the <alphabetic> sorting templates only makes sense in conjunction with a bibliography style which prints the corresponding labels. Note that some bibliography styles initialize this package option to a value different from the package default (\opt{nty}). See \secref{use:xbx:bbx} for details. Please refer to \secref{use:srt} for an in"=depth explanation of the above sorting options as well as the fields considered in the sorting process. See also \secref{aut:ctm:srt} on how to adapt the predefined templates or define new ones.

\boolitem[true]{sortcase}

Whether or not to sort the bibliography and the list of shorthands case"=sensitively.

\boolitem[true]{sortupper}

This option corresponds to \biber's \opt{--sortupper} command-line option. If enabled, the bibliography is sorted in <uppercase before lowercase> order. Disabling this option means <lowercase before uppercase> order.

\optitem[auto]{sortlocale}{\opt{auto}, \prm{locale}}

This option sets the global sorting locale. Every sorting template inherits this locale if none is specified using the \prm{locale} option to \cmd{printbibliography}. Setting this to \opt{auto} requests that it be set to the \sty{babel}/\sty{polyglossia} main document language identifier, if these packages are used and \texttt{en\_US} otherwise. \biber will map \sty{babel}/\sty{polyglossia} language identifiers into sensible locale identifiers (see the \biber documentation). You can therefore specify either a normal locale identifier like \texttt{de\_DE\_phonebook}, \texttt{es\_ES} or one of the supported \sty{babel}/\sty{polyglossia} language identifiers if the mapping \biber makes of this is fine for you.

\boolitem[false]{sortcites}

Whether or not to sort citations if multiple entry keys are passed to a citation command. If this option is enabled, citations are sorted according to the current bibliography context sorting template (see \secref{use:bib:context}). This feature works with all citation styles.

\boolitem[false]{sortsets}

Whether or not to sort set members according to the active reference context sorting scheme. By default this is false and set members appear in the order given in the data source.

\intitem[3]{maxnames}

A threshold affecting all lists of names (\bibfield{author}, \bibfield{editor}, etc.). If a list exceeds this threshold, \ie if it holds more than \prm{integer} names, it is automatically truncated according to the setting of the \opt{minnames} option. \opt{maxnames} is the master option which sets \opt{maxbibnames}, \opt{maxcitenames} and \opt{maxsortnames}. Note that the \opt{uniquelist} feature can locally override \opt{maxnames}, see the documentation of the \opt{uniquelist} option in \secref{use:opt:pre:int} and \secref{aut:cav:amb}.

\intitem[1]{minnames}

A limit affecting all lists of names (\bibfield{author}, \bibfield{editor}, etc.). If a list holds more than \prm{maxnames} names, it is automatically truncated to \prm{minnames} names. The \prm{minnames} value must be smaller than or equal to \prm{maxnames}. \opt{minnames} is the master option which sets both \opt{minbibnames} and \opt{mincitenames}. Like \opt{maxnames} the value of \opt{minnames} can be overridden by \opt{uniquelist}.

\intitem[\prm{maxnames}]{maxbibnames}

Similar to \opt{maxnames} but affects only the bibliography.

\intitem[\prm{minnames}]{minbibnames}

Similar to \opt{minnames} but affects only the bibliography.

\intitem[\prm{maxnames}]{maxcitenames}

Similar to \opt{maxnames} but affects only the citations in the document body.


\intitem[\prm{minnames}]{mincitenames}

Similar to \opt{minnames} but affects only the citations in the document body.

\intitem[\prm{maxbibnames}]{maxsortnames}

Similar to \opt{maxnames} but affects only the names visible to sorting. Since this defaults to \prm{maxbibnames}, you should set this after \opt{maxbibnames} if \opt{maxbibnames} is explicitly set.

\intitem[\prm{minbibnames}]{minsortnames}

Similar to \opt{minnames} but affects only the names visible to sorting. Since this defaults to \prm{minbibnames}, you should set this after \opt{minbibnames} if \opt{minbibnames} is explicitly set.

\intitem[3]{maxitems}

Similar to \opt{maxnames}, but affecting all literal lists (\bibfield{publisher}, \bibfield{location}, etc.).

\intitem[1]{minitems}

Similar to \opt{minnames}, but affecting all literal lists (\bibfield{publisher}, \bibfield{location}, etc.).

\optitem{autocite}{\opt{plain}, \opt{inline}, \opt{footnote}, \opt{superscript}, \opt{...}}

This option controls the behavior of the \cmd{autocite} command discussed in \secref{use:cit:aut}. The \opt{plain} option makes \cmd{autocite} behave like \cmd{cite}, \opt{inline} makes it behave like \cmd{parencite}, \opt{footnote} makes it behave like \cmd{footcite}, and \opt{superscript} makes it behave like \cmd{supercite}. The options \opt{plain}, \opt{inline}, and \opt{footnote} are always available, the \opt{superscript} option is only provided by the numeric citation styles which come with this package. The citation style may also define additional options. The default setting of this option depends on the selected citation style, see \secref{use:xbx:cbx}.

\boolitem[true]{autopunct}

This option controls whether the citation commands scan ahead for punctuation marks. See \secref{use:cit} and \cmd{DeclareAutoPunctuation} in \secref{aut:pct:cfg} for details.

\optitem[autobib]{language}{\opt{autobib}, \opt{autocite}, \opt{auto}, \prm{language}}

This option controls multilingual support. By default \biblatex automatically picks up the active surrounding language from the \sty{babel}/\sty{polyglossia} package\footnote{Note that \biblatex has only limited support for \sty{polyglossia} versions prior to v1.45. If \sty{polyglossia} is used, it should be updated to version~1.45 (2019/10/27) or above.} (and fall back to English if \sty{babel}/\sty{polyglossia} is not available). \opt{autobib} switches the language for each entry in the bibliography using the \bibfield{langid} field and the language environment specified by the \opt{autolang} option. \opt{autocite} switches the language for each citation using the \bibfield{langid} field and the language environment specified by the \opt{autolang} option. \opt{auto} is a shorthand to set both \opt{autobib} and \opt{autocite}. It is also possible to select the package language manually. In this case, the language chosen will override the \bibfield{langid} of entries and you should still choose a language switching environment with the \opt{autolang} option to select how the switch to the manually chosen language is handled. Please refer to \tabref{bib:fld:tab1} for a list of supported languages and the corresponding identifiers.

\boolitem[true]{clearlang}

If this option is enabled, \biblatex will automatically clear the \bibfield{language} field of all entries whose language matches the \sty{babel}/\sty{polyglossia} language of the document (or the language specified explicitly with the \opt{language} option) in order to omit redundant language specifications. The language mappings required by this feature are provided by the \cmd{DeclareRedundantLanguages} command from \secref{aut:lng:cmd}.
This option is also settable on a per-type and per-entry basis.

\optitem[none]{autolang}{\opt{none}, \opt{hyphen}, \opt{other}, \opt{other*}, \opt{langname}}

This option controls which \sty{babel} language environment\footnote{\sty{polyglossia} understands the \sty{babel} language environments too and so this option controls both the \sty{babel} and \sty{polyglossia} language environments.} is used if the \sty{babel}/\sty{polyglossia} package is loaded and a bibliography entry includes a \bibfield{langid} field (see \secref{bib:fld:spc}). Note that regardless of the selected value \biblatex automatically adjusts to the main document language if \sty{babel}/\sty{polyglossia} is loaded. In multilingual documents, it will also continually adjust to the current language as far as citations and the default language of the bibliography is concerned. The effect of additional language adjustment, which can negate the effect of picking up the surrounding language, depends on the language environment selected by this option. The possible choices are:

\begin{valuelist}

\item[none]
Do not use any additional enclosing language environment at all. This means that citations and the bibliography are set in the currently active language---this need not be the main language.

\item[hyphen]
Enclose the entry in a \env{hyphenrules} environment. This will load hyphenation patterns for the language specified in the \bibfield{langid} field of the entry, if available. Localisation strings and extra language definitions are not changed and taken from the surrounding language environment.

\item[other]
Enclose the entry in an \env{otherlanguage} environment. This will load hyphenation patterns for the specified language, enable all extra definitions which \sty{babel}/\sty{polyglossia} and \biblatex provide for the respective language, and translate key terms such as <editor> and <volume>. The extra definitions include localisations of the date format, of ordinals, and similar things.

\item[other*]
Enclose the entry in an \env{otherlanguage*} environment. Please note that \biblatex treats \env{otherlanguage*} like \env{otherlanguage} if \opt{langhook} is set to \opt{extras}.

\item[langname]
\sty{polyglossia} only. Enclose the entry in a \env{$<$languagename$>$} environment. The benefit of this option value for \sty{polyglossia} users is that it takes note of the \bibfield{langidopts} field so that you can add per-language options to an entry (like selecting a language variant). When using \sty{babel}, this option does the same as the \opt{other} option value.

\end{valuelist}

\optitem[captions]{langhook}{\opt{captions}, \opt{extras}}

This option controls whether bibliography strings and extras are written to \cmd{captions$<$language$>$} or \cmd{extras$<$language$>$}. The exact effect of this option depend on the language package (\sty{babel}/\sty{polyglossia}). Broadly speaking, the language switching environments provided by those packages (except \env{hyphenrules}) either switch language captions and extras or only language extras. Hence, if this option is set to \opt{extras}, all language switches will affect \biblatex, whereas with \opt{captions} only language switches that also switch other parts of the document language affect \biblatex.

\optitem[none]{block}{\opt{none}, \opt{space}, \opt{par}, \opt{nbpar}, \opt{ragged}}

This option controls the extra spacing between blocks, \ie larger segments of a bibliography entry. The possible choices are:

\begin{valuelist}

\item[none] Do not add anything at all.

\item[space] Insert additional horizontal space between blocks. This is similar to the default behavior of the standard \latex document classes.

\item[par] Start a new paragraph for every block. This is similar to the \opt{openbib} option of the standard \latex document classes.

\item[nbpar] Similar to the \opt{par} option, but disallows page breaks at block boundaries and within an entry.

\item[ragged] Inserts a small negative penalty to encourage line breaks at block boundaries and sets the bibliography ragged right.

\end{valuelist}

The \cmd{newblockpunct} command may also be redefined directly to achieve different results, see \secref{use:fmt:fmt}. Also see \secref{aut:pct:new} for additional information.

\boolitem[false]{locallabelwidth}

This option controls whether \cmd{printbibliography} uses a locally calculated value for \cmd{labelnumberwidth} and \cmd{labelalphawidth} or the global value calculated from all entries. The local value is calculated separately for each bibliography and takes into account only the entries displayed in that bibliography. This option is useful if there are several bibliographies with wildly varying label lengths in the same document.

\optitem[foot+end]{notetype}{\opt{foot+end}, \opt{footonly}, \opt{endonly}}

This option controls the behavior of \cmd{mkbibfootnote}, \cmd{mkbibendnote}, and similar wrappers from \secref{aut:fmt:ich}. The possible choices are:

\begin{valuelist}
\item[foot+end] Support both footnotes and endnotes, \ie \cmd{mkbibfootnote} will generate footnotes and \cmd{mkbibendnote} will generate endnotes.
\item[footonly] Force footnotes, \ie make \cmd{mkbibendnote} generate footnotes.
\item[endonly] Force endnotes, \ie make \cmd{mkbibfootnote} generate endnotes.
\end{valuelist}

\optitem[auto]{hyperref}{\opt{true}, \opt{false}, \opt{auto}, \opt{manual}}

Whether or not to transform citations and back references into clickable hyperlinks. This feature requires the \sty{hyperref} package. It also requires support by the selected citation style. All standard styles which ship with this package support hyperlinks. \kvopt{hyperref}{auto} automatically detects if the \sty{hyperref} package has been loaded. This is the default setting. \kvopt{hyperref}{false} explicitly disables links even if \sty{hyperref} is loaded. \kvopt{hyperref}{true} enables links when \sty{hyperref} is loaded, it cannot explicitly enable links if \sty{hyperref} is not loaded, as such it works exactly like \kvopt{hyperref}{auto} except that it will issue a warning if \sty{hyperref} is not loaded. \kvopt{hyperref}{manual} gives full manual control over \sty{hyperref} interaction, it should only be needed by package authors in very special circumstances. With the \kvopt{hyperref}{manual} setting you are responsible to enable or disable \sty{hyperref} support manually with \cmd{BiblatexManualHyperrefOn} or \cmd{BiblatexManualHyperrefOff} yourself. One of the two commands must be called exactly once; \cmd{BiblatexManualHyperrefOn} can only be called after \sty{hyperref} is loaded.

\boolitem[false]{backref}

Whether or not to print back references in the bibliography. The back references are a list of page numbers indicating the pages on which the respective bibliography entry is cited. If there are \env{refsection} environments in the document, the back references are local to the reference sections. Strictly speaking, this option only controls whether the \biblatex package collects the data required to print such references. This feature still has to be supported by the selected bibliography style. All standard styles which come with this package do so.

\optitem[three]{backrefstyle}{\opt{none}, \opt{three}, \opt{two}, \opt{two+}, \opt{three+}, \opt{all+}}

This option controls how sequences of consecutive pages in the list of back references are formatted. The following styles are available:

\begin{valuelist}

\item[none] Disable this feature, \ie do not compress the page list.

\item[three] Compress any sequence of three or more consecutive pages to a range, \eg the list <1, 2, 11, 12, 13, 21, 22, 23, 24> is compressed to <1, 2, 11--13, 21--24>.

\item[two] Compress any sequence of two or more consecutive pages to a range, \eg the above list is compressed to <1--2, 11--13, 21--24>.

\item[two+] Similar in concept to \opt{two} but a sequence of exactly two consecutive pages is printed using the starting page and the localisation string \texttt{sequens}, \eg the above list is compressed to <1\,sq., 11--13, 21--24>.

\item[three+] Similar in concept to \opt{two+} but a sequence of exactly three consecutive pages is printed using the starting page and the localisation string \texttt{sequentes}, \eg the above list is compressed to <1\,sq., 11\,sqq., 21--24>.

\item[all+] Similar in concept to \opt{three+} but any sequence of consecutive pages is printed as an open-ended range, \eg the above list is compressed to <1\,sq., 11\,sqq., 21\,sqq.>.

\end{valuelist}

All styles support both Arabic and Roman numerals. In order to avoid potentially ambiguous lists, different sets of numerals will not be mixed when generating ranges, \eg the list <iii, iv, v, 6, 7, 8> is compressed to <iii--v, 6--8>.

\optitem[setonly]{backrefsetstyle}{\opt{setonly}, \opt{memonly}, \opt{setormem}, \opt{setandmem}, \opt{memandset}, \opt{setplusmem}}

This option controls how back references to \bibtype{set} entries and their members are handled. The following options are available:

\begin{valuelist}

\item[setonly] All back references are added to the \bibtype{set} entry. The \bibfield{pageref} lists of set members remain blank.

\item[memonly] References to set members are added to the respective member. References to the \bibtype{set} entry are added to all members. The \bibfield{pageref} list of the \bibtype{set} entry remains blank.

\item[setormem] References to the \bibtype{set} entry are added to the \bibtype{set} entry. References to set members are added to the respective member.

\item[setandmem] References to the \bibtype{set} entry are added to the \bibtype{set} entry. References to set members are added to the respective member and to the \bibtype{set} entry.

\item[memandset] References to the \bibtype{set} entry are added to the \bibtype{set} entry and to all members. References to set members are added to the respective member.

\item[setplusmem] References to the \bibtype{set} entry are added to the \bibtype{set} entry and to all members. References to set members are added to the respective member and to the \bibtype{set} entry.

\end{valuelist}

\boolitem[true]{backreffloats}

Whether to enable back references to citations in floats.


\optitem[false]{indexing}{\opt{true}, \opt{false}, \opt{cite}, \opt{bib}}

This option controls indexing in citations and in the bibliography. More precisely, it affects the \cmd{ifciteindex} and \cmd{ifbibindex} commands from \secref{aut:aux:tst}. The option is settable on a global, a per-type, or on a per-entry basis. The possible choices are:

\begin{valuelist}
\item[true] Enable indexing globally.
\item[false] Disable indexing globally.
\item[cite] Enable indexing in citations only.
\item[bib] Enable indexing in the bibliography only.
\end{valuelist}

This feature requires support by the selected citation style. All standard styles which come with this package support indexing of both citations and entries in the bibliography. Note that you still need to enable indexing globally with \cmd{makeindex} to get an index.

\boolitem[false]{loadfiles}

This option controls whether external files requested by way of the \cmd{printfile} command are loaded. See also \secref{use:use:prf} and \cmd{printfile} in \secref{aut:bib:dat}. Note that this feature is disabled by default for performance reasons.

\optitem[none]{refsection}{\opt{none}, \opt{part}, \opt{chapter}, \opt{chapter+}, \opt{section}, \opt{section+}, \opt{subsection}, \opt{subsection+}}

This option automatically starts a new reference section at a document division such as a chapter or a section. This is equivalent to the \cmd{newrefsection} command, see \secref{use:bib:sec} for details. The following choice of document divisions is available:

\begin{valuelist}
\item[none] Disable this feature.
\item[part] Start a reference section at every \cmd{part} command.
\item[chapter] Start a reference section at every \cmd{chapter} command.
\item[chapter+] Start a reference section at every \cmd{chapter} and every higher level of sectioning, i.e. \cmd{part}.
\item[section] Start a reference section at every \cmd{section} command.
\item[section+] Start a reference section at every \cmd{section} and every higher level of sectioning, i.e. \cmd{part} and \cmd{chapter} (if available).
\item[subsection] Start a reference section at every \cmd{subsection} command.
\item[subsection+] Start a reference section at every \cmd{subsection} and every higher level of sectioning, i.e. \cmd{part}, \cmd{chapter} (if available) and \cmd{section}.
\end{valuelist}
%
The starred versions of these commands will not start a new reference section.

\optitem[none]{refsegment}{\opt{none}, \opt{part}, \opt{chapter}, \opt{chapter+}, \opt{section}, \opt{section+}, \opt{subsection}, \opt{subsection+}}

Similar to the \opt{refsection} option but starts a new reference segment. This is equivalent to the \cmd{newrefsegment} command, see \secref{use:bib:seg} for details. When using both options, note that you can only apply this option to a lower"=level document division than the one \opt{refsection} is applied to and that nested reference segments will be local to the enclosing reference section.

\optitem[none]{citereset}{\opt{none}, \opt{part}, \opt{chapter}, \opt{chapter+}, \opt{section}, \opt{section+}, \opt{subsection}, \opt{subsection+}}

This option automatically executes the \cmd{citereset} command from \secref{use:cit:msc} at a document division such as a chapter or a section. The following choice of document divisions is available:

\begin{valuelist}
\item[none] Disable this feature.
\item[part] Perform a reset at every \cmd{part} command.
\item[chapter] Perform a reset at every \cmd{chapter} command.
\item[chapter+] Perform a reset at every \cmd{chapter} and \cmd{part} command.
\item[section] Perform a reset at every \cmd{section} command.
\item[section+] Perform a reset at every \cmd{section}, \prm{chapter} (if supported by the class) and \cmd{part} command.
\item[subsection] Perform a reset at every \cmd{subsection} command.
\item[subsection+] Perform a reset at every \cmd{subsection}, \cmd{section}, \prm{chapter} (if supported by the class) and \cmd{part} command.
\end{valuelist}

\boolitem[true]{abbreviate}

Whether or not to use long or abbreviated strings in citations and in the bibliography. This option affects the localisation modules. If this option is enabled, key terms such as <editor> are abbreviated. If not, they are written out.
This option is also settable on a per-type or per-entry basis.

\optitem[comp]{date}{\opt{year}, \opt{short}, \opt{long}, \opt{terse}, \opt{comp}, \opt{ymd}, \opt{iso}}

This option controls the basic format of printed date specifications. The following choices are available:

\begin{valuelist}
\item[year] Use only years, for example:\par
2010\par
2010--2012\par
\item[short] Use the short format with verbose ranges, for example:\par
01/01/2010\par
21/01/2010--30/01/2010\par
01/21/2010--01/30/2010
\item[long] Use the long format with verbose ranges, for example:\par
1st January 2010\par
21st January 2010--30th January 2010\par
January 21, 2010--January 30, 2010\par
\item[terse] Use the short format with compact ranges, for example:\par
21--30/01/2010\par
01/21--01/30/2010
\item[comp] Use the long format with compact ranges, for example:\par
21st--30th January 2010\par
January 21--30, 2010\par
\item[iso] Use ISO8601 Extended Format (\texttt{yyyy-mm-dd}), for example:\par
2010-01-01\par
2010-01-21/2010-01-30
\item[ymd] A year-month-day format which can be modified by other options unlike strict \acr{ISO8601-2}, for example:\par
2010-1-1\par
2010-1-21/2010-1-30
\end{valuelist}
%
Note that \opt{iso} format will enforce \kvopt{dateera}{astronomical}, \kvopt{datezeros}{true}, \kvopt{timezeros}{true}, \kvopt{seconds}{true}, \kvopt{$<$datetype$>$time}{24h} and \kvopt{julian}{false}. \opt{ymd} is an EDTF-like format but which can change the various options which the strict \opt{iso} option does not allow for.

As seen in the above examples, the actual date format is language specific. Note that the month name in all long formats is responsive to the \opt{abbreviate} package option. The leading zeros for months and days in all short formats may be controlled separately with the \opt{datezeros} package option. The leading zeros for hours, minutes and seconds in all short formats may be controlled separately with the \opt{timezeros} package option. If outputting times, the printing of seconds and timezones is controlled by the \opt{seconds} and \opt{timezones} options respectively.

The options \opt{julian} and \opt{gregorianstart}  may be used to control when to output Julian Calendar dates.

\optitem[year]{labeldate}{\opt{year}, \opt{short}, \opt{long}, \opt{terse}, \opt{comp}, \opt{ymd}, \opt{iso}}

Similar to the \opt{date} option but controls the format of the date field selected with \cmd{DeclareLabeldate}.

\optitem[comp]{$<$datetype$>$date}{\opt{year}, \opt{short}, \opt{long}, \opt{terse}, \opt{comp}, \opt{ymd}, \opt{iso}}

Similar to the \opt{date} option but controls the format of the \bibfield{$<$datetype$>$date} field in the datamodel.

\optitem{alldates}{\opt{year}, \opt{short}, \opt{long}, \opt{terse}, \opt{comp}, \opt{iso}}

Sets the option for all dates in the datamodel to the same value. The date fields in the default data model are \bibfield{date}, \bibfield{origdate}, \bibfield{eventdate} and \bibfield{urldate}.

\boolitem[false]{julian}

This option controls whether dates before the date specified in the \opt{gregorianstart} option will be converted automatically to the Julian Calendar. Dates so changed will return <true> for the \cmd{ifdatejulian} and \cmd{if$<$datetype$>$datejulian} tests (see \secref{aut:aux:tst}). Please bear in mind that dates consisting of just a year like <1565> will never be converted to a Julian Calendar date because a date without a month and day has an ambiguous Julian Calendar representation\footnote{This is potentially true for dates missing times too but this is not relevant for bibliographic work.}. For example, in the case of <1565>, this is Julian year <1564> until after the Gregorian date <10th January 1565> when the Julian year becomes <1565>.

\valitem{gregorianstart}{YYYY-MM-DD}

This option controls the date before which dates are converted to the Julian Calendar. It is a strict format string, 4-digit year, 2-digit month and day, separated by a single dash character (any valid Unicode character with the <Dash> property). The default is '1582-10-15', the date of the instigation of the standard Gregorian Calendar. This option does not nothing unless \opt{julian} is set to <true>.

\boolitem[true]{datezeros}

This option controls whether \texttt{short} and \texttt{terse} date components are printed with leading zeros unless overridden by specific formatting.

\boolitem[true]{timezeros}

This option controls whether time components are printed with leading zeros unless overridden by specific formatting.

\boolitem[false]{timezones}

This option controls whether timezones are printed when printing times.

\boolitem[false]{seconds}

This option controls whether seconds are printed when printing times.

\boolitem[true]{dateabbrev}

This option controls whether \texttt{long} and \texttt{comp} dates are printed with long or abbreviated month/season names. The option is similar to the generic \opt{abbreviate} option but specific to the date formatting.
This option is also settable on a per-type and per-entry basis.

\boolitem[false]{datecirca}

This option controls whether to output <circa> information about dates. If set to \opt{true}, dates will be preceded by the expansion of the \cmd{datecircaprint} macro (\secref{use:fmt:fmt}).

\boolitem[false]{dateuncertain}

This option controls whether to output uncertainty information about dates. If set to \opt{true}, dates will be followed by the expansion of the \cmd{dateuncertainprint} macro and end dates will be followed by the \cmd{enddateuncertainprint} macro (\secref{use:fmt:fmt}).

\optitem[astronomical]{dateera}{\opt{astronomical}, \opt{secular}, \opt{christian}}

This option controls how date era information is printed. <astronomical> uses \cmd{dateeraprintpre} to print era information \emph{before} start/end dates. <secular> and <christian> uses \cmd{dateeraprint} to print era information \emph{after} the start/end/dates. By default <astronomical> results in a minus sign before BCE/BC dates and <secular>/<christian> results in the relevant localisation strings like <BCE> or <BC> after BCE/BC dates. See the relevant comments in \secref{use:fmt:fmt} and the localisation strings in \secref{aut:lng:key:dt}.

\intitem[0]{dateeraauto}

This option sets the astronomical year, below which era localisation strings are automatically added. This option does nothing without \opt{dateera} being set to <secular> or <christian>.

\optitem[24h]{time}{\opt{12h}, \opt{24h}, \opt{24hcomp}}

This option controls the basic format of printed time specifications. The following choices are available:

\begin{valuelist}
\item[24h] 24-hour format, for example:\par
14:03:23\par
14:3:23\par
14:03:23+05:00\par
14:03:23Z\par
14:21:23--14:23:45\par
14:23:23--14:23:45\par
\item[24hcomp] 24-hour format with compressed ranges, for example:\par
14:21--23 (hours are the same)\par
14:23:23--45 (hour and minute are the same)\par
\item[12h] 12-hour format with (localised) AM/PM markers, for example:\par
2:34 PM\par
2:34 PM--3:50 PM\par
\end{valuelist}
%
As seen in the above examples, the actual time format is language specific. Note that the AM/PM string is responsive to the \opt{abbreviate} package option, if this makes a difference in the specific locale. The leading zeros in the 24-hour formats may be controlled separately with the \opt{timezeros} package option. The separator between time components (\cmd{bibtimesep} and \cmd{bibtzminsep}) and between the time and any timezone (\cmd{bibtimezonesep}) are also language specific and customisable, see \secref{use:fmt:lng}. There are global package options which determine whether seconds and timezones are printed (\opt{seconds} and \opt{timezones}, respectively, see \secref{use:opt:pre:gen}). Timezones, if present, are either <Z> or a numeric positive or negative offset. No default styles print time information. Custom styles may print times by using the \cmd{print$<$datetype$>$time} commands, see \secref{aut:bib:dat}.

\optitem[24h]{labeltime}{\opt{12h}, \opt{24h}, \opt{24hcomp}}

Similar to the \opt{time} option but controls the format of the time part fields obtained from the field selected with \cmd{DeclareLabeldate}.

\optitem[24h]{$<$datetype$>$time}{\opt{12h}, \opt{24h}, \opt{24hcomp}}

Similar to the \opt{time} option but controls the format of the time part fields obtained from the \bibfield{$<$datetype$>$date} field in the datamodel.

\optitem{alltimes}{\opt{12h}, \opt{24h}, \opt{24hcomp}}

Sets \opt{labeltime} and the \opt{$<$datetype$>$time} option for all times in the datamodel to the same value. The date fields supporting time parts in the default data model are \bibfield{date}, \bibfield{origdate}, \bibfield{eventdate} and \bibfield{urldate}.

\boolitem[false]{dateusetime}

Specifies whether to print any time component of a date field after the date component. The separator between the date and time components is \cmd{bibdatetimesep} from \secref{use:fmt:lng}.

\boolitem[false]{labeldateusetime}

Similar to the \opt{dateusetime} option but controls the whether to print time components for the field selected with \cmd{DeclareLabeldate}.

\boolitem[false]{$<$datetype$>$dateusetime}

Similar to the \opt{dateusetime} option but controls the whether to print time components for the \bibfield{$<$datetype$>$date} field in the datamodel.

\boolitem[false]{alldatesusetime}

Sets \opt{labeldateusetime} and the \opt{$<$datetype$>$dateusetime} option for all \bibfield{$<$datetype$>$date} fields in the datamoel.

\boolitem[false]{defernumbers}

In contrast to standard \latex, the numeric labels generated by this package are normally assigned to the full list of references at the beginning of the document body. If this option is enabled, numeric labels (\ie the \bibfield{labelnumber} field discussed in \secref{aut:bbx:fld}) are assigned the first time an entry is printed in any bibliography. See \secref{use:cav:lab} for further explanation.  This option requires two \latex runs after the data has been exported to the \file{bbl} file by the backend (in addition to any other runs required by page breaks changing etc.). An important thing to note is that if you are using this option, then changes to options, the \file{bib} file or certain commands like \cmd{printbibliography} will usually require that you delete your current \file{aux} file and re-run \latex to obtain the correct numbering. See \secref{aut:int}.

\boolitem[false]{punctfont}

This option enables an alternative mechanism for dealing with unit punctuation after a field printed in a different font (for example, a title printed in italics). See \cmd{setpunctfont} in \secref{aut:pct:new} for details.

\optitem[abs]{arxiv}{\opt{abs}, \opt{ps}, \opt{pdf}, \opt{format}}

Path selector for arXiv links. If hyperlink support is enabled, this option controls which version of the document the arXiv \bibfield{eprint} links will point to. The following choices are available:

\begin{valuelist}
\item[abs] Link to the abstract page.
\item[ps] Link to the PostScript version.
\item[pdf] Link to the \pdf version.
\item[format] Link to the format selector page.
\end{valuelist}

See \secref{use:use:epr} for details on support for arXiv and electronic publishing information.

\optitem[auto]{texencoding}{\opt{auto}, \prm{encoding}}

Specifies the encoding of the \file{tex} file. This option affects the data transferred from the backend to \biblatex. This corresponds to \biber's \opt{--output-encoding} option. The following choices are available:

\begin{valuelist}

\item[auto] Try to auto-detect the input encoding. If the \sty{inputenc}\slash \sty{inputenx}\slash \sty{luainputenc} package is available, \biblatex will get the main encoding from that package. If not, it assumes \utf encoding if a \latex format using at least the April 2018 version of the kernel, \xetex or \luatex has been detected, and Ascii otherwise.

\item[\prm{encoding}] Specifies the \prm{encoding} explicitly. This is for odd cases in which auto-detection fails or you want to force a certain encoding for some reason.

\end{valuelist}
%
Note that setting \kvopt{texencoding}{\prm{encoding}} will also affect the \opt{bibencoding} option if \kvopt{bibencoding}{auto}.

\optitem[auto]{bibencoding}{\opt{auto}, \prm{encoding}}

Specifies the default encoding of the \file{bib} files. This can be overridden on a per-datasource basis using the \opt{bibencoding} option to \cmd{addbibresource}, see \secref{use:bib:res}. This option corresponds to \biber's \opt{--input-encoding} option. The following choices are available:

\begin{valuelist}

\item[auto] Use this option if the workflow is transparent, \ie if the encoding of the \file{bib} file is identical to the encoding of the \file{tex} file.

\item[\prm{encoding}] If the encoding of the \file{bib} file is different from the one of the \file{tex} file, you need to specify it explicitly.

\end{valuelist}

By default, \biblatex assumes that the \file{tex} file and the \file{bib} file use the same encoding (\kvopt{bibencoding}{auto}).

\boolitem[false]{safeinputenc}

If this option is enabled, \biblatex will automatically force \kvopt{texencoding}{ascii} if the \sty{inputenc}\slash \sty{inputenx} package has been loaded and the input encoding is \utf, \ie it will ignore any macro-based \utf support and use Ascii only. \biber will then try to convert any non-Ascii data in the \file{bib} file to Ascii. For example, it will convert \texttt{\d{S}} to |\d{S}|. See \secref{bib:cav:enc:enc} for an explanation of why you may want to enable this option.

\boolitem[true]{bibwarn}

By default, \biblatex will report warnings issued by the backend concerning the data in the \file{bib} file as \latex warnings. Use this option to suppress such warnings.

\intitem[2]{mincrossrefs}

Sets the minimum number of cross references to \prm{integer} when requesting a backend run.\footnote{If an entry which is cross-referenced by other entries in the \file{bib} file hits this threshold, it is included in the bibliography even if it has not been cited explicitly. This is a standard feature of the \bibtex format and not specific to \biblatex. See the description of the \bibfield{crossref} field in \secref{bib:fld:spc} for further information.} This option also affects the handling of the \bibfield{xref} field. See the field description in \secref{bib:fld:spc} as well as \secref{bib:cav:ref} for details.

\intitem[2]{minxrefs}

As \opt{mincrossrefs} but for \bibfield{xref} fields.

\boolitem[true]{bibtexcaseprotection}

This option only has an effect when the \sty{expl3} implementation of the case changing functions is selected. If the option is set to \opt{true}, \cmd{MakeSentenceCase*} supports brace protection of words from case change as in classical \bibtex. If the option is set to \opt{false}, pairs of braces no longer imply case protection, which can now be enforced by wrapping the relevant word in \cmd{NoCaseChange}---this makes for a less confusing, if more verbose, markup of case protection.

\end{optionlist}

\paragraph{Style-specific}
\label{use:opt:pre:bbx}

The following options are provided by all standard bibliography styles (as opposed to the core package). The options are available as preamble options like those in \secref{use:opt:pre:gen} and at a per-type and per-entry scope.

\begin{optionlist}

\boolitem[true]{isbn}

This option controls whether the fields \bibfield{isbn}\slash \bibfield{issn}\slash \bibfield{isrn} are printed.

\boolitem[true]{url}

This option controls whether the \bibfield{url} field and the access date is printed. The option only affects entry types whose \bibfield{url} information is optional. The \bibfield{url} field of \bibtype{online} entries is always printed.

\boolitem[true]{doi}

This option controls whether the field \bibfield{doi} is printed.

\boolitem[true]{eprint}

This option controls whether \bibfield{eprint} information is printed.

\boolitem[true]{related}

Whether to use information from related entries or not. See \secref{use:rel}.

\end{optionlist}

\subparagraph{\texttt{alphabetic}/\texttt{numeric}} Additionally, styles of the \texttt{alphabetic} and \texttt{numeric} family support the \opt{subentry} option in global, per-type and per-entry scope.

\begin{optionlist}

\boolitem[false]{subentry}

This option affects the handling of citations to set members and the display of sets in the bibliography. If the option is enabled, citations to individual set members feature an additional letter that identifies the member, that letter is also printed in the bibliography. If the option is disabled, a citation to the member of a set will display just as a citation to the entire set and there will be no additional letters in the bibliography entries enumerating the members.

Suppose \texttt{key1} and \texttt{key2} are members of the set \texttt{set1}. With \opt{subentry} set to \texttt{true} in a numeric style a citation to \texttt{key1} will show as <[1a]> and a citation to \texttt{key2} as <[1b]>, while the entire set \texttt{set1} will be cited as <[1]>. Furthermore <(a)> and <(b)> will be added in front of the entry data for the set members in the bibliography entry for the set. With \opt{subentry} set to \texttt{false} citations to all three keys will show as <[1]>, no additional letter will be printed in the bibliography.

\end{optionlist}

\subparagraph{\texttt{numeric-comp}} The citation style \texttt{numeric-comp}  supports the \opt{subentrycomp} option in global, per-type and per-entry scope.

\begin{optionlist}

\boolitem[true]{subentrycomp}

This option determines whether or not citations to set members are compressed similar to non-set citations. The option only has an effect if \opt{subentry} is set to \texttt{true}.

Suppose \texttt{key1}, \texttt{key2} and \texttt{key3} are members of the set \texttt{set1}. With \opt{subentrycomp} set to \texttt{true} the three entries will be compressed to <[1a--c]> in citations. With \opt{subentry} set to \texttt{false} the citation will show in the more verbose form <[1a, 1b, 1c]>.

The option is intended mainly for backwards compatibility, because earlier versions of \biblatex did not compress set member citations.

\end{optionlist}

\subparagraph{\texttt{authortitle}/\texttt{authoryear}} All bibliography styles of the \texttt{authoryear} and \texttt{authortitle} family as well as all bibliography styles of the \texttt{verbose} family---whose bibliography styles are based on \texttt{authortitle}---support the option \opt{dashed} in global scope.

\begin{optionlist}

\boolitem[true]{dashed}

This option controls whether recurrent the same author\slash editor list in the bibliography are replaced by a dash (\cmd{bibnamdeash}, see \secref{use:fmt:fmt}). If the option is enabled, subsequent mentions of the same name list at the beginning of an entry are replaced by a dash provided the entry is not the first on the current page. If the option is disabled, name lists are never replaced by a dash.

\end{optionlist}

\subparagraph{\texttt{authoryear}} Bibliography styles of the \texttt{authoryear} family provide the option \opt{mergedate} in global, per-type and per-entry scope.

\begin{optionlist}

\optitem[true]{mergedate}{\opt{false}, \opt{minimum}, \opt{basic}, \opt{compact}, \opt{maximum}, \opt{true}}

This option controls whether and how the date specification in the entry is merged with the date label shown directly after the author\slash editor list.

\begin{valuelist}
\item[false] Strictly separate the date specification shown in the entry (styled with \opt{date}) from the date label (styled with \opt{labeldate}). The date will always be shown twice.
\item[minimum] Omit the date specification whenever it coincides \emph{exactly}---including \bibfield{extradate} information---with the output of the date label.
\item[basic] Similar to \opt{minimum}, but the date specification will also be omitted if it differs from the date label only by the absence of the \bibfield{extradate} letter.
\item[compact] Merges all date specifications with the date label. The date format of that merged date label is controlled by \opt{date}, not \opt{labeldate}, even if it is printed in the position of the date label. The \bibfield{issue} field is not merged.
\item[maximum] Like \opt{compact}, but if present the \bibfield{issue} field will also be moved into the date label at the beginning of the entry.
\item[true] An alias for \opt{compact}.
\end{valuelist}

More in-depth examples of this option can be found in the style examples.
\end{optionlist}

\subparagraph{<ibid> styles} Citation styles with <ibid.>\ function, namely \texttt{authortitle-ibid}, \texttt{author\allowbreakhere title-icomp}, \texttt{author\allowbreakhere year-ibid}, \texttt{authoryear-icomp}, \texttt{ver\allowbreakhere bose-ibid}, \texttt{verbose-inote}, \texttt{verbose-trad1}, \texttt{verbose-trad2} and \texttt{verbose-trad3} provide the global \opt{ibidpage} option.

\begin{optionlist}

\boolitem[false]{ibidpage}

Whether \emph{ibidem} without page reference means <same work> or <same work + same page>. If set to \texttt{true} a page range postnote will be suppressed in an \emph{ibidem} citation if the last citation was to the same page range. With \texttt{ibidpage=false} the postnote is not omitted. Citations to different page ranges than the previous always produce the page ranges with either setting.

\end{optionlist}

\subparagraph{\texttt{verbose}} All citation styles of the \texttt{verbose} family provide the global option \opt{citepages}.

\begin{optionlist}

\optitem[permit]{citepages}{\opt{permit}, \opt{suppress}, \opt{omit}, \opt{separate}}

This option controls the output of the \bibfield{page}\slash\bibfield{pagetotal} field in the full citation in combination with a postnote containing a page range. The option can be used to suppress references to two page ranges in full citations like the following

\begin{quote}
Author. \enquote{Title.} In: \emph{Book,} pp.\,100--150, p.\,125.
\end{quote}

Here <p.\,125> is the \bibfield{postnote} argument and <pp.\,100--150> is the value of the \bibfield{pages} field.

\begin{valuelist}
\item[permit] Allow duplication of page specifications, i.e.\ print both \bibfield{page}\slash\bibfield{pagetotal} and \bibfield{postnote}.
\item[suppress] Unconditionally suppress the \bibfield{pages}\slash \bibfield{pagetotal} fields in citations, regardless of the \bibfield{postnote}.
\item[omit] Suppress the \bibfield{pages}\slash \bibfield{pagetotal} if the \bibfield{postnote} contains a page range. They are still printed if there is no \bibfield{postnote} or if the \bibfield{postnote} is not a number or range.
\item[separate] Separate the \bibfield{pages}\slash \bibfield{pagetotal} from the \bibfield{postnote} if the latter contains a page range. The string \texttt{thiscite} is added to separate the two page ranges.
\end{valuelist}

\end{optionlist}

\subparagraph{\texttt{verbose-trad}} The citation styles of the \texttt{verbose-trad} family support the global option \opt{strict}.

\begin{optionlist}

\boolitem[false]{strict}

This option allows to restrict the use of the scholarly abbreviations <ibid.> and <op.~cit.> to avoid ambiguities. If the option is set to \texttt{true} these terms will only be used if the relevant work was cited in the same or previous footnote.

\end{optionlist}

\subparagraph{\texttt{reading}} The \texttt{reading} style supports a number of additional options, but these are not of general interest and can be found in the style example.


\paragraph{Internal}
\label{use:opt:pre:int}

The default settings of the following preamble options are controlled by bibliography and citation styles. Apart from the \opt{pagetracker} and \opt{$<$name$>$inits} options, which you may want to adapt, there is normally no need to set them explicitly.

\begin{optionlist}

\optitem[false]{pagetracker}{\opt{true}, \opt{false}, \opt{page}, \opt{spread}}

This option controls the page tracker which is required by the \cmd{ifsamepage} and \cmd{iffirstonpage} tests from \secref{aut:aux:tst}. The possible choices are:

\begin{valuelist}
\item[true] Enable the tracker in automatic mode. This is like \opt{spread} if \latex is in twoside mode, and like \opt{page} otherwise.
\item[false] Disable the tracker.
\item[page] Enable the tracker in page mode. In this mode, tracking works on a per"=page basis.
\item[spread] Enable the tracker in spread mode. In this mode, tracking works on a per"=spread (double page) basis.
\end{valuelist}

Note that this tracker is disabled in all floats unless explicitly requested with \opt{trackfloats}, see \secref{aut:cav:flt}.

\optitem[false]{citecounter}{\opt{true}, \opt{false}, \opt{context}}

This option controls the citation counter which is required by \cnt{citecounter} from \secref{aut:aux:tst}. The possible choices are:

\begin{valuelist}
\item[true] Enable the citation counter in global mode.
\item[false] Disable the citation counter.
\item[context] Enable the citation counter in context"=sensitive mode. In this mode, citations in footnotes and in the body text are counted independently.
\end{valuelist}

\optitem[false]{citetracker}{\opt{true}, \opt{false}, \opt{context}, \opt{strict}, \opt{constrict}}

This option controls the citation tracker which is required by the \cmd{ifciteseen} and \cmd{ifentryseen} tests from \secref{aut:aux:tst}. The possible choices are:

\begin{valuelist}
\item[true] Enable the tracker in global mode.
\item[false] Disable the tracker.
\item[context] Enable the tracker in context"=sensitive mode. In this mode, citations in footnotes and in the body text are tracked independently.
\item[strict] Enable the tracker in strict mode. In this mode, an item is only considered by the tracker if it appeared in a stand-alone citation, \ie if a single entry key was passed to the citation command.
\item[constrict] This mode combines the features of \opt{context} and \opt{strict}.
\end{valuelist}

Note that this tracker is disabled in all floats unless explicitly requested with \opt{trackfloats}, see \secref{aut:cav:flt}.
This option is also settable on a per-type or per-entry basis.

\optitem[false]{ibidtracker}{\opt{true}, \opt{false}, \opt{context}, \opt{strict}, \opt{constrict}}

This option controls the <ibidem> tracker which is required by the \cmd{ifciteibid} test from \secref{aut:aux:tst}. The possible choices are:

\begin{valuelist}
\item[true] Enable the tracker in global mode.
\item[false] Disable the tracker.
\item[context] Enable the tracker in context"=sensitive mode. In this mode, citations in footnotes and in the body text are tracked separately.
\item[strict] Enable the tracker in strict mode. In this mode, potentially ambiguous references are suppressed. A reference is considered ambiguous if either the current citation (the one including the <ibidem>) or the previous citation (the one the <ibidem> refers to) consists of a list of references.\footnote{For example, suppose the initial citation is Jones, \emph{Title}; Williams, \emph{Title} and the following one ibidem. From a technical point of view, it is fairly clear that the <ibidem> refers to <Williams> because this is the last reference processed by the previous citation command. To a human reader, however, this may not be obvious because the <ibidem> may also refer to both titles. The strict mode avoids such ambiguous references.}
\item[constrict] This mode combines the features of \opt{context} and \opt{strict}. It also keeps track of footnote numbers and detects potentially ambiguous references in footnotes in a stricter way than the \opt{strict} option. In addition to the conditions imposed by the \opt{strict} option, a reference in a footnote will only be considered as unambiguous if the current citation and the previous citation are given in the same footnote or in immediately consecutive footnotes.
\end{valuelist}

Note that this tracker is disabled in all floats unless explicitly requested with \opt{trackfloats}, see \secref{aut:cav:flt}.
This option is also settable on a per-type or per-entry basis.

\optitem[false]{opcittracker}{\opt{true}, \opt{false}, \opt{context}, \opt{strict}, \opt{constrict}}

This option controls the <opcit> tracker which is required by the \cmd{ifopcit} test from \secref{aut:aux:tst}. This feature is similar to the <ibidem> tracker, except that it tracks citations on a per-author/editor basis, \ie \cmd{ifopcit} will yield \texttt{true} if the cited item is the same as the last one by this author\slash editor. The possible choices are:

\begin{valuelist}
\item[true] Enable the tracker in global mode.
\item[false] Disable the tracker.
\item[context] Enable the tracker in context"=sensitive mode. In this mode, citations in footnotes and in the body text are tracked separately.
\item[strict] Enable the tracker in strict mode. In this mode, potentially ambiguous references are suppressed. See \kvopt{ibidtracker}{strict} for details.
\item[constrict] This mode combines the features of \opt{context} and \opt{strict}. See the explanation of \kvopt{ibidtracker}{constrict} for details.
\end{valuelist}

Note that this tracker is disabled in all floats unless explicitly requested with \opt{trackfloats}, see \secref{aut:cav:flt}.
This option is also settable on a per-type or per-entry basis.

\optitem[false]{loccittracker}{\opt{true}, \opt{false}, \opt{context}, \opt{strict}, \opt{constrict}}

This option controls the <loccit> tracker which is required by the \cmd{ifloccit} test from \secref{aut:aux:tst}. This feature is similar to the <opcit> tracker except that it also checks whether the \prm{postnote} arguments match, \ie \cmd{ifloccit} will yield \texttt{true} if the citation refers to the same page cited before. The possible choices are:

\begin{valuelist}
\item[true] Enable the tracker in global mode.
\item[false] Disable the tracker.
\item[context] Enable the tracker in context"=sensitive mode. In this mode, citations in footnotes and in the body text are tracked separately.
\item[strict] Enable the tracker in strict mode. In this mode, potentially ambiguous references are suppressed. See \kvopt{ibidtracker}{strict} for details. In addition to that, this mode also checks if the \prm{postnote} argument is numerical (based on \cmd{ifnumerals} from \secref{aut:aux:tst}).
\item[constrict] This mode combines the features of \opt{context} and \opt{strict}. See the explanation of \kvopt{ibidtracker}{constrict} for details. In addition to that, this mode also checks if the \prm{postnote} argument is numerical (based on \cmd{ifnumerals} from \secref{aut:aux:tst}).
\end{valuelist}

Note that this tracker is disabled in all floats unless explicitly requested with \opt{trackfloats}, see \secref{aut:cav:flt}.
This option is also settable on a per-type or per-entry basis.

\optitem[false]{idemtracker}{\opt{true}, \opt{false}, \opt{context}, \opt{strict}, \opt{constrict}}

This option controls the <idem> tracker which is required by the \cmd{ifciteidem} test from \secref{aut:aux:tst}. The possible choices are:

\begin{valuelist}
\item[true] Enable the tracker in global mode.
\item[false] Disable the tracker.
\item[context] Enable the tracker in context"=sensitive mode. In this mode, citations in footnotes and in the body text are tracked separately.
\item[strict] This is an alias for \texttt{true}, provided only for consistency with the other trackers. Since <idem> replacements do not get ambiguous in the same way as <ibidem> or <op.~cit.>, the \texttt{strict} tracking mode does not apply to them.
\item[constrict] This mode is similar to \opt{context} with one additional condition: a reference in a footnote will only be considered as unambiguous if the current citation and the previous citation are given in the same footnote or in immediately consecutive footnotes.
\end{valuelist}

Note that this tracker is disabled in all floats unless explicitly requested with \opt{trackfloats}, see \secref{aut:cav:flt}.
This option is also settable on a per-type or per-entry basis.

\boolitem[false]{trackfloats}

Whether to enable citation tracking in floats. Citation tracking in floats can be tricky, so this option should only be enabled if absolutely necessary and the output should be scrutinised carefully, see also \secref{aut:cav:flt}.

\boolitem[true]{parentracker}

This option controls the parenthesis tracker which keeps track of nested parentheses and brackets. This information is used by \cmd{parentext} and \cmd{brackettext} from \secref{use:cit:txt}, \cmd{mkbibparens} and \cmd{mkbibbrackets} from \secref{aut:fmt:ich} and \cmd{bibopenparen}, \cmd{bibcloseparen}, \cmd{bibopenbracket}, \cmd{bibclosebracket} (also \secref{aut:fmt:ich}).

\intitem[3]{maxparens}

The maximum permitted nesting level of parentheses and brackets. If parentheses and brackets are nested deeper than this value, \biblatex will issue errors.

\boolitem[false]{$<$namepart$>$inits}

The option sets the \cmd{if$<$namepart$>$inits} test from \secref{aut:aux:tst}. \texttt{$<$namepart$>$} is any valid name part as defined in the data model by the \cmd{DeclareDatamodelConstant} command (\secref{aut:bbx:drv}). For the given name, for example, the option becomes \opt{giveninits}.
This option is also settable on a per-type, per-entry, per-namelist and per-name basis.

If \opt{giveninits} is set to \opt{true}, the default name formats will only render the given name initials and not the full given name. The standard styles only use the test \cmd{ifgiveninits} and hence only respond to the option \opt{giveninits}. Setting the option for a name part different from \texttt{given} has no effect on the default name formats.

Note that sorting and name uniqueness are not automatically affected by this option, that have to be requested explicitly via \cmd{DeclareSortingNamekeyTemplate} and the \opt{uniquename} option (or \cmd{DeclareUniquenameTemplate}), respectively. A warning will be issued if \opt{giveninits} is used together with \opt{uniquename} set to one of the \opt{full} values and \opt{uniquename} is automatically set to the corresponding \opt{init} value.

\boolitem[false]{terseinits}

This option controls the format of all initials generated by \biblatex. If enabled, initials are rendered using a terse format without dots and spaces. For example, the initials of Donald Ervin Knuth would be rendered as <D.~E.> by default, and as <DE> if this option is enabled. The option will affect the \cmd{ifterseinits} test from \secref{aut:aux:tst}. The option works by redefining some macros which control the format of initials. See \secref{use:cav:nam} for details.
This option is also settable on a per-type, per-entry, per-name and per-namelist basis.

\boolitem[false]{labelalpha}

Whether or not to provide the special fields \bibfield{labelalpha} and \bibfield{extraalpha}, see \secref{aut:bbx:fld} for details.
This option is also settable on a per-type and per-entry basis. See also \opt{maxalphanames} and \opt{minalphanames}. Table \ref{use:opt:tab1} summarises the various \opt{extra*} disambiguation counters and what they track.

\intitem[3]{maxalphanames}

Similar to the \opt{maxnames} option but customizes the format of the \bibfield{labelalpha} field.

\intitem[1]{minalphanames}

Similar to the \opt{minnames} option but customizes the format of the \bibfield{labelalpha} field.

\boolitem[false]{labelnumber}

Whether or not to provide the special field \bibfield{labelnumber}, see \secref{aut:bbx:fld} for details.
This option is also settable on a per-type and per-entry basis.

\boolitem[false]{noroman}

Whether or not to try to parse roman numerals encountered in integer fields for sorting purposes. Since \biber also tries to parse alphanumeric values when sorting integer fields, this roman numeral parsing can be a problem when, for example, <C> is encountered as this could be a roman numeral or a simple alphanumeric string which would have a different integer value depending on how it was parsed. It is likely that this is most useful on a per-entry basis for entries that have, for example, a \bibfield{volume} field with values such as <A>, <B>, <C>, <D> which should not be parsed as roman numerals since this would give incorrect integer values for <C> and <D>.

This option is also settable on a per-type and per-entry basis.

\boolitem[false]{labeltitle}

Whether or not to provide the special field \bibfield{extratitle}, see \secref{aut:bbx:fld} for details. Note that the special field \bibfield{labeltitle} is always provided and this option controls rather whether \bibfield{labeltitle} is used to generate \bibfield{extratitle} information. This option is also settable on a per-type and per-entry basis. Table \ref{use:opt:tab1} summarises the various \opt{extra*} disambiguation counters and what they track.

\boolitem[false]{labeltitleyear}

Whether or not to provide the special field \bibfield{extratitleyear}, see \secref{aut:bbx:fld} for details. Note that the special field \bibfield{labeltitle} is always provided and this option controls rather whether \bibfield{labeltitle} is used to generate \bibfield{extratitleyear} information. This option is also settable on a per-type and per-entry basis. Table \ref{use:opt:tab1} summarises the various \opt{extra*} disambiguation counters and what they track.

\boolitem[false]{labeldateparts}

Whether or not to provide the special fields \bibfield{labelyear}, \bibfield{labelmonth}, \bibfield{labelday}, \bibfield{labelendyear}, \bibfield{labelendmonth}, \bibfield{labelendday}, \bibfield{labelhour}, \bibfield{labelendhour}, \bibfield{labelminute}, \bibfield{labelendminute}, \bibfield{labelsecond}, \bibfield{labelendsecond}, \bibfield{labelseason}, \bibfield{labelendseason}, \bibfield{labeltimezone}, \bibfield{labelendtimeone} and \bibfield{extradate}, see \secref{aut:bbx:fld} for details.
This option is also settable on a per-type and per-entry basis. Table \ref{use:opt:tab1} summarises the various \opt{extra*} disambiguation counters and what they track.

\begin{table}
\caption{Work Uniqueness options}
\label{use:opt:wu}
\footnotesize
\ttfamily
\tablesetup
\begin{tabularx}{\textwidth}{XXX}
\toprule
\multicolumn{1}{@{}H}{Option} &
\multicolumn{1}{@{}H}{Test} &
\multicolumn{1}{@{}H}{Tracks} \\
\cmidrule(r){1-1}\cmidrule(r){2-2}\cmidrule(r){3-3}
singletitle & \cmd{ifsingletitle} & labelname\\
uniquetitle & \cmd{ifuniquetitle} & labeltitle\\
uniquebaretitle & \cmd{ifuniquebaretitle} & labeltitle {\rmfamily when} labelname {\rmfamily is null}\\
uniquework  & \cmd{ifuniquework}  & labelname+labeltitle\\
\bottomrule
\end{tabularx}
\end{table}

\boolitem[false]{singletitle}

Whether or not to provide the data required by the \cmd{ifsingletitle} test, see \secref{aut:aux:tst} for details. See \tabref{use:opt:wu} for details on what determines the data for this test.
This option is also settable on a per-type and per-entry basis.

\boolitem[false]{uniquetitle}

Whether or not to provide the data required by the \cmd{ifuniquetitle} test, see \secref{aut:aux:tst} for details. See \tabref{use:opt:wu} for details on what determines the data for this test.
This option is also settable on a per-type and per-entry basis.

\boolitem[false]{uniquebaretitle}

Whether or not to provide the data required by the \cmd{ifuniquebaretitle} test, see \secref{aut:aux:tst} for details. See \tabref{use:opt:wu} for details on what determines the data for this test.
This option is also settable on a per-type and per-entry basis.

\boolitem[false]{uniquework}

Whether or not to provide the data required by the \cmd{ifuniquework} test, see \secref{aut:aux:tst} for details. See \tabref{use:opt:wu} for details on what determines the data for this test.
This option is also settable on a per-type and per-entry basis.

\boolitem[false]{uniqueprimaryauthor}

Whether or not to provide the data required by the \cmd{ifuniqueprimaryauthor} test, see \secref{aut:aux:tst} for details.
This option is also settable on a per-type and per-entry basis.

\optitem[false]{uniquename}{\opt{true}, \opt{false}, \opt{init}, \opt{full}, \opt{allinit}, \opt{allfull},
\opt{mininit}, \opt{minfull}}

Whether or not to update the \cnt{uniquename} counter, see \secref{aut:aux:tst} for details. This feature will disambiguate individual names in the \bibfield{labelname} list. This option is also settable on a per-type, per-entry, per-namelist and per-name basis. The possible choices are:

\begin{valuelist}
\item[true] An alias for \opt{full}.
\item[false] Disable this feature.
\item[init] Disambiguate using initials only.
\item[full] Disambiguate using initials or full names, as required.
\item[allinit] Similar to \opt{init} but disambiguates all names in the \bibfield{labelname} list, beyond \opt{maxnames}\slash \opt{minnames}\slash \opt{uniquelist}.
\item[allfull] Similar to \opt{full} but disambiguates all names in the \bibfield{labelname} list, beyond \opt{maxnames}\slash \opt{minnames}\slash \opt{uniquelist}.
\item[mininit] A variant of \texttt{init} which only disambiguates names in identical lists of base nameparts (by default, lists of family names).
\item[minfull] A variant of \texttt{full} which only disambiguates names in identical lists of base nameparts (by default, lists of family names).
\end{valuelist}
%
Note that the \opt{uniquename} option will also affect \opt{uniquelist}, the \cmd{ifsingletitle} test, and the \bibfield{extradate} and \bibfield{extraname} fields. See \secref{aut:cav:amb} for further details and practical examples.

\optitem[false]{uniquelist}{\opt{true}, \opt{false}, \opt{minyear}}

Whether or not to update the \cnt{uniquelist} counter, see \secref{aut:aux:tst} for details. This feature will disambiguate the \bibfield{labelname} list if it has become ambiguous after \opt{maxnames}\slash \opt{minnames} truncation. Essentially, it overrides \opt{maxnames}\slash \opt{minnames} on a per-field basis. This option is also settable on a per-type, per-entry and per-namelist basis. The possible choices are:

\begin{valuelist}
\item[true] Disambiguate the \bibfield{labelname} list.
\item[false] Disable this feature.
\item[minyear] Disambiguate the \bibfield{labelname} list only if the truncated list is identical to another one with the same \bibfield{labelyear}. This mode of operation is useful for author-year styles and requires \kvopt{labeldateparts}{true}.
\end{valuelist}
%
Note that the \opt{uniquelist} option will also affect the \cmd{ifsingletitle} test and the \bibfield{extradate} and \bibfield{extraname} fields. See \secref{aut:cav:amb} for further details and practical examples.

\boolitem[false]{nohashothers}

By default, name lists which are truncated with <et al>--either explicitly by <and others> in the data source or the \opt{uniquelist} and \opt{min/maxnames} options--result in different name list hashes (and therefore different \opt{extraname} and \opt{extradate} values) and different sorting. This option allows this behaviour to be tuned. When set to \prm{true}, \biber ignores <et al> truncations for the purposes of generating name list hashes. Consider:

\begin{lstlisting}{}
Jones 1972
Jones/and others 1972
Smith 2000
Smith/Vogel/Beast/Tremble 2000
\end{lstlisting}
%
With \kvopt{maxnames}{3}, \kvopt{minnames}{1}, \kvopt{nohashothers}{false}, the result would be:

\begin{lstlisting}{}
  Jones 1972
  Jones et al. 1972
  Smith 2000
  Smith et al. 2000
\end{lstlisting}
%
Whereas with \kvopt{maxnames}{3}, \kvopt{minnames}{1}, \kvopt{nohashothers}{true}, the result would be:

\begin{lstlisting}{}
  Jones 1972a
  Jones et al. 1972b
  Smith 2000a
  Smith et al. 2000b
\end{lstlisting}

If desired, this could be further simplified by removing the <et al.\@>
to obtain:

\begin{lstlisting}{}
  Jones 1972a
  Jones 1972b
  Smith 2000a
  Smith 2000b
\end{lstlisting}
%
Note that the \opt{nohashothers} option will affect the \bibfield{extradate} and \bibfield{extraname} fields.

This option is also settable on a per-type, per-entry and per-namelist basis.

\boolitem[false]{nosortothers}

The option  has a related to effect to \opt{nohashothers} but applies to sorting--the visible list of names (which is the \opt{minsortnames} value) used to determine sorting will ignore any truncation. This means that with \kvopt{nosortothers}{true}, the name lists:

\begin{lstlisting}{}
Jones, Smith
Jones, Smith et al
\end{lstlisting}
%
will sort exactly the same. The default setting of \opt{nosortothers} always sorts in the order shown in the example, that is, by default, truncated names lists always sort after any name lists identical to the point of truncation.

This option is also settable on a per-type, per-entry and per-namelist basis.
\end{optionlist}

\begin{table}
\caption{Disambiguation counters}
\label{use:opt:tab1}
\footnotesize
\ttfamily
\tablesetup
\begin{tabularx}{\textwidth}{XXXX}
\toprule
\multicolumn{1}{@{}H}{Option} &
\multicolumn{1}{@{}H}{Enabled field(s)} &
\multicolumn{1}{@{}H}{Enabled counter} &
\multicolumn{1}{@{}H}{Counter tracks} \\
\cmidrule(r){1-1}\cmidrule(r){2-2}\cmidrule(r){3-3}\cmidrule{4-4}
labelalpha      & labelalpha       & extraalpha     &  label\\
labeldateparts  & labelyear        & extradate      &  labelname+\\
                & labelmonth       &                &  labelyear\\
                & labelday         &                &  \\
                & labelendyear     &                &  \\
                & labelendmonth    &                &  \\
                & labelendday      &                &  \\
                & labelhour        &                &  \\
                & labelminute      &                &  \\
                & labelsecond      &                &  \\
                & labelendhour     &                &  \\
                & labelendminute   &                &  \\
                & labelendsecond   &                &  \\
                & labelseason      &                &  \\
                & labelendseason   &                &  \\
                & labeltimezone    &                &  \\
                & labelendtimezone &                &  \\
labeltitle      & {\rmfamily ---}  & extratitle     &  labelname+labeltitle\\
labeltitleyear  & {\rmfamily ---}  & extratitleyear &  labeltitle+labelyear\\
{\rmfamily ---} & {\rmfamily ---}  & extraname      &  labelname\\
\bottomrule
\end{tabularx}
\end{table}

\subsubsection{Entry Options}
\label{use:opt:bib}

Entry options are package options which determine how bibliography data entries are handled. They may be set at various scopes defined below.

\paragraph{Preamble/Type/Entry Options}
\label{use:opt:bib:hyb}

The following options are settable on a per"=type basis or on a per"=entry in the \bibfield{options} field. In addition to that, they may also be used in the optional argument to \cmd{usepackage} as well as in the configuration file and the document preamble. This is useful if you want to change the default behaviour globally.

\begin{optionlist}

\boolitem[true]{useauthor}

Whether the \bibfield{author} is used in labels and considered during sorting. This may be useful if an entry includes an \bibfield{author} field but is usually not cited by author for some reason. Setting \kvopt{useauthor}{false} does not mean that the \bibfield{author} is ignored completely. It means that the \bibfield{author} is not used in labels and ignored during sorting. The entry will then be alphabetized by \bibfield{editor} or \bibfield{title}. With the standard styles, the \bibfield{author} is printed after the title in this case. See also \secref{use:srt}.
This option is also settable on a per-type and per-entry basis.

\boolitem[true]{useeditor}

Whether the \bibfield{editor} replaces a missing \bibfield{author} in labels and during sorting. This may be useful if an entry includes an \bibfield{editor} field but is usually not cited by editor. Setting \kvopt{useeditor}{false} does not mean that the \bibfield{editor} is ignored completely. It means that the \bibfield{editor} does not replace a missing \bibfield{author} in labels and during sorting. The entry will then be alphabetized by \bibfield{title}. With the standard styles, the \bibfield{editor} is printed after the title in this case. See also \secref{use:srt}.
This option is also settable on a per-type and per-entry basis.

\boolitem[false]{usetranslator}

Whether the \bibfield{translator} replaces a missing \bibfield{author}\slash \bibfield{editor} in labels and during sorting. Setting \kvopt{usetranslator}{true} does not mean that the \bibfield{translator} overrides the \bibfield{author}\slash \bibfield{editor}. It means that the \bibfield{translator} is considered as a fallback if the \bibfield{author}\slash \bibfield{editor} is missing or if \opt{useauthor} and \opt{useeditor} are set to \texttt{false}. In other words, in order to cite a book by translator rather than by author, you need to set the following options:
This option is also settable on a per-type and per-entry basis.

\begin{lstlisting}[style=bibtex]{}
@Book{...,
  options    = {useauthor=false,usetranslator=true},
  author     = {...},
  translator = {...},
  ...
\end{lstlisting}
%
With the standard styles, the \bibfield{translator} is printed after the title by default. See also \secref{use:srt}.

\boolitem[true]{use$<$name$>$}

As per \opt{useauthor}, \opt{useeditor} and \opt{usetranslator}, all name lists defined in the data model have an option controlling their behaviour in sorting and labelling automatically
defined. Global, per-type and per-entry options called <use$<$name$>$>are automatically created.

\boolitem[false]{useprefix}

Whether the default data model name part <prefix> (von, van, of, da, de, della, etc.) is considered when:

\begin{itemize}
\item Printing the family name in citations
\item Sorting
\item Generation of certain types of labels
\item Generating name uniqueness information
\item Formatting aspects of the bibliography
\end{itemize}
%
For example, if this option is enabled, \biblatex precedes the family name with the prefix---Ludwig van Beethoven would be cited as van Beethoven and alphabetized as Van Beethoven, Ludwig. If this option is disabled (the default), he is cited as Beethoven and alphabetized as Beethoven, Ludwig van instead.
This option is also settable on a per-type scope. With \biblatexml datasources and the \bibtex extended name format supported by \biber (see \secref{use:enf}), this is also settable on per-namelist and per-name scopes.

\optitem{indexing}{\opt{true}, \opt{false}, \opt{cite}, \opt{bib}}

The \opt{indexing} option is also settable per-type or per-entry basis. See \secref{use:opt:pre:gen} for details.

\boolitem[false]{skipbib}

If this option is enabled, the entry is excluded from the bibliography but it may still be cited.
This option is also settable on a per-type basis.

\boolitem[false]{skipbiblist}

If this option is enabled, the entry is excluded from bibliography lists. It is still included in the bibliography and it may also be cited by shorthand etc.
This option is also settable on a per-type basis.

\boolitem[false]{skiplab}

If this option is enabled, \biblatex will not assign any labels to the entry. It is not required for normal operation. Use it with care. If enabled, \biblatex can not guarantee unique citations for the respective entry and citations styles which require labels may fail to create valid citations for the entry.
This option is also settable on a per-type basis.

\boolitem[false]{dataonly}

Setting this option is equivalent to \kvopt{uniquename}{false}, \kvopt{uniquelist}{false, }\opt{skipbib}, \opt{skipbiblist}, and \opt{skiplab}. It is not required for normal operation. Use it with care.
This option is also settable on a per-type basis.

\paragraph{Entry Only Options}
\label{use:opt:bib:entry}

The following options are settable only on a per"=entry in the \bibfield{options} field. They are not available globally or per"=type.

\valitem{labelnamefield}{fieldname}

Specifies the field to consider first when looking for a \bibfield{labelname} candidate. It is essentially prepended to the search list created by \cmd{DeclareLabelname} for just this entry.

\valitem{labeltitlefield}{fieldname}

Specifies the field to consider first when looking for a \bibfield{labeltitle} candidate. It is essentially prepended to the search list created by \cmd{DeclareLabeltitle} for just this entry.

\end{optionlist}

\subsubsection{Legacy Options}

The following legacy option may be used globally in the optional argument to \cmd{documentclass} or locally in the optional argument to \cmd{usepackage}:

\begin{optionlist}

\legitem{openbib}\DeprecatedMark  This option is provided for backwards compatibility with the standard LaTeX document classes. \opt{openbib} is similar to \kvopt{block}{par}.

\end{optionlist}

\subsection{Global Customization}
\label{use:cfg}

Apart from writing new citation and bibliography styles, there are numerous ways to customize the styles which come with this package. Customization will usually take place in the preamble, but there is also a configuration file for permanent adaptions. The configuration file may also be used to initialize the package options to a value different from the package default.

\subsubsection{Configuration File}
\label{use:cfg:cfg}

If available, this package will load the configuration file \path{biblatex.cfg}. This file is read at the end of the package, immediately after the citation and bibliography styles have been loaded.

\subsubsection{Setting Package Options}
\label{use:cfg:opt}

The load-time package options in \secref{use:opt:ldt} must be given in the optional argument to \cmd{usepackage}. The package options in \secref{use:opt:pre} may also be given in the preamble. The options are executed with the following command:

\begin{ltxsyntax}

\cmditem{ExecuteBibliographyOptions}[entrytype, \dots]{key=value, \dots}

This command may also be used in the configuration file to modify the default setting of a package option. Certain options are also settable on a per-type basis. In this case, the optional \prm{entrytype} argument specifies the entry type. The \prm{entrytype} argument may be a comma"=separated list of values.

\end{ltxsyntax}

\subsection{Standard Styles}
\label{use:xbx}

This section provides a short description of all bibliography and citation styles which come with the \biblatex package. Each style is further illustrated in a style example which is linked in the right margin. The local link may not be available if this document does not reside in the expected folder structure. If you want to write your own styles, see \secref{aut}.

\subsubsection{Citation Styles}
\label{use:xbx:cbx}

The citation styles which come with this package implement several common citation schemes. All standard styles cater for the \bibfield{shorthand} field and support hyperlinks as well as indexing.

\begin{marglist}

\item[numeric]\seestyleexample{30-style-numeric-biber}
This style implements a numeric citation scheme similar to the standard bibliographic facilities of \latex. It should be employed in conjunction with a numeric bibliography style which prints the corresponding labels in the bibliography. It is intended for in-text citations. The style will set the following package options at load time: \kvopt{autocite}{inline}, \kvopt{labelnumber}{true}. This style also provides an additional preamble option called \opt{subentry} which affects the handling of entry sets. If this option is disabled, citations referring to a member of a set will point to the entire set. If it is enabled, the style supports citations like [5c] which point to a subentry in a set (the third one in this example). See the style example for details.

\item[numeric-comp]\seestyleexample{31-style-numeric-comp-biber}
A compact variant of the \texttt{numeric} style which prints a list of more than two consecutive numbers as a range. This style is similar to the \sty{cite} package and the \opt{sort\&compress} option of the \sty{natbib} package in numerical mode. For example, instead of [8, 3, 1, 7, 2] this style would print [1--3, 7, 8]. It is intended for in-text citations. The style will set the following package options at load time: \kvopt{autocite}{inline}, \kvopt{sortcites}{true}, \kvopt{labelnumber}{true}. It also provides the \opt{subentry} and \opt{subentrycomp} options.

\item[numeric-verb]\seestyleexample{32-style-numeric-verb-biber}
A verbose variant of the \texttt{numeric} style. The difference affects the handling of a list of citations and is only apparent when multiple entry keys are passed to a single citation command. For example, instead of [2, 5, 6] this style would print [2]; [5]; [6]. It is intended for in-text citations. The style will set the following package options at load time: \kvopt{autocite}{inline}, \kvopt{labelnumber}{true}. It also provides the \opt{subentry} option.

\item[alphabetic]\seestyleexample{40-style-alphabetic-biber}
This style implements an alphabetic citation scheme similar to the \path{alpha.bst} style of traditional \bibtex. The alphabetic labels resemble a compact author"=year style to some extent, but the way they are employed is similar to a numeric citation scheme. For example, instead of Jones 1995 this style would use the label [Jon95]. Jones and Williams 1986 would be rendered as [JW86]. This style should be employed in conjunction with an alphabetic bibliography style which prints the corresponding labels in the bibliography. It is intended for in-text citations. The style will set the following package options at load time: \kvopt{autocite}{inline}, \kvopt{labelalpha}{true}. This style also provides an additional preamble option called \opt{subentry} which affects the handling of entry sets. If this option is disabled, citations referring to a member of a set will point to the entire set. If it is enabled, the style supports citations like [SGW(c)] which point to a subentry in a set (the third one in this example). See the style example for details.

\item[alphabetic-verb]\seestyleexample{41-style-alphabetic-verb-biber}
A verbose variant of the \texttt{alphabetic} style. The difference affects the handling of a list of citations and is only apparent when multiple entry keys are passed to a single citation command. For example, instead of [Doe92; Doe95; Jon98] this style would print [Doe92]; [Doe95]; [Jon98]. It is intended for in-text citations. The style will set the following package options at load time: \kvopt{autocite}{inline}, \kvopt{labelalpha}{true}. It also provides the subentry option.

\item[authoryear]\seestyleexample{50-style-authoryear-biber}
This style implements an author"=year citation scheme. If the bibliography contains two or more works by the same author which were all published in the same year, a letter is appended to the year. For example, this style would print citations such as Doe 1995a; Doe 1995b; Jones 1998. This style should be employed in conjunction with an author"=year bibliography style which prints the corresponding labels in the bibliography. It is primarily intended for in-text citations, but it could also be used with citations given in footnotes. The style will set the following package options at load time: \kvopt{autocite}{inline}, \kvopt{labeldateparts}{true}, \kvopt{uniquename}{full}, \kvopt{uniquelist}{true}.

\item[authoryear-comp]\seestyleexample{52-style-authoryear-comp-biber}
A compact variant of the \texttt{authoryear} style which prints the author only once if subsequent references passed to a single citation command share the same author. If they share the same year as well, the year is also printed only once. For example, instead of Doe 1995b; Doe 1992; Jones 1998; Doe 1995a this style would print Doe 1992, 1995a,b; Jones 1998. It is primarily intended for in-text citations, but it could also be used with citations given in footnotes. The style will set the following package options at load time: \kvopt{autocite}{inline}, \kvopt{sortcites}{true}, \kvopt{labeldateparts}{true}, \kvopt{uniquename}{full}, \kvopt{uniquelist}{true}.

\item[authoryear-ibid]\seestyleexample{51-style-authoryear-ibid-biber}
A variant of the \texttt{authoryear} style which replaces repeated citations by the abbreviation \emph{ibidem} unless the citation is the first one on the current page or double-page spread, or the \emph{ibidem} would be ambiguous in the sense of the package option \kvopt{ibidtracker}{constrict}. The style will set the following package options at load time: \kvopt{autocite}{inline}, \kvopt{labeldateparts}{true}, \kvopt{uniquename}{full}, \kvopt{uniquelist}{true}, \kvopt{ibidtracker}{constrict}, \kvopt{pagetracker}{true}. This style also provides an additional preamble option called \opt{ibidpage}. See the style example for details.

\item[authoryear-icomp]\seestyleexample{53-style-authoryear-icomp-biber}
A style combining \texttt{authoryear-comp} and \texttt{authoryear-ibid}. The style will set the following package options at load time: \kvopt{autocite}{inline}, \kvopt{labeldateparts}{true}, \kvopt{uniquename}{full}, \kvopt{uniquelist}{true}, \kvopt{ibidtracker}{constrict}, \kvopt{pagetracker}{true}, \kvopt{sortcites}{true}. This style also provides an additional preamble option called \opt{ibidpage}. See the style example for details.

\item[authortitle]\seestyleexample{60-style-authortitle-biber}
This style implements a simple author"=title citation scheme. It will make use of the \bibfield{shorttitle} field, if available. It is intended for citations given in footnotes. The style will set the following package options at load time: \kvopt{autocite}{footnote}, \kvopt{uniquename}{full}, \kvopt{uniquelist}{true}.

\item[authortitle-comp]\seestyleexample{62-style-authortitle-comp-biber}
A compact variant of the \texttt{authortitle} style which prints the author only once if subsequent references passed to a single citation command share the same author. For example, instead of Doe, \emph{First title}; Doe, \emph{Second title} this style would print Doe, \emph{First title}, \emph{Second title}. It is intended for citations given in footnotes. The style will set the following package options at load time: \kvopt{autocite}{footnote}, \kvopt{sortcites}{true}, \kvopt{uniquename}{full}, \kvopt{uniquelist}{true}.

\item[authortitle-ibid]\seestyleexample{61-style-authortitle-ibid-biber}
A variant of the \texttt{authortitle} style which replaces repeated citations by the abbreviation \emph{ibidem} unless the citation is the first one on the current page or double-page spread, or the \emph{ibidem} would be ambiguous in the sense of the package option \kvopt{ibidtracker}{constrict}. It is intended for citations given in footnotes. The style will set the following package options at load time: \kvopt{autocite}{footnote}, \kvopt{uniquename}{full}, \kvopt{uniquelist}{true}, \kvopt{ibidtracker}{constrict}, \kvopt{pagetracker}{true}. This style also provides an additional preamble option called \opt{ibidpage}. See the style example for details.

\item[authortitle-icomp]\seestyleexample{63-style-authortitle-icomp-biber}
A style combining the features of \texttt{authortitle-comp} and \texttt{authortitle-ibid}. The style will set the following package options at load time: \kvopt{autocite}{footnote}, \kvopt{uniquename}{full}, \kvopt{uniquelist}{true}, \kvopt{ibidtracker}{constrict}, \kvopt{pagetracker}{true}, \kvopt{sortcites}{true}. This style also provides an additional preamble option called \opt{ibidpage}. See the style example for details.

\item[authortitle-terse]\seestyleexample{64-style-authortitle-terse-biber}
A terse variant of the \texttt{authortitle} style which only prints the title if the bibliography contains more than one work by the respective author\slash editor. This style will make use of the \bibfield{shorttitle} field, if available. It is suitable for in-text citations as well as citations given in footnotes. The style will set the following package options at load time: \kvopt{autocite}{inline}, \kvopt{singletitle}{true}, \kvopt{uniquename}{full}, \kvopt{uniquelist}{true}.

\item[authortitle-tcomp]\seestyleexample{65-style-authortitle-tcomp-biber}
A style combining the features of \texttt{authortitle-comp} and \texttt{authortitle-terse}. This style will make use of the \bibfield{shorttitle} field, if available. It is suitable for in-text citations as well as citations given in footnotes. The style will set the following package options at load time: \kvopt{autocite}{inline}, \kvopt{sortcites}{true}, \kvopt{singletitle}{true}, \kvopt{uniquename}{full}, \kvopt{uniquelist}{true}.

\item[authortitle-ticomp]\seestyleexample{66-style-authortitle-ticomp-biber}
A style combining the features of \texttt{authortitle-icomp} and \texttt{authortitle-terse}. In other words: a variant of the \texttt{authortitle-tcomp} style with an \emph{ibidem} feature. This style is suitable for in-text citations as well as citations given in footnotes. It will set the following package options at load time: \kvopt{autocite}{inline}, \kvopt{ibidtracker}{constrict}, \kvopt{pagetracker}{true}, \kvopt{sortcites}{true}, \kvopt{singletitle}{true}, \kvopt{uniquename}{full}, \kvopt{uniquelist}{true}. This style also provides an additional preamble option called \opt{ibidpage}. See the style example for details.

\item[verbose]\seestyleexample{70-style-verbose-biber}
A verbose citation style which prints a full citation similar to a bibliography entry when an entry is cited for the first time, and a short citation afterwards. If available, the \bibfield{shorttitle} field is used in all short citations. If the \bibfield{shorthand} field is defined, the shorthand is introduced on the first citation and used as the short citation thereafter. This style may be used without a list of references and shorthands since all bibliographic data is provided on the first citation. It is intended for citations given in footnotes. The style will set the following package options at load time: \kvopt{autocite}{footnote}, \kvopt{citetracker}{context}. This style also provides an additional preamble option called \opt{citepages}. See the style example for details.

\item[verbose-ibid]\seestyleexample{71-style-verbose-ibid-biber}
A variant of the \texttt{verbose} style which replaces repeated citations by the abbreviation \emph{ibidem} unless the citation is the first one on the current page or double-page spread, or the \emph{ibidem} would be ambiguous in the sense of \kvopt{ibidtracker}{strict}. This style is intended for citations given in footnotes. The style will set the following package options at load time: \kvopt{autocite}{footnote}, \kvopt{citetracker}{context}, \kvopt{ibidtracker}{constrict}, \kvopt{pagetracker}{true}. This style also provides additional preamble options called \opt{ibidpage} and \opt{citepages}. See the style example for details.

\item[verbose-note]\seestyleexample{72-style-verbose-note-biber}
This style is similar to the \texttt{verbose} style in that it prints a full citation similar to a bibliography entry when an entry is cited for the first time, and a short citation afterwards. In contrast to the \texttt{verbose} style, the short citation is a pointer to the footnote with the full citation. If the bibliography contains more than one work by the respective author\slash editor, the pointer also includes the title. If available, the \bibfield{shorttitle} field is used in all short citations. If the \bibfield{shorthand} field is defined, it is handled as with the \texttt{verbose} style. This style may be used without a list of references and shorthands since all bibliographic data is provided on the first citation. It is exclusively intended for citations given in footnotes. The style will set the following package options at load time: \kvopt{autocite}{footnote}, \kvopt{citetracker}{context}, \kvopt{singletitle}{true}. This style also provides additional preamble options called \opt{pageref} and \opt{citepages}. See the style example for details.

\item[verbose-inote]\seestyleexample{73-style-verbose-inote-biber}
A variant of the \texttt{verbose"=note} style which replaces repeated citations by the abbreviation \emph{ibidem} unless the citation is the first one on the current page or double-page spread, or the \emph{ibidem} would be ambiguous in the sense of \kvopt{ibidtracker}{strict}. This style is exclusively intended for citations given in footnotes. It will set the following package options at load time: \kvopt{autocite}{footnote}, \kvopt{citetracker}{context}, \kvopt{ibidtracker}{constrict}, \kvopt{singletitle}{true}, \kvopt{pagetracker}{true}. This style also provides additional preamble options called \opt{ibidpage}, \opt{pageref}, and \opt{citepages}. See the style example for details.

\item[verbose-trad1]\seestyleexample{74-style-verbose-trad1-biber}
This style implements a traditional citation scheme. It is similar to the \texttt{verbose} style in that it prints a full citation similar to a bibliography entry when an item is cited for the first time, and a short citation afterwards. Apart from that, it uses the scholarly abbreviations \emph{ibidem}, \emph{idem}, \emph{op.~cit.}, and \emph{loc.~cit.} to replace recurrent authors, titles, and page numbers in repeated citations in a special way. If the \bibfield{shorthand} field is defined, the shorthand is introduced on the first citation and used as the short citation thereafter. This style may be used without a list of references and shorthands since all bibliographic data is provided on the first citation. It is intended for citations given in footnotes. The style will set the following package options at load time: \kvopt{autocite}{footnote}, \kvopt{citetracker}{context}, \kvopt{ibidtracker}{constrict}, \kvopt{idemtracker}{constrict}, \kvopt{opcittracker}{context}, \kvopt{loccittracker}{context}. This style also provides additional preamble options called \opt{ibidpage}, \opt{strict}, and \opt{citepages}. See the style example for details.

\item[verbose-trad2]\seestyleexample{75-style-verbose-trad2-biber}
Another traditional citation scheme. It is also similar to the \texttt{verbose} style but uses scholarly abbreviations like \emph{ibidem} and \emph{idem} in repeated citations. In contrast to the \texttt{verbose-trad1} style, the logic of the \emph{op.~cit.} abbreviations is different in this style and \emph{loc.~cit.} is not used at all. It is in fact more similar to \texttt{verbose-ibid} and \texttt{verbose-inote} than to \texttt{verbose-trad1}. The style will set the following package options at load time: \kvopt{autocite}{footnote}, \kvopt{citetracker}{context}, \kvopt{ibidtracker}{constrict}, \kvopt{idemtracker}{constrict}. This style also provides additional preamble options called \opt{ibidpage}, \opt{strict}, and \opt{citepages}. See the style example for details.

\item[verbose-trad3]\seestyleexample{76-style-verbose-trad3-biber}
Yet another traditional citation scheme. It is similar to the \texttt{verbose-trad2} style but uses the scholarly abbreviations \emph{ibidem} and \emph{op.~cit.} in a slightly different way. The style will set the following package options at load time: \kvopt{autocite}{footnote}, \kvopt{citetracker}{context}, \kvopt{ibidtracker}{constrict}, \kvopt{loccittracker}{constrict}. This style also provides additional preamble options called \opt{strict} and \opt{citepages}. See the style example for details.

\item[reading]\seestyleexample{80-style-reading-biber}
A citation style which goes with the bibliography style by the same name. It simply loads the \texttt{authortitle} style.

\end{marglist}

The following citation styles are special purpose styles. They are not intended for the final version of a document:

\begin{marglist}

\item[draft]\seestyleexample{81-style-draft-biber}
A draft style which uses the entry keys in citations. The style will set the following package options at load time: \kvopt{autocite}{plain}.

\item[debug]\seestyleexample{82-style-debug-biber}
This style prints the entry key rather than some kind of label. It is intended for debugging only and will set the following package options at load time: \kvopt{autocite}{plain}.

\end{marglist}

\subsubsection{Bibliography Styles}
\label{use:xbx:bbx}

All bibliography styles which come with this package use the same basic format for the individual bibliography entries. They only differ in the kind of label printed in the bibliography and the overall formatting of the list of references. There is a matching bibliography style for every citation style. Note that some bibliography styles are not mentioned below because they simply load a more generic style. For example, the bibliography style \texttt{authortitle-comp} will load the \texttt{authortitle} style.

\begin{marglist}

\item[numeric]\seestyleexample{30-style-numeric-biber}
This style prints a numeric label similar to the standard bibliographic facilities of \latex. It is intended for use in conjunction with a numeric citation style. Note that the \bibfield{shorthand} field overrides the default label. The style will set the following package options at load time: \kvopt{labelnumber}{true}. This style also provides an additional preamble option called \opt{subentry} which affects the formatting of entry sets. If this option is enabled, all members of a set are marked with a letter which may be used in citations referring to a set member rather than the entire set. See the style example for details.

\item[alphabetic]\seestyleexample{40-style-alphabetic-biber}
This style prints an alphabetic label similar to the \path{alpha.bst} style of traditional \bibtex. It is intended for use in conjunction with an alphabetic citation style. Note that the \bibfield{shorthand} field overrides the default label. The style will set the following package options at load time: \kvopt{labelalpha}{true}, \kvopt{sorting}{anyt}.

\item[authoryear]\seestyleexample{50-style-authoryear-biber}
This style differs from the other styles in that the publication date is not printed towards the end of the entry but rather after the author\slash editor. It is intended for use in conjunction with an author"=year citation style. Recurring author and editor names are replaced by a dash unless the entry is the first one on the current page or double-page spread. This style provides an additional preamble option called \opt{dashed} which controls this feature. It also provided a preamble option called \opt{mergedate}. See the style example for details. The style will set the following package options at load time: \kvopt{labeldateparts}{true}, \kvopt{sorting}{nyt}, \kvopt{pagetracker}{true}, \kvopt{mergedate}{true}.

\item[authortitle]\seestyleexample{60-style-authortitle-biber}
This style does not print any label at all. It is intended for use in conjunction with an author"=title citation style. Recurring author and editor names are replaced by a dash unless the entry is the first one on the current page or double-page spread. This style also provides an additional preamble option called \opt{dashed} which controls this feature. See the style example for details. The style will set the following package options at load time: \kvopt{pagetracker}{true}.

\item[verbose]\seestyleexample{70-style-verbose-biber}
This style is similar to the \texttt{authortitle} style. It also provides an additional preamble option called \opt{dashed}. See the style example for details. The style will set the following package options at load time: \kvopt{pagetracker}{true}.

\item[reading]\seestyleexample{80-style-reading-biber}
This special bibliography style is designed for personal reading lists, annotated bibliographies, and similar applications. It optionally includes the fields \bibfield{annotation}, \bibfield{abstract}, \bibfield{library}, and \bibfield{file} in the bibliography. If desired, it also adds various kinds of short headers to the bibliography. This style also provides the additional preamble options \opt{entryhead}, \opt{entrykey}, \opt{annotation}, \opt{abstract}, \opt{library}, and \opt{file} which control whether or not the corresponding items are printed in the bibliography. See the style example for details. See also \secref{use:use:prf}. The style will set the following package options at load time: \kvopt{loadfiles}{true}, \kvopt{entryhead}{true}, \kvopt{entrykey}{true}, \kvopt{annotation}{true}, \kvopt{abstract}{true}, \kvopt{library}{true}, \kvopt{file}{true}.

\end{marglist}

The following bibliography styles are special purpose styles. They are not intended for the final version of a document:

\begin{marglist}

\item[draft]\seestyleexample{81-style-draft-biber}
This draft style includes the entry keys in the bibliography. The bibliography will be sorted by entry key. The style will set the following package options at load time: \kvopt{sorting}{debug}.

\item[debug]\seestyleexample{82-style-debug-biber}
This style prints all bibliographic data in tabular format. It is intended for debugging only and will set the following package options at load time: \kvopt{sorting}{debug}.

\end{marglist}

\subsection{Extended Name Format}
\label{use:enf}
The parsing rules for \bibtex names are rather archaic and not suited to
many international name formats. \biber supports an extended name format
which allows explicit specification of the parts of names. This allows the
use of custom name parts apart from the four standard \bibtex parts. Extended
name formats are supported in all name fields and can be used along with
the usual \bibtex name format. Recognition of extended name format can be
disabled with the \biber option \opt{--noxname} in case you do not need the
extended format and the auto-detection causes problems with normal name
parsing. The separator \verb+=+ which comes between the namepart names and
values is customisable with the \biber option \opt{--xnamesep}. Here is an
example:

\begin{lstlisting}[style=bibtex]{}
AUTHOR = {Hans Harman and Simon de Beumont}
AUTHOR = {given=Hans, family=Harman and given=Simon, prefix=de, family=Beumont}
\end{lstlisting}
% 
These two name specifications are equivalent but the extended format
explicitly names the parts. The supported parts are those specified by the
\biblatex data mode constant \opt{nameparts}, see \secref{aut:bbx:drv}. As
with traditional \bibtex name parsing, initials are automatically generated
but it is also possible to specify these explicitly:

\begin{lstlisting}[style=bibtex]{}
AUTHOR = {given=Jean, prefix=de la, prefix-i=d, family=Rousse}
AUTHOR = {given={Jean Pierre Simon}, given-i=JPS}
\end{lstlisting}
% 
Initials are specified by adding the suffix \verb+-i+ to the namepart name.
Compound parts may be protected with braces:

\begin{lstlisting}[style=bibtex]{}
AUTHOR = {given={Jean Pierre}}
\end{lstlisting}
% 
If a namepart contains a comma, the whole namepart should be protected with
quotes:

\begin{lstlisting}[style=bibtex]{}
AUTHOR = {"family={Robert and Sons, Inc.}"}
\end{lstlisting}
% 
Traditional \bibtex name formats and the extended form may be used together:

\begin{lstlisting}[style=bibtex]{}
AUTHOR = {Hans Harman and given=Simon, prefix=de, family=Beumont}
\end{lstlisting}
% 
Per-namelist and per-name options may be specified in the extended name
format, see \secref{use:opt:bib:hyb}:

\begin{lstlisting}[style=bibtex]{}
AUTHOR = {nosortothers=true and Hans Harman and
          given=Simon, family=Beumont, prefix=de, useprefix=true}
\end{lstlisting}
      
\subsection{Related Entries}
\label{use:rel}

Almost all bibliography styles require authors to specify certain types of relationship between entries such as Reprint of, Reprinted in etc. It is impossible to provide data fields to cover all of these relationships and so \biblatex provides a general mechanism for this using the entry fields \bibfield{related}, \bibfield{relatedtype} and \bibfield{relatedstring}. A related entry does not need to be cited and does not appear in the bibliography itself (unless of course it is also cited itself independently) as a clone is taken of the related entry to be used as a data source. The \bibfield{relatedtype} field should specify a localisation string which will be printed before the information from the related entries is printed, for example Orig. Pub. as. The \bibfield{relatedstring} field can be used to override the string determined via \bibfield{relatedtype}. Some examples:

\begin{lstlisting}[style=bibtex]{}
@Book{key1,
  ...
  related     = {key2},
  relatedtype = {reprintof},
  ...
}

@Book{key2,
  ...
}
\end{lstlisting}
%
Here we specify that entry \texttt{key1} is a reprint of entry \texttt{key2}. In the bibliography driver for \texttt{Book} entries, when \cmd{usebibmacro\{related\}} is called for entry \texttt{key1}:

\begin{itemize}
\item If the localisation string \texttt{reprintof} is defined, it is printed in the \texttt{relatedstring:reprintof} format. If this formatting directive is undefined, the string is printed in the \texttt{relatedstring:default} format.
\item If the \texttt{related:reprintof} macro is defined, it is used to format the information contained in entry \texttt{key2}, otherwise the \texttt{related:default} macro is used
\item If the \texttt{related:reprintof} format is defined, it is used to format both the localisation string and data. If this format is not defined, then the \texttt{related} format is used instead.
\end{itemize}
%
It is also supported to have cascading and/or circular relations:

\begin{lstlisting}[style=bibtex]{}
@Book{key1,
  ...
  related     = {key2},
  relatedtype = {reprintof},
  ...
}

@Book{key2,
  ...
  related     = {key3},
  relatedtype = {translationof},
  ...
}

@Book{key3,
  ...
  related     = {key2},
  relatedtype = {translatedas},
  ...
}
\end{lstlisting}
%
Multiple relations to the same entry are also possible:
\begin{lstlisting}[style=bibtex]{}
@MVBook{key1,
  ...
  related     = {key2,key3},
  relatedtype = {multivolume},
  ...
}

@Book{key2,
  ...
}

@Book{key3,
  ...
}
\end{lstlisting}
%
Note the order of the keys in lists of multiple related entries is important. The data from multiple related entries is printed in the order of the keys listed in this field. See \secref{aut:ctm:rel} for a more details on the mechanisms behind this feature. You can turn this feature off using the package option \opt{related} from \secref{use:opt:pre:gen}.

You can use the \bibfield{relatedoptions} to set options on the related entry data clone. This is useful if you need to override the \opt{dataonly} option which is set by default on all related entry clones. For example, if you will expose some of the names in the related clone in your document, you may want to have them disambiguated from names in other entries but normally this won't happen as related clones have the per"=entry \opt{dataonly} option set and this in turn sets \kvopt{uniquename}{false} and \kvopt{uniquelist}{false}. In such a case, you can set \bibfield{relatedoptions} to just \opt{skiplab, skipbib, skipbiblist}.

\subsection{Sorting Options}
\label{use:srt}

This package supports fully customisable sorting templates for the bibliography. The default global sorting template is selected with the \opt{sorting} package option from \secref{use:opt:pre:gen}. Apart from the regular data fields there are also some special fields which may be used to optimize the sorting of the bibliography. \Apxref{apx:srt:a1, apx:srt:a2} give an outline of the default alphabetic sorting templates supported by \biblatex. Chronological sorting templates are listed in \apxref{apx:srt:chr}. A few explanations concerning the default templates are in order.

The first item considered in the sorting process is always the \bibfield{presort} field of the entry. If this field is undefined, \biblatex will use the default value <\texttt{mm}> as a presort string. The next item considered is the \bibfield{sortkey} field. If this field is defined, it serves as the master sort key. Apart from the \bibfield{presort} field, no further data is considered in this case. If the \bibfield{sortkey} field is undefined, sorting continues with the name. The package will try using the \bibfield{sortname}, \bibfield{author}, \bibfield{editor}, and \bibfield{translator} fields, in this order. Which fields are considered also depends on the setting of the \opt{use$<$name$>$} options. If all such options are disabled, the \bibfield{sortname} field is ignored as well. Note that all name fields are responsive to \opt{maxnames} and \opt{minnames}. If no name field is available, either because all of them are undefined or because all \opt{use$<$name$>$} options are disabled, \biblatex will fall back to the \bibfield{sorttitle} and \bibfield{title} fields as a last resort. The remaining items are, in various order: the \bibfield{sortyear} field, if defined, or the first four digits of the \bibfield{year} field otherwise; the \bibfield{sorttitle} field, if defined, or the \bibfield{title} field otherwise; the \bibfield{volume} field. Note that the sorting templates shown in \apxref{apx:srt:a2} include an additional item: \bibfield{labelalpha} is the label used by <alphabetic> bibliography styles. Strictly speaking, the string used for sorting is \bibfield{labelalpha}~+ \bibfield{extraalpha}. The sorting templates in \apxref{apx:srt:a2} are intended to be used in conjunction with alphabetic styles only.

The chronological sorting templates presented in \apxref{apx:srt:chr} also make use of the \bibfield{presort} and \bibfield{sortkey} fields, if defined. The next item considered is the \bibfield{sortyear} or the \bibfield{year} field, depending on availability. The \opt{ynt} template extracts the first four Arabic figures from the field. If both fields are undefined, the string \texttt{9999} is used as a fallback value. This means that all entries without a year will be moved to the end of the list. The \opt{ydnt} template is similar in concept but sorts the year in descending order. As with the \opt{ynt} template, the string \texttt{9999} is used as a fallback value. The remaining items are similar to the alphabetic sorting templates discussed above. Note that the \opt{ydnt} sorting template will only sort the date in descending order. All other items are sorted in ascending order as usual.

Using special fields such as \bibfield{sortkey}, \bibfield{sortname}, or \bibfield{sorttitle} is usually not required. The \biblatex package is quite capable of working out the desired sorting order by using the data found in the regular fields of an entry. You will only need them if you want to manually modify the sorting order of the bibliography or if any data required for sorting is missing. Please refer to the field descriptions in \secref{bib:fld:spc} for details on possible uses of the special fields.

\subsection{Data Annotations}
\label{use:annote}
Ideally, there should be no formatting information in a bibliography data file, however, sometimes such questionable practice seems to the only way in which the desired results can be achieved. Data annotations are a way of addressing this by allowing users to attach semantic information (rather than typographical markup) to information in a bibliography data source so that the information can be used at markup time by a style. For example, if you wanted to highlight certain names in a work depending on whether they were a student author (indicated by a superscript asterisk in the references) or a corresponding author (indicated by bold face), then you might be tempted to try:

\begin{lstlisting}[style=bibtex]{}
@MISC{Article1,
  AUTHOR = {Last1\textsuperscript{*}, First1 and \textbf{Last2}, \textbf{First2} and Last3, First3}
}
\end{lstlisting}
%
There are several problems with this. Firstly, it will break \bibtex's fragile name parsing routines and probably won't compile at all. Secondly, it is not only mixing up data with markup, it does so in a hard-coded way: this data can't easily be shared and used with other styles. While it is possible to achieve this formatting using \biblatex internals in a style or document, this is a complex and unreliable method which many users will not wish to use.

In order to address these issues, \biblatex has a general data annotation facility which allows you to attach any number of a comma"=separated list of annotations to data fields, items within data field lists (like names) and even parts of specific items such as parts of names (given name, family name etc.). There are macros provided to check for annotations which can be used in formatting directives.

There are three scopes for data annotations, in order of increasing specificity:

\begin{itemize}
\item \opt{field}---applied to top-level fields in a data source entry
\item \opt{item}---applied to items within a list field in a data source entry
\item \opt{part}---applied to parts within items within a list field in a data source entry
\end{itemize}
%
Data annotations are supported for \bibtex and \biblatexml data sources.

\begin{lstlisting}[style=bibtex]{}
@MISC{ann1,
  AUTHOR           = {Last1, First1 and Last2, First2 and Last3, First3},
  AUTHOR+an        = {1:family=student;2=corresponding},
  TITLE            = {The Title},
  TITLE+an:default = {=titleannotation},
  TITLE+an:french  = {="Le titre"},
  TITLE+an:german  = {="Der Titel"},
}
\end{lstlisting}
%
Here the field name suffix \texttt{+an} is a user-definable\footnote{See \biber's \opt{--annotation-marker} option.} suffix which marks a data field as an annotation of the unsuffixed field. Multiple annotations can be provided for the same field since all annotations are named. After the annotation marker is the optional named annotation marker \footnote{See \biber's \opt{--named-annotation-marker} option.} and an optional annotation name. The annotation name is <default> if not specified and so in the above example the following two are equivalent:

\begin{lstlisting}[style=bibtex]{}
TITLE+an         = {=titleannotation},
TITLE+an:default = {=titleannotation},
\end{lstlisting}
%
The format of annotation fields in \bibtex data sources is is as follows:

\begin{lstlisting}
<annotationspecs> ::= <annotationspec> [ ";" <annotationspec> ]
<annotationspec>  ::= [ <itemcount> [ ":" <part> ] ] "=" <annotations>
<annotations>     ::= <annotation> [ "," <annotation> ]
<annotation>      ::= ["] (string) ["]
\end{lstlisting}
%
That is, one or more specifications separated by semicolons. Each specification is an equals sign followed by a comma"=separated list of annotation keywords or a string enclosed in double-quotes (a <literal> annotation, see below). To annotate a specific item in a list, put the number of the list item before the equals sign (lists start at 1). If you need to annotate a specific part of the list item, give its name after the list item number, preceded by a colon. Name part names are defined in the data model, see \secref{aut:bbx:drv}. Some further examples:

\begin{lstlisting}[style=bibtex]{}
AUTHOR      = {Last1, First1 and Last2, First2 and Last3, First3},
AUTHOR+an   = {3:given=annotation1, annotation2},
TITLE       = {A title},
TITLE+an    = {=a title annotation, another title annotation},
LANGUAGE    = {english and french},
LANGUAGE+an = {1=annotation3; 2=annotation4}
}
\end{lstlisting}
%
Attaching annotations to data is similar in \biblatexml data sources. Using the example above, we would have:

\begin{lstlisting}[language=xml]
<bltx:entries xmlns:bltx="http://biblatex-biber.sourceforge.net/biblatexml">
  <bltx:entry id="test" entrytype="misc">
    <bltx:names type="author">
      <bltx:name>
        <bltx:namepart type="given" initial="F">First1</bltx:namepart>
        <bltx:namepart type="family" initial="L">Last1</bltx:namepart>
      </bltx:name>
      <bltx:name>
        <bltx:namepart type="given" initial="F">First2</bltx:namepart>
        <bltx:namepart type="family" initial="L">Last2</bltx:namepart>
      </bltx:name>
      <bltx:name>
        <bltx:namepart type="given" initial="F">First3</bltx:namepart>
        <bltx:namepart type="family" initial="L">Last3</bltx:namepart>
      </bltx:name>
    </bltx:names>
    </bltx:annotation field="author" item="1" part="family">student</bltx:annotation>
    </bltx:annotation field="author" item="2">corresponding</bltx:annotation>
  </bltx:entry>
</bltx:entries>
\end{lstlisting}
%
To access the annotation information when formatting bibliography data, macros are provided, corresponding to the three annotation scopes:

\begin{ltxsyntax}

\cmditem{iffieldannotation}[field][annotationname]{annotation}{true}{false}

Executes \prm{true} if the data field \prm{field} has an annotation \prm{annotation} for the annotation called \prm{annotationname} and false otherwise. If \prm{annotationname} is not given, then the annotation named <default> is assumed (this is the name given to annotations defined without an explicit name). If \prm{field} is not given, the current data field as indicated by \cmd{currentfield}, \cmd{currentlist} or \cmd{currentname} (see \secref{aut:bib:fmt}) is assumed. Of course, this is only possible if these commands are defined, that is, inside formatting directives.

\cmditem{ifitemannotation}[field][annotationname][item]{annotation}{true}{false}

Executes \prm{true} if the item \prm{item} in the data field \prm{field} has an annotation \prm{annotation} and false otherwise. If \prm{annotationname} is not given, then the annotation named <default> is assumed (this is the name given to annotations defined without an explicit name). The optional argument \prm{field} can be inferred if not provided as with \cmd{iffieldannotation}. If \prm{item} is not given, the number of the item currently being processed as given by \cnt{listcount} is used.

\cmditem{ifpartannotation}[field][annotationname][item]{part}{annotation}{true}{false}

Executes \prm{true} if the part named \prm{part} in item \prm{item} in the data field \prm{field} has an annotation \prm{annotation} and false otherwise. If \prm{annotationname} is not given, then the annotation named <default> is assumed (this is the name given to annotations defined without an explicit name). The two optional arguments \prm{field} and \prm{item} can be inferred as in \cmd{ifitemannotation}. The parameter \prm{part} can never be inferred and is therefore a mandatory argument.

Date fields are special and handled in a context where \cmd{currentfield} is not accessible. Thus there is a fourth command to test annotations for dates.

\cmditem{ifdateannotation}[annotationname]{datetype}{annotation}{true}{false}

Executes \prm{true} if the date field \prm{datetype} has an annotation \prm{annotation} and false otherwise. If \prm{annotationname} is not given, then the annotation named <default> is assumed (this is the name given to annotations defined without an explicit name). The \prm{datetype} argument is mandatory, because it cannot be inferred in most contexts where \cmd{ifdateannotation} will be used.

\cmditem{hasfieldannotation}[field][annotationname]{true}{false}

Executes \prm{true} if the data field \prm{field} has a literal annotation \prm{annotationname} defined and false otherwise. If \prm{annotationname} is not given, then the annotation named <default> is assumed (this is the name given to annotations defined without an explicit name). If \prm{field} is not given, the current data field as indicated by \cmd{currentfield}, \cmd{currentlist} or \cmd{currentname} (see \secref{aut:bib:fmt}) is assumed. Of course, this is only possible if these commands are defined, that is, inside formatting directives.

\cmditem{hasitemannotation}[field][annotationname][item]{true}{false}

Executes \prm{true} if the item \prm{item} in the data field \prm{field} has a literal annotation \prm{annotationname} defined and false otherwise. If \prm{annotationname} is not given, then the annotation named <default> is assumed (this is the name given to annotations defined without an explicit name). The optional argument \prm{field} can be inferred if not provided as with \cmd{iffieldannotation}. If \prm{item} is not given, the number of the item currently being processed as given by \cnt{listcount} is used.

\cmditem{haspartannotation}[field][annotationname][item]{part}{true}{false}

Executes \prm{true} if the part named \prm{part} in the item \prm{item} in the data field \prm{field} has a literal annotation \prm{annotationname} defined and false otherwise. If \prm{annotationname} is not given, then the annotation named <default> is assumed (this is the name given to annotations defined without an explicit name). The two optional arguments \prm{field} and \prm{item} can be inferred as in \cmd{ifitemannotation}. The parameter \prm{part} can never be inferred and is therefore a mandatory argument.

Date fields are special and handled in a context where \cmd{currentfield} is not accessible. Thus there is a fourth command to test the existence of annotations for dates.

\cmditem{hasdateannotation}[annotationname]{datetype}{true}{false}

Executes \prm{true} if the date field \prm{datetype} has any annotation \prm{annotationname} defined and false otherwise. If \prm{annotationname} is not given, then the annotation named <default> is assumed (this is the name given to annotations defined without an explicit name). The \prm{datetype} argument is mandatory, because it cannot be inferred in most contexts where \cmd{ifdateannotation} will be used.

\end{ltxsyntax}
%
As an example of how to use the annotation information to solve the problem originally presented in this section, this could be used in the name formatting directives to put an asterisk after all family names annotated as student:

\begin{lstlisting}[style=latex]{}
  \ifpartannotation{family}{student}
    {\textsuperscript{*}}
    {}%
\end{lstlisting}
%
To put the given and family names of name list items annotated as corresponding in boldface:

\begin{lstlisting}[style=latex]{}
\renewcommand*{\mkbibnamegiven}[1]{%
  \ifitemannotation{corresponding}
    {\textbf{#1}}
    {#1}}

\renewcommand*{\mkbibnamefamily}[1]{%
  \ifitemannotation{corresponding}
    {\textbf{#1}}
    {#1}}
\end{lstlisting}

\subsubsection{Literal Annotations}

If the annotation is a string enclosed in double-quotes, the annotation is a <literal> annotation. In this case the annotation can be retrieved and used as a string rather than as meta-information used to determine formatting. This is useful in order to be able to attached specific annotations to data which are to be printed as-is. For example:

\begin{lstlisting}[style=bibtex]{}
AUTHOR = {{American Educational Research Association} and {American Psychological Association}
            and {National Council on Measurement in Education}},
AUTHOR+an = {1:family="AERA"; 2:family="APA"; 3:family="NCME"}
}
\end{lstlisting}
%
Such annotations are not keys whose presence can be tested for but are rather literal information attached to the data. The values are retrieved by the following macros

\begin{ltxsyntax}

\cmditem{getfieldannotation}[field][annotationname]

Retrieves any literal annotation for the field \prm{field}. If \prm{annotationname} is not given, then the annotation named <default> is assumed (this is the name given to annotations defined without an explicit name). If \prm{field} is not given, the current data field as indicated by \cmd{currentfield}, \cmd{currentlist} or \cmd{currentname} (see \secref{aut:bib:fmt}) is assumed. Of course, this is only possible if these commands are defined, that is, inside formatting directives.

\cmditem{getitemannotation}[field][annotationname][item]

Retrieves any literal annotation for the item \prm{item} in the field \prm{field}. If \prm{annotationname} is not given, then the annotation named <default> is assumed (this is the name given to annotations defined without an explicit name). The optional argument \prm{field} can be inferred if not provided as with \cmd{getfieldannotation}. If \prm{item} is not given, the number of the item currently being processed as given by \cnt{listcount} is used.

\cmditem{getpartannotation}[field][annotationname][item]{part}

Retrieves any literal annotation for the part \prm{part}. If \prm{annotationname} is not given, then the annotation named <default> is assumed (this is the name given to annotations defined without an explicit name). The two optional arguments \prm{field} and \prm{item} can be inferred as in \cmd{getitemannotation}. The parameter \prm{part} can never be inferred and is therefore a mandatory argument.

Date fields are special and handled in a context where \cmd{currentfield} is not accessible. Thus there is a fourth command to access literal annotations for dates.

\cmditem{getdateannotation}[annotationname]{datetype}

Retrieve a literal annotation for the datefield \prm{datetype}. If \prm{annotationname} is not given, then the annotation named <default> is assumed (this is the name given to annotations defined without an explicit name). The \prm{datetype} argument is mandatory, because it cannot be inferred in most contexts where \cmd{getdateannotation} will be used.

\end{ltxsyntax}
%
So, for example, given the bibliography entry above, we could put the following in the preamble:

\begin{ltxexample}[style=latex]
\renewcommand*{\mkbibnamefamily}[1]{%
  #1\space\mkbibparens{\getpartannotation{family}}}
\end{ltxexample}
%
In order to get something like this in the bibliography when formatting names:

\begin{lstlisting}[style=bibtex]{}
  American Educational Research Association (AERA) and
  American Psychological Association (APA), and
  National Council on Measurement in Education (NCME)
}
\end{lstlisting}
%
Naturally there are semantically more elegant ways of dealing with corporate authors without using the <family> namepart (see \secref{aut:bbx:drv}) but this example demonstrates clearly a use for literal annotations.

\subsection{Bibliography Commands}
\label{use:bib}

\subsubsection{Resources}
\label{use:bib:res}

\begin{ltxsyntax}

\cmditem{addbibresource}[options]{resource}

Adds a \prm{resource}, such as a \file{.bib} file, to the default resource list. This command is only available in the preamble. It replaces the \cmd{bibliography} legacy command. Note that files must be specified with their full name, including the extension. With \biber, the resource name can be a BSD-style glob pattern. This only makes sense when resources refer to files with an absolute or relative path and does not work when looking for data resources in \biber s input/output directories or with resources located by \prm{kpsewhich} etc. When running on Windows, \biber will switch to a Windows compatible globbing mode where backslashes are also useable as path separators and case does not matter. If the resources contain duplicate entries (that is, duplicate \bibfield{entrykey}s), it is backend dependent what then happens. For example, by default \biber will ignore further occurrence of \bibfield{entrykey}s unless its \opt{--noskipduplicates} options is used. Invoke \cmd{addbibresource} multiple times to add more resources, for example:

\begin{ltxexample}
\addbibresource{bibfile1.bib}
\addbibresource{bibfile2.bib}
\addbibresource[glob]{bibfiles/bibfile*.bib}
\addbibresource[glob]{bibfile-num?.bib}
\addbibresource[glob]{bibfile{1,2,3}.bib}
\addbibresource[location=remote]{https://raw.githubusercontent.com/plk/biblatex/master/bibtex/bib/biblatex/biblatex-examples.bib}
\addbibresource[location=remote,label=lan]{ftp://192.168.1.57/~user/file.bib}
\end{ltxexample}
%
Since the \prm{resource} string is read in a verbatim-like mode, it may contain arbitrary characters. The only restriction is that any curly braces must be balanced. The following \prm{options} are available:

\begin{optionlist*}

\valitem{bibencoding}{bibencoding}

This option can be used to override the global \opt{bibencoding} option for a particular \prm{resource}.

\valitem{label}{identifier}

Assigns a label to a resource. The \prm{identifier} may be used in place of the full resource name in the optional argument of \env{refsection} (see \secref{use:bib:sec}). The label is a \emph{unique} identifier for the \prm{resource}, so each label should only be used once.

\valitem[local]{location}{location}

The location of the resource. The \prm{location} may be either \texttt{local} for local resources or \texttt{remote} for \acr{URL}s. Remote resources require \biber. The protocols \acr{HTTP/HTTPS} and \acr{FTP} are supported. The remote \acr{URL} must be a fully qualified path to a \file{bib} file or a \acr{URL} which returns a \file{bib} file.

\valitem[file]{type}{type}

The type of resource. Currently, the only supported type is \texttt{file}.

\valitem[bibtex]{datatype}{datatype}

The data type (format) of the resource. The following formats are currently supported:

\begin{valuelist}

\item[bibtex] \bibtex format.

\item[biblatexml]  Experimental XML format for \biblatex. See \secref{apx:biblatexml}.

\end{valuelist}

\boolitem{glob}

Whether \biber should glob (expand according to pattern) the datasource name. There is
a global setting for this in \biber (false by default and settable to true
using the \opt{--glob-datasources} option). This option allows overriding
the \biber default on a per-resource basis.

\end{optionlist*}

\cmditem{addglobalbib}[options]{resource}

This command differs from \cmd{addbibresource} in that the \prm{resource} is added to the global resource list. The difference between default resources and global resources is only relevant if there are reference sections in the document and the optional argument of \env{refsection} (\secref{use:bib:sec}) is used to specify alternative resources which replace the default resource list. Any global resources are added to all reference sections.

\cmditem{addsectionbib}[options]{resource}

This command differs from \cmd{addbibresource} in that the resource \prm{options} are registered but the \prm{resource} not added to any resource list. This is only required for resources which 1) are given exclusively in the optional argument of \env{refsection} (\secref{use:bib:sec}) and 2) require options different from the default settings. In this case, \cmd{addsectionbib} is employed to qualify the \prm{resource} prior to using it by setting the appropriate \prm{options} in the preamble. The \opt{label} option may be useful to assign a short name to the resource.

\cmditem{bibliography}{bibfile, \dots}\DeprecatedMark

The legacy command for adding bibliographic resources, supported for backwards compatibility. Like \cmd{addbibresource}, this command is only available in the preamble and adds resources to the default resource list. Its argument is a comma"=separated list of \file{bib} files. The \file{.bib} extension may be omitted from the filename. Invoking this command multiple times to add more files is permissible. This command is deprecated. Please consider using \cmd{addbibresource} instead.

\subsubsection{The Bibliography}
\label{use:bib:bib}

\cmditem{printbibliography}[key=value, \dots]

This command prints the bibliography. It takes one optional argument, which is a list of options given in \keyval notation. The following options are available:

\end{ltxsyntax}

\begin{optionlist*}

\valitem[bibliography/shorthands]{env}{name}

The <high-level> layout of the bibliography and the list of shorthands is controlled by environments defined with \cmd{defbibenvironment}. This option selects an environment. The \prm{name} corresponds to the identifier used when defining the environment with \cmd{defbibenvironment}. By default, the \cmd{printbibliography} command uses the identifier \texttt{bibliography}; \cmd{printbiblist} uses \texttt{shorthands}. See also \secref{use:bib:biblist,use:bib:hdg}.

\valitem[bibliography/shorthands]{heading}{name}

The bibliography and the list of shorthands typically have a chapter or section heading. This option selects the heading \prm{name}, as defined with \cmd{defbibheading}. By default, the \cmd{printbibliography} command uses the heading \texttt{bibliography}; \cmd{printbiblist} uses \texttt{shorthands}. See also \secref{use:bib:biblist,use:bib:hdg}.

\valitem{title}{text}

This option overrides the default title provided by the heading selected with the \opt{heading} option, if supported by the heading definition. See \secref{use:bib:hdg} for details.

\valitem{label}{label}

If \prm{label} is nonempty, issue \texttt{\textbackslash label\{\prm{label}\}} after typesetting the heading. No sanity checking is done whether or not it is useful to set a label after the heading (\eg if the heading is not numbered a \texttt{\textbackslash ref} to the label might not result in useful output).

\optitem[\normalfont\emph{global setting} (\opt{none})]{block}{\opt{none}, \opt{space}, \opt{par}, \opt{nbpar}, \opt{ragged}}

This option overrides the global \opt{block} option (see \secref{use:opt:pre:gen}, the meaning of the settings is explained there as well).

\valitem{prenote}{name}

The prenote is an arbitrary piece of text to be printed after the heading but before the list of references. This option selects the prenote \prm{name}, as defined with \cmd{defbibnote}. By default, no prenote is printed. The note is printed in the standard text font. It is not affected by \cmd{bibsetup} and \cmd{bibfont} but it may contain its own font declarations. See \secref{use:bib:nts} for details.

\valitem{postnote}{name}

The postnote is an arbitrary piece of text to be printed after the list of references. This option selects the postnote \prm{name}, as defined with \cmd{defbibnote}. By default, no postnote is printed. The note is printed in the standard text font. It is not affected by \cmd{bibsetup} and \cmd{bibfont} but it may contain its own font declarations. See \secref{use:bib:nts} for details.

\intitem[\normalfont\em current section]{section}

Print only entries cited in reference section \prm{integer}. The reference sections are numbered starting at~1. All citations given outside a \env{refsection} environment are assigned to section~0. See \secref{use:bib:sec} for details and \secref{use:use:mlt} for usage examples.

\intitem{segment}

Print only entries cited in reference segment \prm{integer}. The reference segments are numbered starting at~1. All citations given outside a \env{refsegment} environment are assigned to segment~0. See \secref{use:bib:seg} for details and \secref{use:use:mlt} for usage examples. Remember that segments within a section are numbered local to the section so the segment you request will be the nth segment in the requested (or currently active enclosing) section.

\valitem{type}{entrytype}

Print only entries whose entry type is \prm{entrytype}.

\valitem{nottype}{entrytype}

Print only entries whose entry type is not \prm{entrytype}. This option may be used multiple times.

\valitem{subtype}{subtype}

Print only entries whose \bibfield{entrysubtype} is defined and \prm{subtype}.

\valitem{notsubtype}{subtype}

Print only entries whose \bibfield{entrysubtype} is undefined or not \prm{subtype}. This option may be used multiple times.

\valitem{keyword}{keyword}

Print only entries whose \bibfield{keywords} field includes \prm{keyword}. This option may be used multiple times.

\valitem{notkeyword}{keyword}

Print only entries whose \bibfield{keywords} field does not include \prm{keyword}. This option may be used multiple times.

\valitem{category}{category}

Print only entries assigned to category \prm{category}. This option may be used multiple times.

\valitem{notcategory}{category}

Print only entries not assigned to category \prm{category}. This option may be used multiple times.

\valitem{filter}{name}

Filter the entries with filter \prm{name}, as defined with \cmd{defbibfilter}. See \secref{use:bib:flt} for details.

\valitem{check}{name}

Filter the entries with check \prm{name}, as defined with \cmd{defbibcheck}. See \secref{use:bib:flt} for details.

\valitem{resetnumbers}{true,false,number}

This option applies to numerical citation\slash bibliography styles only and requires that the \opt{defernumbers} option from \secref{use:opt:pre:gen} be enabled globally. If enabled, it will reset the numerical labels assigned to the entries in the respective bibliography, \ie the numbering will restart at~1. You can also pass a number to this option, for example: \texttt{resetnumbers=10} to reset numbering to the specified number to aid numbering continuity across documents. Use this option with care as \biblatex can not guarantee unique labels globally if they are reset manually.

\boolitem{omitnumbers}

This option applies to numerical citation\slash bibliography styles only and requires that the \opt{defernumbers} option from \secref{use:opt:pre:gen} be enabled globally. If enabled, \biblatex will not assign a numerical label to the entries in the respective bibliography. This is useful when mixing a numerical subbibliography with one or more subbibliographies using a different scheme (\eg author-title or author-year).

\boolitem[false]{locallabelwidth}

Calculate \cmd{labelnumberwidth}, \cmd{labelalphawidth} and similar lengths locally for the present bibliography and not globally for all entries. See also \opt{labelnumberwidth} in \secref{use:opt:pre:gen}.

\end{optionlist*}

\begin{ltxsyntax}

\cmditem{bibbysection}[key=value, \dots]

This command automatically loops over all reference sections. This is equivalent to giving one \cmd{printbibliography} command for every section but has the additional benefit of automatically skipping sections without references. Note that \cmd{bibbysection} starts looking for references in section \texttt{1}. It will ignore references given outside of \env{refsection} environments since they are assigned to section~0. See \secref{use:use:mlt} for usage examples. The options are a subset of those supported by \cmd{printbibliography}. Valid options are \opt{env}, \opt{heading}, \opt{prenote}, \opt{postnote}. The current bibliography context sorting template is used for all sections (see \secref{use:bib:context}).

\cmditem{bibbysegment}[key=value, \dots]

This command automatically loops over all reference segments. This is equivalent to giving one \cmd{printbibliography} command for every segment in the current \env{refsection} but has the additional benefit of automatically skipping segments without references. Note that \cmd{bibbysegment} starts looking for references in segment \texttt{1}. It will ignore references given outside of \env{refsegment} environments since they are assigned to segment~0. See \secref{use:use:mlt} for usage examples. The options are a subset of those supported by \cmd{printbibliography}. Valid options are \opt{env}, \opt{heading}, \opt{prenote}, \opt{postnote}. The current bibliography context sorting template is used for all segments (see \secref{use:bib:context}).

\cmditem{bibbycategory}[key=value, \dots]

This command loops over all bibliography categories. This is equivalent to giving one \cmd{printbibliography} command for every category but has the additional benefit of automatically skipping empty categories. The categories are processed in the order in which they were declared. See \secref{use:use:mlt} for usage examples. The options are a subset of those supported by \cmd{printbibliography}. Valid options are \opt{env}, \opt{prenote}, \opt{postnote}, \opt{section}. Note that \opt{heading} is not available with this command. The name of the current category is automatically used as the heading name. This is equivalent to passing \texttt{heading=\prm{category}} to \cmd{printbibliography} and implies that there must be a matching heading definition for every category. The current bibliography context sorting template is used for all categories (see \secref{use:bib:context}).

\cmditem{printbibheading}[key=value, \dots]

This command prints a bibliography heading defined with \cmd{defbibheading}. It takes one optional argument, which is a list of options given in \keyval notation. The options are a small subset of those supported by \cmd{printbibliography}. Valid options are \opt{heading}, \opt{title} and \opt{label}. By default, this command uses the heading \texttt{bibliography}. See \secref{use:bib:hdg} for details. Also see \secref{use:use:mlt,use:use:div} for usage examples.

\cmditem{DeclarePrintbibliographyDefaults}{key=value, \dots}

This command can be used to globally set defaults for some options to \cmd{printbibliography}, the \cmd{bibby...} bibliography commands and \cmd{printbibheading}.
The supported keys are
\begin{itemize}
  \item \opt{env}
  \item \opt{heading}
  \item \opt{title}
  \item \opt{prenote}
  \item \opt{postnote}
  \item \opt{filter}
\end{itemize}

\end{ltxsyntax}
%
To print a bibliography with a different sorting template than the global sorting template, use the bibliography context switching commands from \secref{use:bib:context}.

\subsubsection{Bibliography Lists}
\label{use:bib:biblist}

\biblatex can, in addition to printing normal bibliographies, also print arbitrary lists of information derived from the bibliography data such as a list of shorthand abbreviations for particular entries or a list of abbreviations of journal titles.

A bibliography list differs from a normal bibliography in that the same bibliography driver is used to print all entries rather than a specific driver being used for each entry depending on the entry type.

\begin{ltxsyntax}

\cmditem{printbiblist}[key=value, \dots]{biblistname}

This command prints a bibliography list. It takes an optional argument, which is a list of options given in \keyval notation. Valid options are all options supported by \cmd{printbibliography} (\secref{use:bib:bib}) except \opt{resetnumbers} and \opt{omitnumbers}. Additionaly, the two options \opt{driver} and \opt{biblistfilter} are available. If there are any \env{refsection} environments in the document, the bibliography list will be local to these environments; see \secref{use:bib:sec} for details. By default, this command uses the heading \texttt{biblist}. See \secref{use:bib:hdg} for details.


The \prm{biblistname} is a mandatory argument which names the bibliography list. This name is used to identify:
\begin{itemize}
\item The default bibliography driver used to print the list entries
\item A default bibliography list filter declared with \cmd{DeclareBiblistFilter} (see \secref{aut:ctm:bibfilt}) used to filter the entries returned from \biber
\item A default check declared with \cmd{defbibcheck} (see \secref{use:bib:flt}) used to post-process the list entries
\item The default bib environment to use
\item The default sorting template to use
\end{itemize}

The two additional options can be used to change some of the defaults set by the mandatory argument.

\begin{optionlist*}
\valitem[\prm{biblistname}]{driver}{driver}

Change the bibliography driver used to print the list entries.
% \prm{driver} must be a valid driver declared with \cmd{DeclareBibliographyDriver} (see \secref{aut:bbx:bbx}).

\valitem[\prm{biblistname}]{biblistfilter}{biblistfilter}

Change the bibliography list filter used to filter the entries. \prm{biblistfilter} must be a valid bibliography list filter defined with \cmd{DeclareBiblistFilter} (see \secref{aut:ctm:bibfilt}).
\end{optionlist*}

In terms of sorting the list, the default is to sort using the sorting template named after the bibliography list (if it exists) and only then to fall back to the current context sorting template if this is not defined (see \secref{use:bib:context}).

The most common bibliography list is a list of shorthand abbreviations for certain entries and so this has a convenience alias \cmd{printshorthands[\dots]} for backwards compatibility which is defined as:

\begin{ltxexample}
\printbiblist[...]{shorthand}
\end{ltxexample}

\biblatex provides automatic support for data source fields in the default data model marked as <Label fields> (See \secref{bib:fld:dat}). Such fields automatically have defined for them:

\begin{itemize}
\item A default bib environment (See \secref{use:bib:hdg})
\item A bibliography list filter (See \secref{aut:ctm:bibfilt})
\item Some supporting formats and lengths (See \secref{aut:fmt:ilc} and \secref{aut:fmt:ich})
\end{itemize}
%
Therefore only a minimal setup is required to print bibliography lists with such fields. For example, to print a list of journal title abbreviations, you can minimally put this in your preamble:

\begin{ltxexample}
\DeclareBibliographyDriver{shortjournal}{%
  \printfield{journaltitle}}
\end{ltxexample}
%
Then you can put this in your document where you want to print the list:

\begin{ltxexample}
\printbiblist[title={Journal Shorthands}]{shortjournal}
\end{ltxexample}
%
Since \bibfield{shortjournal} is defined in the default data model as a <Label field>, this example:
\begin{itemize}
\item Uses the automatically created <shortjournal> bib environment
\item Uses the automatically created <shortjournal> bibliography list filter to return only entries with a \bibfield{shortjournal} field in the \file{.bbl}
\item Uses the defined <shortjournal> bibliography driver to print the entries
\item Uses the default <biblist> heading but overrides the title with <Journal Shorthands>
\item Uses the current bibliography context sorting template if no template exists with the name \bibfield{shortjournal}
\end{itemize}
%
Often, you will want to sort on the label field of the list and since a sorting template is automatically picked up if it is named after the list, in this case you could simply do:

\begin{ltxexample}
\DeclareSortingTemplate{shortjournal}{
  \sort{
        \field{shortjournal}
  }
}
\end{ltxexample}

Naturally all defaults can be overridden by options to \cmd{printbiblist} and definitions of the environments, filters etc. and in this way arbitrary types of bibliography lists can be printed containing a variety of information from the bibliography data.
\end{ltxsyntax}

Bibliography lists are often used to print lists of various kinds of shorthands and this can result in duplicate entries if more than one bibliography entry has the same shorthand. For example, several journal articles in the same journal would result in duplicate entries in a list of journal shorthands. You can use the fact that such lists automatically pick up a \cmd{bibcheck} with the same name as the list to define a check to remove duplicates. If you are defining a list to print all of the journal shorthands using the \bibfield{shortjournal} field, you could define a \cmd{bibcheck} like this:

\begin{ltxexample}[basicstyle=\displayverbfont\footnotesize]
\defbibcheck{shortjournal}{%
   \iffieldundef{shortjournal}
     {\skipentry}
     {\iffieldundef{journaltitle}
       {\skipentry}
       {\ifcsdef{sjcheck@\therefsection
          -\strfield{shortjournal}=\strfield{journaltitle}}
         {\skipentry}
         {\savefieldcs{journaltitle}{sjcheck@\therefsection
            -\strfield{shortjournal}=\strfield{journaltitle}}}}}}
\end{ltxexample}

\subsubsection{Bibliography Sections}
\label{use:bib:sec}

The \env{refsection} environment is used in the document body to mark a reference section. This environment is useful if you want separate, independent bibliographies and bibliography lists in each chapter, section, or any other part of a document. Within a reference section, all cited works are assigned labels which are local to the environment. Technically, reference sections are completely independent from document divisions such as \cmd{chapter} and \cmd{section} even though they will most likely be used per chapter or section. See the \opt{refsection} package option in \secref{use:opt:pre:gen} for a way to automate this. Also see \secref{use:use:mlt} for usage examples.

\begin{ltxsyntax}

\envitem{refsection}[resource, \dots]

The optional argument is a comma"=separated list of resources specific to the reference section. If the argument is omitted, the reference section will use the default resource list, as specified with \cmd{addbibresource} in the preamble. If the argument is provided, it replaces the default resource list. Global resources specified with \cmd{addglobalbib} are always considered. \env{refsection} environments may not be nested, but you may use \env{refsegment} environments within a \env{refsection} to subdivide it into segments. Use the \opt{section} option of \cmd{printbibliography} to select a section when printing the bibliography, and the corresponding option of \cmd{printbiblist} when printing bibliography lists. Bibliography sections are numbered starting at~\texttt{1}. The number of the current section is also written to the transcript file. All citations given outside a \env{refsection} environment are assigned to section~0. If \cmd{printbibliography} is used within a \env{refsection}, it will automatically select the current section. The \opt{section} option is not required in this case. This also applies to \cmd{printbiblist}. Beginning a new reference section automatically ends the active reference context (see \secref{use:bib:context}).

\cmditem{newrefsection}[resource, \dots]

This command is similar to the \env{refsection} environment except that it is a stand"=alone command rather than an environment. It automatically ends the previous reference section (if any) and immediately starts a new one. Note that the reference section started by the last \cmd{newrefsection} command in the document will extend to the very end of the document. Use \cmd{endrefsection} if you want to terminate it earlier.

\end{ltxsyntax}

\subsubsection{Bibliography Segments}
\label{use:bib:seg}

The \env{refsegment} environment is used in the document body to mark a reference segment. This environment is useful if you want one global bibliography which is subdivided by chapter, section, or any other part of the document. Technically, reference segments are completely independent from document divisions such as \cmd{chapter} and \cmd{section} even though they will typically be used per chapter or section. See the \opt{refsegment} package option in \secref{use:opt:pre:gen} for a way to automate this. Also see \secref{use:use:mlt} for usage examples.

\begin{ltxsyntax}

\envitem{refsegment}

The difference between a \env{refsection} and a \env{refsegment} environment is that the former creates labels which are local to the environment whereas the latter provides a target for the \opt{segment} filter of \cmd{printbibliography} without affecting the labels. They will be unique across the entire document. \env{refsegment} environments may not be nested, but you may use them in conjunction with \env{refsection} to subdivide a reference section into segments. In this case, the segments are local to the enclosing \env{refsection} environment. Use the \env{segment} option of \cmd{printbibliography} to select a segment when printing the bibliography. Within a section, the reference segments are numbered starting at~\texttt{1} and the number of the current segment will be written to the transcript file. All citations given outside a \env{refsegment} environment are assigned to segment~0. In contrast to the \env{refsection} environment, the current segment is not selected automatically if \cmd{printbibliography} is used within a \env{refsegment} environment.

\csitem{newrefsegment}

This command is similar to the \env{refsegment} environment except that it is a stand"=alone command rather than an environment. It automatically ends the previous reference segment (if any) and immediately starts a new one. Note that the reference segment started by the last \cmd{newrefsegment} command will extend to the end of the document. Use \cmd{endrefsegment} if you want to terminate it earlier.

\end{ltxsyntax}

\subsubsection{Bibliography Categories}
\label{use:bib:cat}

Bibliography categories allow you to split the bibliography into multiple parts dedicated to different topics or different types of references, for example primary and secondary sources. See \secref{use:use:div} for usage examples.

\begin{ltxsyntax}

\cmditem{DeclareBibliographyCategory}{category}

Declares a new \prm{category}, to be used in conjunction with \cmd{addtocategory} and the
\opt{category} and \opt{notcategory} filters of \cmd{printbibliography}. This command is used in the document preamble.

\cmditem{addtocategory}{category}{key}

Assigns a \prm{key} to a \prm{category}, to be used in conjunction with the \opt{category} and \opt{notcategory} filters of \cmd{printbibliography}. This command may be used in the preamble and in the document body. The \prm{key} may be a single entry key or a comma"=separated list of keys. The assignment is global.

\end{ltxsyntax}

\subsubsection{Bibliography Headings and Environments}
\label{use:bib:hdg}

\begin{ltxsyntax}

\cmditem{defbibenvironment}{name}{begin code}{end code}{item code}

This command defines bibliography environments. The \prm{name} is an identifier passed to the \opt{env} option of \cmd{printbibliography} and \cmd{printbiblist} when selecting the environment. The \prm{begin code} is \latex code to be executed at the beginning of the environment; the \prm{end code} is executed at the end of the environment; the \prm{item code} is code to be executed at the beginning of each entry in the bibliography or a bibliography list. Here is an example of a definition based on the standard \latex \env{list} environment:

\begin{ltxexample}
\defbibenvironment{bibliography}
  {\list{}
     {\setlength{\leftmargin}{\bibhang}%
      \setlength{\itemindent}{-\leftmargin}%
      \setlength{\itemsep}{\bibitemsep}%
      \setlength{\parsep}{\bibparsep}}}
  {\endlist}
  {\item}
\end{ltxexample}
%
As seen in the above example, usage of \cmd{defbibenvironment} is roughly similar to \cmd{newenvironment} except that there is an additional mandatory argument for the \prm{item code}.

\cmditem{defbibheading}{name}[title]{code}

This command defines bibliography headings. The \prm{name} is an identifier to be passed to the \opt{heading} option of \cmd{printbibliography} or \cmd{printbibheading} and \cmd{printbiblist} when selecting the heading. The \prm{code} should be \latex code generating a fully"=fledged heading, including page headers and an entry in the table of contents, if desired. If \cmd{printbibliography} or \cmd{printbiblist} are invoked with a \opt{title} option, the title will be passed to the heading definition as |#1|. If not, the default title specified by the optional \prm{title} argument is passed as |#1| instead. The \prm{title} argument will typically be \cmd{bibname}, \cmd{refname}, or \cmd{biblistname} (see \secref{aut:lng:key:bhd}). This command is often needed after changes to document headers in the preamble. Here is an example of a simple heading definition:

\begin{ltxexample}
\defbibheading{bibliography}[\bibname]{%
  \chapter*{#1}%
  \markboth{#1}{#1}}
\end{ltxexample}

\end{ltxsyntax}

The following headings, which are intended for use with \cmd{printbibliography} and \cmd{printbibheading}, are predefined:

\begin{valuelist*}

\item[bibliography]
This is the default heading used by \cmd{printbibliography} if the \opt{heading} option is not given. Its default definition depends on the document class. If the class provides a \cmd{chapter} command, the heading is similar to the bibliography heading of the standard \latex \texttt{book} class, \ie it uses \cmd{chapter*} to create an unnumbered chapter heading which is not included in the table of contents. If there is no \cmd{chapter} command, it is similar to the bibliography heading of the standard \latex \texttt{article} class, \ie it uses \cmd{section*} to create an unnumbered section heading which is not included in the table of contents. The string used in the heading also depends on the document class. With \texttt{book}-like classes the localisation string \texttt{bibliography} is used, with other classes it is \texttt{references} (see \secref{aut:lng:key}). See also \secref{use:cav:scr, use:cav:mem} for class-specific hints.

\item[subbibliography]
Similar to \texttt{bibliography} but one sectioning level lower. This heading definition uses \cmd{section*} instead of \cmd{chapter*} with a \texttt{book}-like class and \cmd{subsection*} instead of \cmd{section*} otherwise.

\item[bibintoc]
Similar to \texttt{bibliography} above but adds an entry to the table of contents.

\item[subbibintoc]
Similar to \texttt{subbibliography} above but adds an entry to the table of contents.

\item[bibnumbered]
Similar to \texttt{bibliography} above but uses \cmd{chapter} or \cmd{section} to create a numbered heading which is also added to the table of contents.

\item[subbibnumbered]
Similar to \texttt{subbibliography} above but uses \cmd{section} or \cmd{subsection} to create a numbered heading which is also added to the table of contents.

\item[none]
A blank heading definition. Use this to suppress the heading.

\end{valuelist*}

The following headings intended for use with \cmd{printbiblist} are predefined:

\begin{valuelist*}

\item[biblist]
This is the default heading used by \cmd{printbiblist} if the \opt{heading} option is not given. It is similar to \texttt{bibliography} above except that it uses the localisation string \texttt{shorthands} instead of \texttt{bibliography} or \texttt{references} (see \secref{aut:lng:key}). See also \secref{use:cav:scr, use:cav:mem} for class-specific hints.

\item[biblistintoc]
Similar to \texttt{biblist} above but adds an entry to the table of contents.

\item[biblistnumbered]
Similar to \texttt{biblist} above but uses \cmd{chapter} or \cmd{section} to create a numbered heading which is also added to the table of contents.

\end{valuelist*}

\subsubsection{Bibliography Notes}
\label{use:bib:nts}

\begin{ltxsyntax}

\cmditem{defbibnote}{name}{text}

Defines the bibliography note \prm{name}, to be used via the \opt{prenote} and \opt{postnote} options of \cmd{printbibliography} and \cmd{printbiblist}. The \prm{text} may be any arbitrary piece of text, possibly spanning several paragraphs and containing font declarations. Also see \secref{use:cav:act}.

\end{ltxsyntax}

\subsubsection{Bibliography Filters and Checks}
\label{use:bib:flt}

\begin{ltxsyntax}

\cmditem{defbibfilter}{name}{expression}

Defines the custom bibliography filter \prm{name}, to be used via the \opt{filter} option of \cmd{printbibliography}. The \prm{expression} is a complex test based on the logical operators \texttt{and}, \texttt{or}, \texttt{not}, the group separator \texttt{(...)}, and the following atomic tests:

\end{ltxsyntax}

\begin{optionlist*}

\valitem{segment}{integer}

Matches all entries cited in reference segment \prm{integer}.

\valitem{type}{entrytype}

Matches all entries whose entry type is \prm{entrytype}.

\valitem{subtype}{subtype}

Matches all entries whose \bibfield{entrysubtype} is \prm{subtype}.

\valitem{keyword}{keyword}

Matches all entries whose \bibfield{keywords} field includes \prm{keyword}. If the \prm{keyword} contains spaces, it needs to be wrapped in braces.

\valitem{category}{category}

Matches all entries assigned to \prm{category} with \cmd{addtocategory}.

\end{optionlist*}

Here is an example of a filter expression:

\begin{ltxexample}[style=latex,keywords={and,or,not,type,keyword}]
\defbibfilter{example}{%
  ( type=book or type=inbook )
  and keyword=abc
  and not keyword={x y z}
}
\end{ltxexample}
%
This filter will match all entries whose entry type is either \bibtype{book} or \bibtype{inbook} and whose \bibfield{keywords} field includes the keyword <\texttt{abc}> but not <\texttt{x y z}>. As seen in the above example, all elements are separated by whitespace (spaces, tabs, or line endings). There is no spacing around the equal sign. The logical operators are evaluated with the \cmd{ifboolexpr} command from the \sty{etoolbox} package. See the \sty{etoolbox} manual for details about the syntax. The syntax of the \cmd{ifthenelse} command from the \sty{ifthen} package, which has been employed in older versions of \biblatex, is still supported. This is the same test using \sty{ifthen}-like syntax:

\begin{ltxexample}[style=ifthen,morekeywords={\\type,\\keyword}]
\defbibfilter{example}{%
  \( \type{book} \or \type{inbook} \)
  \and \keyword{abc}
  \and \not \keyword{x y z}
}
\end{ltxexample}
%
Note that custom filters are local to the reference section in which they are used. Use the \texttt{section} filter of \cmd{printbibliography} to select a different section. This is not possible from within a custom filter.

\begin{ltxsyntax}

\cmditem{defbibcheck}{name}{code}

Defines the custom bibliography filter \prm{name}, to be used via the \opt{check} option of \cmd{printbibliography}. \cmd{defbibcheck} is similar in concept to \cmd{defbibfilter} but much more low-level. Rather than a high-level expression, the \prm{code} is \latex code, much like the code used in driver definitions, which may perform arbitrary tests to decide whether or not a given entry is to be printed. The bibliographic data of the respective entry is available when the \prm{code} is executed. Issuing the command \cmd{skipentry} in the \prm{code} will cause the current entry to be skipped. For example, the following filter will only output entries with an \bibfield{abstract} field:

\begin{ltxexample}
\defbibcheck{<<abstract>>}{%
  \iffieldundef{abstract}{<<\skipentry>>}{}}
...
\printbibliography[<<check=abstract>>]
\end{ltxexample}
%
The following check will exclude all entries published before the year 2000:

\begin{ltxexample}
\defbibcheck{recent}{%
  \iffieldint{year}
    {\ifnumless{\thefield{year}}{2000}
       {\skipentry}
       {}}
    {\skipentry}}
\end{ltxexample}
%
See the author guide, in particular \secref{aut:aux:tst,aut:aux:ife}, for further details.

\end{ltxsyntax}

\subsubsection{Reference Contexts}
\label{use:bib:context}

References in a bibliography are cited and printed in a <context>. The context determines the data which is actually used to cite or provide bibliographic data for an entry. A context consists of the following information:

\begin{itemize}
 \item A sorting template
 \item A template for constructing the sorting keys for names
 \item A string prefix for citation schemes which use alphabetic or numeric labels
 \item A template for calculating name uniqueness information
 \item A template for constructing alphabetic labels for names
\end{itemize}
%
The purpose of bibliography contexts is twofold. Firstly, they are used to set options which influence a printed bibliography and secondly to influence the data printed by citation commands.
The former use is the most common when one needs to print more than one bibliography list with different, for example, sorting.

\begin{ltxexample}
\usepackage[sorting=nyt]{biblatex}
\begin{document}
\cite{one}
\cite{two}
\printbibliography
\newrefcontext[sorting=ydnt]
\printbibliography
\end{ltxexample}
%
Here we print two bibliographies, one with the default <nyt> sorting template and one with the <ydnt> sorting template.

To demonstrate the second type of use of bibliography contexts, we have to understand that the actual data for an entry can vary depending on the context. This is most obvious in the case of the \bibfield{extra*} fields like \bibfield{extradate} which are generated by the backend according to the order of entries \emph{after} sorting so that they come out in the expected <a, b, c> order. This clearly shows that the \emph{data} in an entry can be different between sorting templates. If a document contains more than one bibliography list with different sorting templates, it can happen then that the \file{.bbl} contains sorting lists with the same entry but containing different data (a different value for \bibfield{extradate}, for example). The purpose of bibliography contexts is to encapsulate things inside a context so that \biblatex can use the correct entry data. An example is printing a bibliography list with a different sorting order to the global sorting order where the \opt{extra*} fields are different for the same entry between sorting lists:

\begin{ltxexample}
\usepackage[sorting=nyt,style=authoryear]{biblatex}
\DeclareSortingTemplate{yntd}{
  \sort{
    \field[strside=left,strwidth=4]{sortyear}
    \field[strside=left,strwidth=4]{year}
    \literal{9999}
  }
  \sort{
    \field{sortname}
    \field{author}
    \field{editor}
  }
  \sort[direction=descending]{
    \field{sorttitle}
    \field{title}
  }
}
\begin{document}
\cite{one}
\cite{two}
\printbibliography
\newrefcontext[sorting=yntd]
\cite{one}
\cite{two}
\printbibliography
\end{ltxexample}
%
Here, the second use of the citations, along with the \cmd{printbibliography} command will use data from the context of the custom <yntd> sorting template which may well be different from the data associated with the default <nyt> template. That is, the citation labels (in an authoryear style which uses \opt{extradate}) may be different \emph{for the exact same entries} between different bibliography contexts and so the citations themselves may look different.

Reference contexts can be declared with \cmd{DeclareRefcontext} and referred to by name, see below.

By default, data for a citation is drawn from the reference context of the last bibliography in which it was printed. For example:

\begin{ltxexample}[style=latex]
\DeclareRefcontext{ap}{labelprefix=A}
\begin{document}

\cite{book, article, misc}

\printbibliography[type=book]

\newrefcontext{ap}
\printbibliography[type=article]

\newrefcontext[sorting=ydnt]
\printbibliography[type=misc]

\end{document}
\end{ltxexample}
%
This example also shows the declaration and use of a named reference context. Assuming the entrykeys are indicative of their entrytypes, this is the default situation for the citations which corresponds to what users normally expect:

\begin{itemize}
\item The citation of entry \bibfield{book} would draw its data from the global reference context, because the last bibliography in which it was printed was the one in the global reference context.
\item The citation of entry \bibfield{article} would draw its data from reference context with \opt{labelprefix=A} and would therefore have a <A> prefix when cited.
\item The citation of entry \bibfield{misc} would draw its data from the reference context with \opt{sorting=ydnt}
\end{itemize}
%
In cases where the user has entries which occur in multiple bibliographies in different forms or with potentially different labels (in a numeric scheme with different \bibfield{labelprefix} values for example), it may be necessary to tell \biblatex from which reference context you wish to draw the citation information. As shown above this can be done by explicitly putting citations inside reference contexts. This can be onerous in a large document and so there is specific functionality for assigning citations to reference contexts programatically, see the \cmd{assignrefcontext*} macros below.

\begin{ltxsyntax}

\cmditem{DeclareRefcontext}{name}{key=value, \dots}

Declares a named reference context with name \prm{name}. The \prm{key=value} options define the context attributes. All context attributes are optional and default to the global settings if absent. The valid options are:

\begin{optionlist*}

\valitem{sorting}{name}

Specify a sorting template defined previously with \cmd{DeclareSortingTemplate}. This template is used to determine which data to retrieve and/or print for an entry in the commands inside the context.

\valitem{sortingnamekeytemplatename}{name}

Specify a sorting name key template defined previously with \cmd{DeclareSortingNamekeyTemplate}. This template is used to construct sorting keys for names inside the context. The template name can also be specified (in increasing order of preference) per"=entry, per"=name list and per"=name. See \secref{apx:opt} for information on setting per"=option, per"=namelist and per"=name options.

\valitem{uniquenametemplatename}{name}

Specify a uniquename template defined previously with \cmd{DeclareUniquenameTemplate} (see \secref{aut:cav:amb}). This template is used to calculate uniqueness information for names inside the context. The template name can also be specified (in increasing order of preference) per"=entry, per"=name list and per"=name. See \secref{apx:opt} for information on setting per"=option, per"=namelist and per"=name options.

\valitem{labelalphanametemplatename}{name}

Specify a template defined previously with \cmd{DeclareLabelalphaNameTemplate} (see \secref{aut:ctm:lab}). This template is used to construct name parts of alphabetic labels for names inside the context. The template name can also be specified (in increasing order of preference) per"=entry, per"=name list and per"=name. See \secref{apx:opt} for information on setting per"=option, per"=namelist and per"=name options.

\valitem{nametemplates}{name}

A convenience meta-option which sets \opt{sortingnamekeytemplate}, \opt{uniquenametemplate} and \opt{labelalphanametemplate} to the same template name. This option can also be specified (in increasing order of preference) per"=entry, per"=name list and per"=name. See \secref{apx:opt} for information on setting per"=option, per"=namelist and per"=name options.

\valitem{labelprefix}{string}

This option applies to numerical citation\slash bibliography styles only and requires that the \opt{defernumbers} option from \secref{use:opt:pre:gen} be enabled globally. Setting this option will implicitly enable \opt{resetnumbers} for the any \cmd{printbibliography} in the scope of the context (unless overridden by a user-specified value for \opt{resetnumbers}). The option assigns the \prm{string} as a prefix to all entries in the reference context. For example, if the \prm{string} is \texttt{A}, the numerical labels printed will be \texttt{[A1]}, \texttt{[A2]}, \texttt{[A3]}, etc. This is useful for subdivided numerical bibliographies where each subbibliography uses a different prefix. The \prm{string} is available to styles in the \bibfield{labelprefix} field of all affected entries. Note that the \prm{string} is fully expanded, which means that you can use context-dependent macros like \cmd{thechapter}, but not unexpandable commands such as \cmd{dag}. If you need to pass unexpandable code to \prm{string}, protect it from expansion with \cmd{detokenize}. See \secref{aut:bbx:fld:lab} for details.

\end{optionlist*}
%

\envitem{refcontext}[key=value, \dots]{name}

Wraps a reference context environment. The possible \prm{key=value} optional arguments are as for \cmd{DeclareRefcontext} and override options given for the named reference context \prm{name}. \prm{name} can also be omitted as \verb+{}+ or by omitting even the empty braces\footnote{This slightly odd syntax possibility is a result of backwards compatibility with \biblatex $<$3.5}.

The \opt{refcontext} environment cannot be nested and \biblatex will generate an error if you try to do so.

\cmditem{newrefcontext}[key=value, \dots]{name}

This command is similar to the \env{refcontext} environment except that it is a stand"=alone command rather than an environment. It automatically ends any previous reference context section begun with \cmd{newrefcontext} (if any) and immediately starts a new one. Note that the context section started by the last \cmd{newrefcontext} command in the document will extend to the very end of the document. Use \cmd{endrefcontext} if you want to terminate it earlier.
\end{ltxsyntax}
%
At the beginning of the document, there is always a global context containing global settings for each of the reference context options. Here is an example summarising the reference contexts with various settings:

\begin{ltxexample}[style=latex]
\usepackage[sorting=nty]{biblatex}

\DeclareRefcontext{testrc}{sorting=nyt}

% Global reference context:
%   sorting=nty
%   sortingnamekeytemplate=global
%   labelprefix=

\begin{document}

\begin{refcontext}{testrc}
% reference context:
%   sorting=nyt
%   sortingnamekeytemplate=global
%   labelprefix=
\end{refcontext}

\begin{refcontext}[labelprefix=A]{testrc}
% reference context:
%   sorting=nyt
%   sortingnamekeytemplate=global
%   labelprefix=A
\end{refcontext}

\begin{refcontext}[sorting=ydnt,labelprefix=A]
% reference context:
%   sorting=ydnt
%   sortingnamekeytemplate=global
%   labelprefix=A
\end{refcontext}

\newrefcontext}[labelprefix=B]
% reference context:
%   sorting=nty
%   sortingnamekeytemplate=global
%   labelprefix=B
\endrefcontext

\newrefcontext}[sorting=ynt,labelprefix=C]{testrc}
% reference context:
%   sorting=ynt
%   sortingnamekeytemplate=global
%   labelprefix=C
\endrefcontext
\end{ltxexample}

\begin{ltxsyntax}

\cmditem{assignrefcontextkeyws}[key=value, \dots]{keyword1,keyword2, ...}
\cmditem{assignrefcontextkeyws*}[key=value, \dots]{keyword1,keyword2, ...}
\cmditem{assignrefcontextcats}[key=value, \dots]{category1, category2, ...}
\cmditem{assignrefcontextcats*}[key=value, \dots]{category1, category2, ...}
\cmditem{assignrefcontextentries}[key=value, \dots]{entrykey1, entrykey2, ...}
\cmditem{assignrefcontextentries*}[key=value, \dots]{entrykey1, entrykey2, ...}
\cmditem{assignrefcontextentries}[key=value, \dots]{*}
\cmditem{assignrefcontextentries*}[key=value, \dots]{*}

\end{ltxsyntax}
These commands automate putting citations into refcontexts when the default behaviour is not sufficient. The \prm{key=value} options are as for \cmd{DeclareRefcontext} with the addition of the \kvopt{name}{refcontextname} option which sets all options from those defined for the named refcontext \prm{refcontextname}. Use \kvopt{name}{default} to use the global default refcontext options. The specific \keyval options override those set by any named \prm{refcontextname}. The default behaviour is that the data for a citation is drawn from the refcontext of the most recently processed bibliography in which it was printed\footnote{This does not always mean what one might think. In a document containing multiple bibliographies, the last bibliography will be the context for any citations \emph{before} the first bibliography because all bibliographies are processed when the \file{.bbl} is read.}. For citations that are used in some way but not printed in a bibliography or bibliography list, they default to drawing their data from the global refcontext established at the beginning of the document. To override this behaviour, instead of manually wrapping citation commands in \env{refcontext} environments, which might be error-prone and tedious, you can register a comma"=separated list of \prm{keywords}, \prm{categories} or \prm{entrykeys} which, respectively, make the entries with any of the specified keywords, entries in any of the specified categories (see \secref{use:use:div}) or entries with any of the specified citation keys draw their data from a particular named refcontext and/or specified \prm{refcontext key/values}. Such refcontext auto-assignments are specific to the current refsection. You may specify the same citation key in any of these commands but be aware that assignment is done in the order \prm{keywords}, \prm{categories}, \prm{entrykeys} with the later specifications overriding the earlier. \cmd{assignrefcontextentries} accepts a single asterisk instead of a list of entrykeys which allows the assignment of all keys in a refsection to a refcontext with having to explicitly list them. An example:

\begin{ltxexample}[style=latex]
\assignrefcontextentries[labelprefix=A]{key2}
\cite{key1}
\begin{refcontext}[labelprefix=B]
\cite{key2}
\end{refcontext}
\end{ltxexample}
%
Here, the data for the citation of \bibfield{key2} will be drawn from refcontext \opt{labelprefix=A} and not \opt{labelprefix=B} (resulting in a label with prefix <A> and not <B>).
The starred versions do not override a local refcontext and so with:

\begin{ltxexample}[style=latex]
\assignrefcontextentries*[labelprefix=A]{key2}
\cite{key1}
\begin{refcontext}[labelprefix=B]
\cite{key2}
\end{refcontext}
\end{ltxexample}
%
the data for the citation of \bibfield{key2} will be drawn from refcontext \opt{labelprefix=B}. Note that these commands are rarely necessary unless you have multiple bibliographies in which the same citations occur and \biblatex\ cannot by default tell which bibliography list a citation should refer to. See the example file \file{94-labelprefix.tex} for more details.

\begin{ltxexample}[style=latex]
\DeclareRefcontext{testrc}{labelprefix=A}
\assignrefcontextentries[name=testrc]{key2}
\cite{key1}
\begin{refcontext}[labelprefix=B]
\cite{key2}
\end{refcontext}
\end{ltxexample}
%
Here, the data for the citation of \bibfield{key2} will be drawn from the refcontext named <testrc> which has \opt{labelprefix=A} and not \opt{labelprefix=B} (resulting in a label with prefix <A> and not <B>).

\begin{ltxexample}[style=latex]
\DeclareRefcontext{testrc}{labelprefix=A}
\assignrefcontextentries[name=testrc,labelprefix=C]{key2}
\cite{key1}
\begin{refcontext}[labelprefix=B]
\cite{key2}
\end{refcontext}
\end{ltxexample}
%
Here, the data for the citation of \bibfield{key2} will be drawn from refcontext with \opt{labelprefix=C} and not \opt{labelprefix=A} since the explicit options override the named refcontext (resulting in a label with prefix <C> and not <A> or <B>).

\subsubsection{Dynamic Entry Sets}
\label{use:bib:set}

In addition to the \bibtype{set} entry type, \biblatex also supports dynamic entry sets defined on a per-document\slash per-refsection basis. The following command, which may be used in the document preamble or the document body, defines the set \prm{key}:

\begin{ltxsyntax}

\cmditem{defbibentryset}{key}{key1,key2,key3, \dots}

The \prm{key} is the entry key of the set, which is used like any other entry key when referring to the set. The \prm{key} must be unique and it must not conflict with any other entry key. The second argument is a comma"=separated list of the entry keys which make up the set. \cmd{defbibentryset} implies the equivalent of a \cmd{nocite} command, \ie all sets which are declared are also added to the bibliography. When declaring the same set more than once, only the first invocation of \cmd{defbibentryset} will define the set. Subsequent definitions of the same \prm{key} are ignored and work like \cmd{nocite}\prm{key}. Dynamic entry sets defined in the document body are local to the enclosing \env{refsection} environment, if any. Otherwise, they are assigned to reference section~0. Those defined in the preamble are assigned to reference section~0. See \secref{use:use:set} for further details.

\end{ltxsyntax}

\subsection{Citation Commands}
\label{use:cit}

All citation commands generally take one mandatory and two optional arguments. The \prm{prenote} is text to be printed at the beginning of the citation. This is usually a notice such as <see> or <compare>. The \prm{postnote} is text to be printed at the very end of the citation. This is usually a page number. If only one of these arguments is given, it is taken as a postnote. If you want to specify a prenote but no postnote, you need to leave the second optional argument empty, as in |\cite[see][]{key}|. The \prm{key} argument to all citation commands is mandatory. This is the entry key or a comma"=separated list of keys corresponding to the entry keys in the \file{bib} file. In sum, all basic citations commands listed further down have the following syntax:

\begin{ltxsyntax}

\cmditem*{command}[prenote][postnote]{keys}<punctuation>

If the \opt{autopunct} package option from \secref{use:opt:pre:gen} is enabled, they will scan ahead for any \prm{punctuation} immediately following their last argument. This is useful to avoid spurious punctuation marks after citations. This feature is configured with \cmd{DeclareAutoPunctuation}, see \secref{aut:pct:cfg} for details.

\end{ltxsyntax}

\subsubsection{Standard Commands}
\label{use:cit:std}

The following commands are defined by the citation style. Citation styles may provide any arbitrary number of specialized commands, but these are the standard commands typically provided by general-purpose styles.

\begin{ltxsyntax}

\cmditem{cite}[prenote][postnote]{key}
\cmditem{Cite}[prenote][postnote]{key}

These are the bare citation commands. They print the citation without any additions such as parentheses. The numeric and alphabetic styles still wrap the label in square brackets since the reference may be ambiguous otherwise. \cmd{Cite} is similar to \cmd{cite} but capitalizes the name prefix of the first name in the citation if the \opt{useprefix} option is enabled, provided that there is a name prefix and the citation style prints any name at all.

\cmditem{parencite}[prenote][postnote]{key}
\cmditem{Parencite}[prenote][postnote]{key}

These commands use a format similar to \cmd{cite} but enclose the entire citation in parentheses. The numeric and alphabetic styles use square brackets instead. \cmd{Parencite} is similar to \cmd{parencite} but capitalizes the name prefix of the first name in the citation if the \opt{useprefix} option is enabled, provided that there is a name prefix and the citation style prints any name at all.

\cmditem{footcite}[prenote][postnote]{key}
\cmditem{footcitetext}[prenote][postnote]{key}

These command use a format similar to \cmd{cite} but put the entire citation in a footnote and add a period at the end. In the footnote, they automatically capitalize the name prefix of the first name if the \opt{useprefix} option is enabled, provided that there is a name prefix and the citation style prints any name at all. \cmd{footcitetext} differs from \cmd{footcite} in that it uses \cmd{footnotetext} instead of \cmd{footnote}.

\end{ltxsyntax}

\subsubsection{Style-specific Commands}
\label{use:cit:cbx}

The following additional citation commands are only provided by some of the citation styles which come with this package.

\begin{ltxsyntax}

\cmditem{textcite}[prenote][postnote]{key}
\cmditem{Textcite}[prenote][postnote]{key}

These citation commands are provided by all styles that come with this package. They are intended for use in the flow of text, replacing the subject of a sentence. They print the authors or editors followed by a citation label which is enclosed in parentheses. Depending on the citation style, the label may be a number, the year of publication, an abridged version of the title, or something else. The numeric and alphabetic styles use square brackets instead of parentheses. In the verbose styles, the label is provided in a footnote. Trailing punctuation is moved between the author or editor names and the footnote mark. \cmd{Textcite} is similar to \cmd{textcite} but capitalizes the name prefix of the first name in the citation if the \opt{useprefix} option is enabled, provided that there is a name prefix.

\cmditem{smartcite}[prenote][postnote]{key}
\cmditem{Smartcite}[prenote][postnote]{key}

Like \cmd{parencite} in a footnote and like \cmd{footcite} in the body.

\cmditem{cite*}[prenote][postnote]{key}

This command is provided by all author-year and author-title styles. It is similar to the regular \cmd{cite} command but merely prints the year or the title, respectively.

\cmditem{parencite*}[prenote][postnote]{key}

This command is provided by all author-year and author-title styles. It is similar to the regular \cmd{parencite} command but merely prints the year or the title, respectively.

\cmditem{supercite}{key}

This command, which is only provided by the numeric styles, prints numeric citations as superscripts without brackets. It uses \cmd{supercitedelim} instead of \cmd{multicitedelim} as citation delimiter. Note that any \prm{prenote} and \prm{postnote} arguments are ignored. If they are given, \cmd{supercite} will discard them and issue a warning message.

\end{ltxsyntax}

\subsubsection{Qualified Citation Lists}
\label{use:cit:mlt}

This package supports a class of special citation commands called <multicite> commands. The point of these commands is that their argument is a list of citations where each item forms a fully qualified citation with a pre- and\slash or postnote. This is particularly useful with parenthetical citations and citations given in footnotes. It is also possible to assign a pre- and\slash or postnote to the entire list. The multicite commands are built on top of backend commands like \cmd{parencite} and \cmd{footcite}. The citation style provides a multicite definition with \cmd{DeclareMultiCiteCommand} (see \secref{aut:cbx:cbx}). The following example illustrates the syntax of multicite commands:

\begin{ltxexample}
\parencites[35]{key1}[88--120]{key2}[23]{key3}
\end{ltxexample}
%
The format of the arguments is similar to that of the regular citation commands, except that only one citation command is given. If only one optional argument is given for an item in the list, it is taken as a postnote. If you want to specify a prenote but no postnote, you need to leave the second optional argument of the respective item empty:

\begin{ltxexample}
\parencites[35]{key1}[chapter 2 in][]{key2}[23]{key3}
\end{ltxexample}
%
In addition to that, the entire citation list may also have a pre- and\slash or postnote. The syntax of these global notes differs from other optional arguments in that they are given in parentheses rather than the usual brackets:

\begin{ltxexample}
\parencites<<(>>and chapter 3<<)>>[35]{key1}[78]{key2}[23]{key3}
\parencites<<(>>Compare<<)()>>[35]{key1}[78]{key2}[23]{key3}
\parencites<<(>>See<<)(>>and the introduction<<)>>[35]{key1}[78]{key2}[23]{key3}
\end{ltxexample}
%
Note that the multicite commands keep on scanning for arguments until they encounter a token that is not the start of an optional or mandatory argument. If a left brace or bracket follows a multicite command, you need to mask it by adding \cmd{relax} or a control space (a backslash followed by a space) after the last valid argument. This will cause the scanner to stop.

\begin{ltxexample}[style=latex,showspaces]
\parencites[35]{key1}[78]{key2}<<\relax>>[...]
\parencites[35]{key1}[78]{key2}<<\ >>{...}
\end{ltxexample}
%
By default, this package provides the following multicite commands which correspond to regular commands from \secref{use:cit:std, use:cit:cbx}:

\begin{ltxsyntax}

\cmditem{cites}(multiprenote)(multipostnote)[prenote][postnote]{key}|...|[prenote][postnote]{key}
\cmditem{Cites}(multiprenote)(multipostnote)[prenote][postnote]{key}|...|[prenote][postnote]{key}

The multicite version of \cmd{cite} and \cmd{Cite}, respectively.

\cmditem{parencites}(multiprenote)(multipostnote)[prenote][postnote]{key}|...|[prenote][postnote]{key}
\cmditem{Parencites}(multiprenote)(multipostnote)[prenote][postnote]{key}|...|[prenote][postnote]{key}

The multicite version of \cmd{parencite} and \cmd{Parencite}, respectively.

\cmditem{footcites}(multiprenote)(multipostnote)[prenote][postnote]{key}|...|[prenote][postnote]{key}
\cmditem{footcitetexts}(multiprenote)(multipostnote)[prenote][postnote]{key}|...|[prenote][postnote]{key}

The multicite version of \cmd{footcite} and \cmd{footcitetext}, respectively.

\cmditem{smartcites}(multiprenote)(multipostnote)[prenote][postnote]{key}|...|[prenote][postnote]{key}
\cmditem{Smartcites}(multiprenote)(multipostnote)[prenote][postnote]{key}|...|[prenote][postnote]{key}

The multicite version of \cmd{smartcite} and \cmd{Smartcite}, respectively.

\cmditem{textcites}(multiprenote)(multipostnote)[prenote][postnote]{key}|...|[prenote][postnote]{key}
\cmditem{Textcites}(multiprenote)(multipostnote)[prenote][postnote]{key}|...|[prenote][postnote]{key}

The multicite version of \cmd{textcite} and \cmd{Textcite}, respectively.

\cmditem{supercites}(multiprenote)(multipostnote)[prenote][postnote]{key}|...|[prenote][postnote]{key}

The multicite version of \cmd{supercite}. This command is only provided by the numeric styles.

\end{ltxsyntax}

\subsubsection{Style-independent Commands}
\label{use:cit:aut}

Sometimes it is desirable to give the citations in the source file in a format that is not tied to a specific citation style and can be modified globally in the preamble. The format of the citations is easily changed by loading a different citation style. However, when using commands such as \cmd{parencite} or \cmd{footcite}, the way the citations are integrated with the text is still effectively hard"=coded. The idea behind the \cmd{autocite} command is to provide higher"=level citation markup which makes global switching from inline citations to citations given in footnotes (or as superscripts) possible. The \cmd{autocite} command is built on top of backend commands like \cmd{parencite} and \cmd{footcite}. The citation style provides an \cmd{autocite} definition with \cmd{DeclareAutoCiteCommand} (see \secref{aut:cbx:cbx}). This definition may be activated with the \opt{autocite} package option from \secref{use:opt:pre:gen}. The citation style will usually initialize this package option to a value which is suitable for the style, see \secref{use:xbx:cbx} for details. Note that there are certain limits to high"=level citation markup. For example, inline author-year citation schemes often integrate citations so tightly with the text that it is virtually impossible to automatically convert them to footnotes. The \cmd{autocite} command is only applicable in cases in which you would normally use \cmd{parencite} or \cmd{footcite} (or \cmd{supercite}, with a numeric style). The citations should be given at the end of a sentence or a partial sentence, immediately preceding the terminal punctuation mark, and they should not be a part of the sentence in a grammatical sense (like \cmd{textcite}, for example).

\begin{ltxsyntax}

\cmditem{autocite}[prenote][postnote]{key}
\cmditem{Autocite}[prenote][postnote]{key}

In contrast to other citation commands, the \cmd{autocite} command does not only scan ahead for punctuation marks following its last argument to avoid double punctuation marks, it actually moves them around if required. For example, with \kvopt{autocite}{footnote}, a trailing punctuation mark will be moved such that the footnote mark is printed after the punctuation. \cmd{Autocite} is similar to \cmd{autocite} but capitalizes the name prefix of the first name in the citation if the \opt{useprefix} option is enabled, provided that there is a name prefix and the citation style prints any name at all.

\cmditem*{autocite*}[prenote][postnote]{key}
\cmditem*{Autocite*}[prenote][postnote]{key}

The starred variants of \cmd{autocite} do not behave differently from the regular ones. The asterisk is simply passed on to the backend command. For example, if \cmd{autocite} is configured to use \cmd{parencite}, then \cmd{autocite*} will execute \cmd{parencite*}.

\cmditem{autocites}(multiprenote)(multipostnote)[prenote][postnote]{key}|...|[prenote][postnote]{key}
\cmditem{Autocites}(multiprenote)(multipostnote)[prenote][postnote]{key}|...|[prenote][postnote]{key}

This is the multicite version of \cmd{autocite}. It also detects and moves punctuation if required. Note that there is no starred variant. \cmd{Autocites} is similar to \cmd{autocites} but capitalizes the name prefix of the first name in the citation if the \opt{useprefix} option is enabled, provided that there is a name prefix and the citation style prints any name at all.

\end{ltxsyntax}

\subsubsection{Text Commands}
\label{use:cit:txt}

The following commands are provided by the core of \biblatex. They are intended for use in the flow of text. Note that all text commands are excluded from citation tracking.

\begin{ltxsyntax}

\cmditem{citeauthor}[prenote][postnote]{key}
\cmditem*{citeauthor*}[prenote][postnote]{key}
\cmditem{Citeauthor}[prenote][postnote]{key}
\cmditem*{Citeauthor*}[prenote][postnote]{key}

These commands print the authors. Strictly speaking, it prints the \bibfield{labelname} list, which may be the
\bibfield{author}, the \bibfield{editor}, or the \bibfield{translator}. \cmd{Citeauthor} is similar to \cmd{citeauthor} but capitalizes the name prefix of the first name in the citation if the \opt{useprefix} option is enabled, provided that there is a name prefix. The starred variants effectively force maxcitenames to 1 for just this command on so only print the first name in the labelname list (potentially followed by the et al string if there are more names). This allows more natural textual flow when refering to a paper in the singular when otherwise \cmd{citeauthor} would generate a (naturally plural) list of names.

\cmditem{citetitle}[prenote][postnote]{key}
\cmditem*{citetitle*}[prenote][postnote]{key}

This command prints the title. It will use the abridged title in the \bibfield{shorttitle} field, if available. Otherwise it falls back to the full title found in the \bibfield{title} field. The starred variant always prints the full title.

\cmditem{citeyear}[prenote][postnote]{key}
\cmditem*{citeyear*}[prenote][postnote]{key}

This command prints the year (\bibfield{year} field or year component of \bibfield{date}). The starred variant includes the \bibfield{extradate} information, if any.

\cmditem{citedate}[prenote][postnote]{key}
\cmditem*{citedate*}[prenote][postnote]{key}

This command prints the full date (\bibfield{date} or \bibfield{year}). The starred variant includes the \bibfield{extradate} information, if any.

\cmditem{citeurl}[prenote][postnote]{key}

This command prints the \bibfield{url} field.

\cmditem{parentext}{text}

This command wraps the \prm{text} in context sensitive parentheses.

\cmditem{brackettext}{text}

This command wraps the \prm{text} in context sensitive brackets.

\end{ltxsyntax}

\subsubsection{Special Commands}
\label{use:cit:spc}

The following special commands are also provided by the core of \biblatex.

\begin{ltxsyntax}

\cmditem{nocite}{key}
\cmditem*{nocite}|\{*\}|

This command is similar to the standard \latex \cmd{nocite} command. It
adds the \prm{key} to the bibliography without printing a citation. If the
\prm{key} is an asterisk, all entries available in the in-scope bibliography datasource(s) are added to the bibliography. Like all other citation commands, \cmd{nocite} commands in the document body are local to the enclosing \env{refsection} environment, if any. In contrast to standard \latex, \cmd{nocite} may also be used in the document preamble. In this case, the references are assigned to reference section~0. For the purposes of ordering citations by appearance \cmd{nocite} will behave like all other cite commands, with the added rule that a \cmd{nocite} issued in the preamble is treated as coming before all explicit citations in reference section~0 from the document body.

\cmditem{fullcite}[prenote][postnote]{key}

This command uses the bibliography driver for the respective entry type to create a full citation similar to the bibliography entry. It is thus related to the bibliography style rather than the citation style.

\cmditem{footfullcite}[prenote][postnote]{key}

Similar to \cmd{fullcite} but puts the entire citation in a footnote and adds a period at the end.

\cmditem{volcite}[prenote]{volume}[pages]{key}
\cmditem{Volcite}[prenote]{volume}[pages]{key}

These commands are similar to \cmd{cite} and \cmd{Cite} but intended for references to multi"=volume works which are cited by volume and page number. Instead of the \prm{postnote}, they take a mandatory \prm{volume} and an optional \prm{pages} argument. Since they merely compose the postnote and pass it to the \cmd{cite} command provided by the citation style as a \prm{postnote} argument, these commands are style independent. The volume and pages/text portion are formatted with the macro \cmd{mkvolcitenote} when they are passed on to the citation command. Additionally they are made available in the special fields \bibfield{volcitevolume} and \bibfield{volcitevolume} (\secref{aut:cbx:fld}) The format of the volume portion is controlled by the field formatting directive \opt{volcitevolume}, the format of the pages/text portion is controlled by the field formatting directive \opt{volcitepages} (\secref{aut:fmt:ich}). The delimiter printed between the volume portion and the pages/text portion may be modified by redefining the macro \cmd{volcitedelim} (\secref{aut:fmt:fmt}).

\cmditem{volcites}(multiprenote)(multipostnote)[prenote]{volume}[pages]{key}|\\...|[prenote]{volume}[pages]{key}
\cmditem{Volcites}(multiprenote)(multipostnote)[prenote]{volume}[pages]{key}|\\...|[prenote]{volume}[pages]{key}

The multicite version of \cmd{volcite} and \cmd{Volcite}, respectively.

\cmditem{pvolcite}[prenote]{volume}[pages]{key}
\cmditem{Pvolcite}[prenote]{volume}[pages]{key}

Similar to \cmd{volcite} but based on \cmd{parencite}.

\cmditem{pvolcites}(multiprenote)(multipostnote)[prenote]{volume}[pages]{key}|\\...|[prenote]{volume}[pages]{key}
\cmditem{Pvolcites}(multiprenote)(multipostnote)[prenote]{volume}[pages]{key}|\\...|[prenote]{volume}[pages]{key}

The multicite version of \cmd{pvolcite} and \cmd{Pvolcite}, respectively.

\cmditem{fvolcite}[prenote]{volume}[pages]{key}
\cmditem{ftvolcite}[prenote]{volume}[pages]{key}

Similar to \cmd{volcite} but based on \cmd{footcite} and \cmd{footcitetext}, respectively.

\cmditem{fvolcites}(multiprenote)(multipostnote)[prenote]{volume}[pages]{key}|\\...|[prenote]{volume}[pages]{key}
\cmditem{Fvolcites}(multiprenote)(multipostnote)[prenote]{volume}[pages]{key}|\\...|[prenote]{volume}[pages]{key}

The multicite version of \cmd{fvolcite} and \cmd{Fvolcite}, respectively.

\cmditem{svolcite}[prenote]{volume}[pages]{key}
\cmditem{Svolcite}[prenote]{volume}[pages]{key}

Similar to \cmd{volcite} but based on \cmd{smartcite}.

\cmditem{svolcites}(multiprenote)(multipostnote)[prenote]{volume}[pages]{key}|\\...|[prenote]{volume}[pages]{key}
\cmditem{Svolcites}(multiprenote)(multipostnote)[prenote]{volume}[pages]{key}|\\...|[prenote]{volume}[pages]{key}

The multicite version of \cmd{svolcite} and \cmd{Svolcite}, respectively.

\cmditem{tvolcite}[prenote]{volume}[pages]{key}
\cmditem{Tvolcite}[prenote]{volume}[pages]{key}

Similar to \cmd{volcite} but based on \cmd{textcite}.

\cmditem{tvolcites}(multiprenote)(multipostnote)[prenote]{volume}[pages]{key}|\\...|[prenote]{volume}[pages]{key}
\cmditem{Tvolcites}(multiprenote)(multipostnote)[prenote]{volume}[pages]{key}|\\...|[prenote]{volume}[pages]{key}

The multicite version of \cmd{tvolcite} and \cmd{Tvolcite}, respectively.

\cmditem{avolcite}[prenote]{volume}[pages]{key}
\cmditem{Avolcite}[prenote]{volume}[pages]{key}

Similar to \cmd{volcite} but based on \cmd{autocite}.

\cmditem{avolcites}(multiprenote)(multipostnote)[prenote]{volume}[pages]{key}|\\...|[prenote]{volume}[pages]{key}
\cmditem{Avolcites}(multiprenote)(multipostnote)[prenote]{volume}[pages]{key}|\\...|[prenote]{volume}[pages]{key}

The multicite version of \cmd{avolcite} and \cmd{Avolcite}, respectively.

\cmditem{notecite}[prenote][postnote]{key}
\cmditem{Notecite}[prenote][postnote]{key}

These commands print the \prm{prenote} and \prm{postnote} arguments but no citation. Instead, a \cmd{nocite} command is issued for every \prm{key}. This may be useful for authors who incorporate implicit citations in their writing, only giving information not mentioned before in the running text, but who still want to take advantage of the automatic \prm{postnote} formatting and the implicit \cmd{nocite} function. This is a generic, style"=independent citation command. Special citation styles may provide smarter facilities for the same purpose. The capitalized version forces capitalization (note that this is only applicable if the note starts with a command which is sensitive to \biblatex's punctuation tracker).

\cmditem{pnotecite}[prenote][postnote]{key}
\cmditem{Pnotecite}[prenote][postnote]{key}

Similar to \cmd{notecite} but the notes are printed in parentheses.

\cmditem{fnotecite}[prenote][postnote]{key}

Similar to \cmd{notecite} but the notes are printed in a footnote.

\end{ltxsyntax}

\subsubsection{Low-level Commands}
\label{use:cit:low}

The following commands are also provided by the core of \biblatex. They grant access to all lists and fields at a lower level.

\begin{ltxsyntax}

\cmditem{citename}[prenote][postnote]{key}[format]{name list}

The \prm{format} is a formatting directive defined with \cmd{DeclareNameFormat}. Formatting directives are discussed in \secref{aut:bib:fmt}. If this optional argument is omitted, this command falls back to the format \texttt{citename}. The last argument is the name of a \prm{name list}, in the sense explained in \secref{bib:fld}.

\cmditem{citelist}[prenote][postnote]{key}[format]{literal list}

The \prm{format} is a formatting directive defined with \cmd{DeclareListFormat}. Formatting directives are discussed in \secref{aut:bib:fmt}. If this optional argument is omitted, this command falls back to the format \texttt{citelist}. The last argument is the name of a \prm{literal list}, in the sense explained in \secref{bib:fld}.

\cmditem{citefield}[prenote][postnote]{key}[format]{field}

The \prm{format} is a formatting directive defined with \cmd{DeclareFieldFormat}. Formatting directives are discussed in \secref{aut:bib:fmt}. If this optional argument is omitted, this command falls back to the format \texttt{citefield}. The last argument is the name of a \prm{field}, in the sense explained in \secref{bib:fld}.

\end{ltxsyntax}

\subsubsection{Miscellaneous Commands}
\label{use:cit:msc}

The commands in this section are little helpers related to citations.

\begin{ltxsyntax}

\csitem{citereset}

This command resets the citation style. This may be useful if the style replaces repeated citations with abbreviations like \emph{ibidem}, \emph{idem}, \emph{op. cit.}, etc. and you want to force a full citation at the beginning of a new chapter, section, or some other location. The command executes a style specific initialization hook defined with the \cmd{InitializeCitationStyle} command from \secref{aut:cbx:cbx}. It also resets the internal citation trackers of this package. The reset will affect the \cmd{ifciteseen}, \cmd{ifentryseen}, \cmd{ifciteibid}, and \cmd{ifciteidem} tests discussed in \secref{aut:aux:tst}. When used inside a \env{refsection} environment, the reset of the citation tracker is local to the current \env{refsection} environment. Also see the \opt{citereset} package option in \secref{use:opt:pre:gen}.

\csitem{citereset*}

Similar to \cmd{citereset} but only executes the style's initialization hook, without resetting the internal citation trackers.

\csitem{mancite}

Use this command to mark manually inserted citations if you mix automatically generated and manual citations. This is particularly useful if the citation style replaces repeated citations by an abbreviation like \emph{ibidem} which may get ambiguous or misleading otherwise. Always use \cmd{mancite} in the same context as the manual citation, \eg if the citation is given in a footnote, include \cmd{mancite} in the footnote. The \cmd{mancite} command executes a style specific reset hook defined with the \cmd{OnManualCitation} command from \secref{aut:cbx:cbx}. It also resets the internal <ibidem> and <idem> trackers of this package. The reset will affect the \cmd{ifciteibid} and \cmd{ifciteidem} tests discussed in \secref{aut:aux:tst}.

\csitem{pno}

This command forces a single page prefix in the \prm{postnote} argument to a citation command. See \secref{use:cav:pag} for further details and usage instructions. Note that this command is only available locally in citations and the bibliography.

\csitem{ppno}

Similar to \cmd{pno} but forces a range prefix. See \secref{use:cav:pag} for further details and usage instructions. Note that this command is only available locally in citations and the bibliography.

\csitem{nopp}

Similar to \cmd{pno} but suppresses all prefixes. See \secref{use:cav:pag} for further details and usage instructions. Note that this command is only available locally in citations and the bibliography.

\csitem{psq}

In the \prm{postnote} argument to a citation command, this command indicates a range of two pages where only the starting page is given. See \secref{use:cav:pag} for further details and usage instructions. The suffix printed is the localisation string \texttt{sequens}, see \secref{aut:lng:key}. The spacing inserted between the suffix and the page number may be modified by redefining the macro \cmd{sqspace}. The default is an unbreakable interword space. Note that this command is only available locally in citations and the bibliography.

\csitem{psqq}

Similar to \cmd{psq} but indicates an open-ended page range. See \secref{use:cav:pag} for further details and usage instructions. The suffix printed is the localisation string \texttt{sequentes}, see \secref{aut:lng:key}. This command is only available locally in citations and the bibliography.

\cmditem{pnfmt}{text}

This command formats its argument \prm{text} in the same format as \bibfield{postnote}. The command can be used to format a page range while adding additional text in the postnote argument of a cite command.

\begin{ltxexample}
\autocite[\pnfmt{378-381, 383} and more]{sigfridsson}
\end{ltxexample}

\cmditem{RN}{integer}

This command prints an integer as an uppercase Roman numeral. The formatting applied to the numeral may be modified by redefining the macro \cmd{RNfont}.

\cmditem{Rn}{integer}

Similar to \cmd{RN} but prints a lowercase Roman numeral. The formatting applied to the numeral may be modified by redefining the macro \cmd{Rnfont}.

\end{ltxsyntax}

\subsubsection{\sty{natbib} Compatibility Commands}
\label{use:cit:nat}

The \opt{natbib} package option loads a \sty{natbib} compatibility module. The module defines aliases for the citation commands provided by the \sty{natbib} package. This includes aliases for the core citation commands \cmd{citet} and \cmd{citep} as well as the variants \cmd{citealt} and \cmd{citealp}. The starred variants of these commands, which print the full author list, are also supported. The \cmd{cite} command, which is handled in a particular way by \sty{natbib}, is not treated in a special way. The text commands (\cmd{citeauthor}, \cmd{citeyear}, etc.) are also supported, as are all commands which capitalize the name prefix (\cmd{Citet}, \cmd{Citep}, \cmd{Citeauthor}, etc.). Aliasing with \cmd{defcitealias}, \cmd{citetalias}, and \cmd{citepalias} is possible as well. Note that the compatibility commands will not emulate the citation format of the \sty{natbib} package. They merely alias \sty{natbib}'s commands to functionally equivalent facilities of the \biblatex package. The citation format depends on the main citation style. However, the compatibility style will adapt \cmd{nameyeardelim} to match the default style of the \sty{natbib} package.

\subsubsection[\sty{mcite}-like Citation Commands]{\sty{mcite}-like Citation Commands}
\label{use:cit:mct}

The \opt{mcite} package option loads a special citation module which provides \sty{mcite}\slash \sty{mciteplus}-like citation commands. Strictly speaking, what the module provides are wrappers for the commands of the main citation style. For example, the following command:

\begin{ltxexample}
\mcite{key1,setA,*keyA1,*keyA2,*keyA3,key2,setB,*keyB1,*keyB2,*keyB3}
\end{ltxexample}
%
is essentially equivalent to this:

\begin{ltxexample}
\defbibentryset{setA}{keyA1,keyA2,keyA3}%
\defbibentryset{setB}{keyB1,keyB2,keyB3}%
\cite{key1,setA,key2,setB}
\end{ltxexample}
%
The \cmd{mcite} command will work with any style since the \cmd{cite} backend command is controlled by the main citation style as usual. The \texttt{mcite} module provides wrappers for the standard commands in \secref{use:cit:std,use:cit:cbx}. See \tabref{use:cit:mct:tab2} for an overview. Pre and postnotes as well as starred variants of all commands are also supported. The parameters will be passed to the backend command. For example:

\begin{ltxexample}
\mcite*[pre][post]{setA,*keyA1,*keyA2,*keyA3}
\end{ltxexample}
%
will execute:

\begin{ltxexample}
\defbibentryset{setA}{keyA1,keyA2,keyA3}%
\cite*[pre][post]{setA}
\end{ltxexample}
%
Note that the \texttt{mcite} module is not a compatibility module. It provides commands which are very similar but not identical in syntax and function to \sty{mcite}'s commands. When migrating from \sty{mcite}\slash\sty{mciteplus} to \biblatex, legacy files must be updated. With \sty{mcite}, the first member of the citation group is also the identifier of the group as a whole. Borrowing an example from the \sty{mcite} manual, this group:

\begin{table}
\caption{\sty{mcite}-like commands}
\label{use:cit:mct:tab1}
\tablesetup
\begin{tabular}{@{}V{0.5\textwidth}@{}V{0.5\textwidth}@{}}
\toprule
\multicolumn{1}{@{}H}{Standard Command} &
\multicolumn{1}{@{}H}{\sty{mcite}-like Command} \\
\cmidrule(r){1-1}\cmidrule{2-2}
|\cite|		& |\mcite| \\
|\Cite|		& |\Mcite| \\
|\parencite|	& |\mparencite| \\
|\Parencite|	& |\Mparencite| \\
|\footcite|	& |\mfootcite| \\
|\footcitetext|	& |\mfootcitetext| \\
|\textcite|	& |\mtextcite| \\
|\Textcite|	& |\Mtextcite| \\
|\supercite|	& |\msupercite| \\
|\autocite|	& |\mautocite| \\
|\Autocite|	& |\Mautocite| \\
\bottomrule
\end{tabular}
\end{table}

\begin{ltxexample}
\cite{<<glashow>>,*salam,*weinberg}
\end{ltxexample}
%
consists of three entries and the entry key of the first one also serves as identifier of the entire group. In contrast to that, a \biblatex entry set is an entity in its own right. Therefore, it requires a unique entry key which is assigned to the set as it is defined:

\begin{ltxexample}
\mcite{<<set1>>,*glashow,*salam,*weinberg}
\end{ltxexample}
%
Once defined, an entry set is handled like any regular entry in a \file{bib} file. When using one of the \texttt{numeric} styles which come with \biblatex and activating its \opt{subentry} option, it is even possible to refer to set members. See \tabref{use:cit:mct:tab2} for some examples. Restating the original definition of the set is redundant, but permissible. In contrast to \sty{mciteplus}, however, restating a part of the original definition is invalid. Use the entry key of the set instead.

\begin{table}
\caption[\sty{mcite}-like syntax]
{\sty{mcite}-like syntax (sample output with \kvopt{style}{numeric} and \opt{subentry} option)}
\label{use:cit:mct:tab2}
\tablesetup
\begin{tabular}{@{}V{0.6\textwidth}@{}V{0.1\textwidth}@{}p{0.3\textwidth}@{}}
\toprule
\multicolumn{1}{@{}H}{Input} &
\multicolumn{1}{@{}H}{Output} &
\multicolumn{1}{@{}H}{Comment} \\
\cmidrule(r){1-1}\cmidrule(r){2-2}\cmidrule{3-3}
|\mcite{set1,*glashow,*salam,*weinberg}|& [1]	& Defining and citing the set \\
|\mcite{set1}|				& [1]	& Subsequent citation of the set \\
|\cite{set1}|				& [1]	& Regular |\cite| works as usual \\
|\mcite{set1,*glashow,*salam,*weinberg}|& [1]	& Redundant, but permissible \\
|\mcite{glashow}|			& [1a]	& Citing a set member \\
|\cite{weinberg}|			& [1c]	& Regular |\cite| works as well \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Localization Commands}
\label{use:lng}

The \biblatex package provides translations for key terms such as <edition> or <volume> as well as definitions for language specific features such as the date format and ordinals. These definitions, which are loaded automatically, may be modified or extended in the document preamble or the configuration file with the commands introduced in this section.

\begin{ltxsyntax}

\cmditem{DefineBibliographyStrings}{language}{definitions}

This command is used to define localisation strings. The \prm{language} must be a language name known to the \sty{babel}/\sty{polyglossia} packages, \ie one of the identifiers listed in \tabref{bib:fld:tab1} on page \pageref{bib:fld:tab1}. The \prm{definitions} are \keyval pairs which assign an expression to an identifier:

\begin{ltxexample}
\DefineBibliographyStrings{american}{%
  bibliography  = {Bibliography},
  shorthands    = {Abbreviations},
  editor        = {editor},
  editors       = {editors},
}
\end{ltxexample}
%
A complete list of all keys supported by default is given is \secref{aut:lng:key}. Note that all expressions should be capitalized as they usually are when used in the middle of a sentence. The \biblatex package will automatically capitalize the first word when required at the beginning of a sentence. Expressions intended for use in headings should be capitalized in a way that is suitable for titling. In contrast to \cmd{DeclareBibliographyStrings}, \cmd{DefineBibliographyStrings} overrides both the full and the abbreviated version of the string. See \secref{aut:lng:cmd} for further details.

\cmditem{DefineBibliographyExtras}{language}{code}

This command is used to adapt language specific features such as the date format and ordinals. The \prm{language} must be a language name known to the \sty{babel}/\sty{polyglossia} packages. The \prm{code}, which may be arbitrary \latex code, will usually consist of redefinitions of the formatting commands from \secref{use:fmt:lng}.

\cmditem{UndefineBibliographyExtras}{language}{code}

This command is used to restore the original definition of any commands modified with \cmd{DefineBibliographyExtras}. If a redefined command is included in \secref{use:fmt:lng}, there is no need to restore its previous definition since these commands are adapted by all language modules anyway.

\cmditem{DefineHyphenationExceptions}{language}{text}

This is a \latex frontend to \tex's \cmd{hyphenation} command which defines hyphenation exceptions.
The \prm{language} must be a language name known to the \sty{babel}/\sty{polyglossia} packages. The \prm{text} is a whitespace"=separated list of words. Hyphenation points are marked with a dash:

\begin{ltxexample}
\DefineHyphenationExceptions{american}{%
  hy-phen-ation ex-cep-tion
}
\end{ltxexample}

\cmditem{NewBibliographyString}{key}

This command declares new localisation strings, \ie it initializes a new \prm{key} to be used in the
\prm{definitions} of \cmd{DefineBibliographyStrings}. The \prm{key} argument may also be a comma"=separated
list of key names. The keys listed in \secref{aut:lng:key} are defined by default.

\end{ltxsyntax}

\subsection{Entry Querying Commands}
\label{use:eq}
The commands in this section are user-facing equivalents of the identically-named commands in section \secref{aut:aux:tst}. They can be used to test for the presence and attributes of specific bibliography entries. See section \secref{aut:aux:tst} for usage.

\begin{ltxsyntax}
\cmditem{ifentryseen}{entrykey}{true}{false}
\cmditem{ifentryinbib}{entrykey}{true}{false}
\cmditem{ifentrycategory}{entrykey}{category}{true}{false}
\cmditem{ifentrykeyword}{entrykey}{keyword}{true}{false}
\end{ltxsyntax}

\subsection{Formatting Commands}
\label{use:fmt}

The commands and facilities presented in this section may be used to adapt the format of citations and the bibliography.

\subsubsection{Generic Commands and Hooks}
\label{use:fmt:fmt}

The commands in this section may be redefined with \cmd{renewcommand} in the document preamble. Those marked as <Context Sensitive> in the margin can also be customised with \cmd{DeclareDelimFormat} and are printed with \cmd{printdelim} (\secref{use:fmt:csd}). Note that all commands starting with \cmd{mk\dots} take one argument. All of these commands are defined in \path{biblatex.def}.

\begin{ltxsyntax}

\csitem{bibsetup}
Arbitrary code to be executed at the beginning of the bibliography, intended for commands which affect the layout of the bibliography.

\csitem{bibfont}
Arbitrary code setting the font used in the bibliography. This is very similar to \cmd{bibsetup} but intended for switching fonts.

\csitem{citesetup}
Arbitrary code to be executed at the beginning of each citation command.

\csitem{newblockpunct}
The separator inserted between <blocks> in the sense explained in \secref{aut:pct:new}. The default definition is controlled by the package option \opt{block} (see \secref{use:opt:pre:gen}).

\csitem{newunitpunct}
The separator inserted between <units> in the sense explained in \secref{aut:pct:new}. This will usually be a period or a comma plus an interword space. The default definition is a period and a space.

\csitem{finentrypunct}
The punctuation printed at the very end of every bibliography entry, usually a period. The default definition is a period.

\csitem{entrysetpunct}
The punctuation printed between bibliography subentries of an entry set. The default definition is a semicolon and a space.

\csitem{bibnamedelima}
This delimiter controls the spacing between the elements which make up a name part. It is inserted automatically by the backend after the first name element if the element is less than three characters long and before the last element. The default definition is \cmd{addhighpenspace}, \ie a space penalized by the value of the \cnt{highnamepenalty} counter (\secref{use:fmt:len}). Please refer to \secref{use:cav:nam} for further details.

\csitem{bibnamedelimb}
This delimiter controls the spacing between the elements which make up a name part. It is inserted automatically by the backend between all name elements where \cmd{bibnamedelima} does not apply. The default definition is \cmd{addlowpenspace}, \ie a space penalized by the value of the \cnt{lownamepenalty} counter (\secref{use:fmt:len}). Please refer to \secref{use:cav:nam} for further details.

\csitem{bibnamedelimc}
This delimiter controls the spacing between name parts. The default name formats use it between the name prefix and the family name if \kvopt{useprefix}{true}. The default definition is \cmd{addhighpenspace}, \ie a space penalized by the value of the \cnt{highnamepenalty} counter (\secref{use:fmt:len}). Please refer to \secref{use:cav:nam} for further details.

\csitem{bibnamedelimd}
This delimiter controls the spacing between name parts. The default name formats use it between all name parts where \cmd{bibnamedelimc} does not apply. The default definition is \cmd{addlowpenspace}, \ie a space penalized by the value of the \cnt{lownamepenalty} counter (\secref{use:fmt:len}). Please refer to \secref{use:cav:nam} for further details.

\csitem{bibnamedelimi}
This delimiter replaces \cmd{bibnamedelima/b} after initials. Note that this only applies to initials given as such in the \file{bib} file, not to the initials automatically generated by \biblatex which use their own set of delimiters.

\csitem{bibinitperiod}
The punctuation inserted automatically by the backend after all initials unless \cmd{bibinithyphendelim} applies. The default definition is a period (\cmd{adddot}). Please refer to \secref{use:cav:nam} for further details.

\csitem{bibinitdelim}
The spacing inserted automatically by the backend between multiple initials unless \cmd{bibinithyphendelim} applies. The default definition is an unbreakable interword space. Please refer to \secref{use:cav:nam} for further details.

\csitem{bibinithyphendelim}
The punctuation inserted automatically by the backend between the initials of hyphenated name parts, replacing \cmd{bibinitperiod} and \cmd{bibinitdelim}. The default definition is a period followed by an unbreakable hyphen. Please refer to \secref{use:cav:nam} for further details.

\csitem{bibindexnamedelima}
Replaces \cmd{bibnamedelima} in the index.

\csitem{bibindexnamedelimb}
Replaces \cmd{bibnamedelimb} in the index.

\csitem{bibindexnamedelimc}
Replaces \cmd{bibnamedelimc} in the index.

\csitem{bibindexnamedelimd}
Replaces \cmd{bibnamedelimd} in the index.

\csitem{bibindexnamedelimi}
Replaces \cmd{bibnamedelimi} in the index.

\csitem{bibindexinitperiod}
Replaces \cmd{bibinitperiod} in the index.

\csitem{bibindexinitdelim}
Replaces \cmd{bibinitdelim} in the index.

\csitem{bibindexinithyphendelim}
Replaces \cmd{bibinithyphendelim} in the index.

\csitem{revsdnamepunct}
The punctuation to be printed between the given and family name parts when a name is reversed. The default is a comma. Here is an example showing a name with the default comma as \cmd{revsdnamedelim}:

\begin{ltxexample}
Jones<<,>> Edward
\end{ltxexample}

This command should be used with \cmd{bibnamedelimd} as a reversed-name separator in formatting directives for name lists. Please refer to \secref{use:cav:nam} for further details.

\csitem{bibnamedash}
The dash to be used as a replacement for recurrent authors or editors in the bibliography. The default is an <em> or an <en> dash, depending on the indentation of the list of references.

\csitem{labelnamepunct}\DeprecatedMark
A separator to be printed after the name used for alphabetizing in the bibliography (\bibfield{author} or \bibfield{editor}, if the \bibfield{author} field is undefined) instead of \cmd{newunitpunct}. The default is \cmd{newunitpunct}, \ie it is not handled differently from regular unit punctuation but permits convenient reconfiguration. This punctuation command is deprecated and has been superseded by the context-sensitive \cmd{nametitledelim} (see \secref{use:fmt:csd}). For backwards compatibility reasons, however, \cmd{nametitledelim} still defaults to \cmd{labelnamepunct} in the \texttt{bib} and \texttt{biblist} contexts. Style authors may want to consider replacing \cmd{labelnampunct} with \texttt{\textbackslash printdelim\{nametitledelim\}} and users may want to prefer modifying the context-sensitive \texttt{nametitledelim} in the \texttt{bib} context with \cmd{DeclareDelimFormat} over redefining \cmd{labelnamepunct}, e.g.
\begin{ltxexample}
\DeclareDelimFormat[bib]{nametitledelim}{%
  \addcolon\space}
\end{ltxexample}

\csitem{subtitlepunct}
The separator printed between the fields \bibfield{title} and \bibfield{subtitle}, \bibfield{booktitle} and \bibfield{booksubtitle}, as well as \bibfield{maintitle} and \bibfield{mainsubtitle}. With the default styles, this separator replaces \cmd{newunitpunct} at this location. The default definition is \cmd{newunitpunct}, \ie it is not handled differently from regular unit punctuation.

\csitem{intitlepunct}
The separator between the word in and the following title in entry types such as \bibtype{article}, \bibtype{inbook}, \bibtype{incollection}, etc. The default definition is a colon plus an interword space (\eg Article, in: \emph{Journal} or Title, in: \emph{Book}). Note that this is the separator string, not only the punctuation mark. If you don't want a colon after in, \cmd{intitlepunct} should still insert a space.

\csitem{bibpagespunct}
The separator printed before the \bibfield{pages} field. The default is a comma plus an interword space.

\csitem{bibpagerefpunct}
The separator printed before the \bibfield{pageref} field. The default is an interword space.

\csitem{bibeidpunct}
The separator printed before the \bibfield{eid} field (similar to \cmd{bibpagespunct}). The default is a comma plus an interword space.

\csitem{multinamedelim}\CSdelimMark
The delimiter printed between multiple items in a name list like \bibfield{author} or \bibfield{editor} if there are more than two names in the list. The default is a comma plus an interword space. See \cmd{finalnamedelim} for an example.\footnote{Note that \cmd{multinamedelim} is not used at all if there are only two names in the list. In this case, the default styles use the \cmd{finalnamedelim}.}

\csitem{finalnamedelim}\CSdelimMark
The delimiter printed instead of \cmd{multinamedelim} before the final name in a name list. The default is the localised term <and>, separated by interword spaces. Here is an example:

\begin{ltxexample}
Michel Goossens<<,>> Frank Mittelbach <<and>> Alexander Samarin
Edward Jones <<and>> Joe Williams
\end{ltxexample}
%
The comma in the first example is the \cmd{multinamedelim} whereas the string <and> in both examples is the \cmd{finalnamedelim}. See also \cmd{finalandcomma} in \secref{use:fmt:lng}.

\csitem{revsdnamedelim}\CSdelimMark
An extra delimiter printed after the first name in a name list if the first name is reversed (only in lists with two names). The default is an empty string, \ie no extra delimiter will be printed. Here is an example showing a name list with a comma as \cmd{revsdnamedelim}:

\begin{ltxexample}
Jones, Edward<<, and>> Joe Williams
\end{ltxexample}
%
In this example, the comma after <Edward> is the \cmd{revsdnamedelim} whereas the string <and> is the \cmd{finalnamedelim}, printed in addition to the former.

\csitem{andothersdelim}\CSdelimMark
The delimiter printed before the localisation string <\texttt{andothers}> if a name list like \bibfield{author} or \bibfield{editor} is truncated. The default is an interword space.

\csitem{multilistdelim}\CSdelimMark
The delimiter printed between multiple items in a literal list like \bibfield{publisher} or \bibfield{location} if there are more than two items in the list. The default is a comma plus an interword space. See \cmd{multinamedelim} for further explanation.

\csitem{finallistdelim}\CSdelimMark
The delimiter printed instead of \cmd{multilistdelim} before the final item in a literal list. The default is the localised term <and>, separated by interword spaces. See \cmd{finalnamedelim} for further explanation.

\csitem{andmoredelim}\CSdelimMark
The delimiter printed before the localisation string <\texttt{andmore}> if a literal list like \bibfield{publisher} or \bibfield{location} is truncated. The default is an interword space.

\csitem{multicitedelim}
The delimiter printed between citations if multiple entry keys are passed to a single citation command. The default is a semicolon plus an interword space.

\csitem{multiciterangedelim}
The delimiter printed between two citations if they are compressed to a range. The default is \cmd{bibrangedash}.

\csitem{multicitesubentrydelim}
The delimiter printed between subentry citations of the same set. This delimiter is only used in citation styles that reduce citations of the same set to a more compact form (\opt{subentry} of \texttt{numeric-comp}). The default is a comma.

\csitem{multicitesubentryrangedelim}
The delimiter printed between two citations of the same set if they are compressed to a range. The default is \cmd{multiciterangedelim}.

\csitem{supercitedelim}
Similar to \cmd{multicitedelim}, but used by the \cmd{supercite} command only. The default is a comma.

\csitem{superciterangedelim}
Analogue of \cmd{multiciterangedelim} for \cmd{supercite}. The default is \cmd{bibrangedash}.

\csitem{supercitesubentrydelim}
Analogue of \cmd{multicitesubentrydelim} for \cmd{supercite}. The default is \cmd{supercitedelim}.

\csitem{supercitesubentryrangedelim}
Analogue of \cmd{multicitesubentryrangedelim} for \cmd{supercite}. The default is \cmd{superciterangedelim}.

\csitem{compcitedelim}
Similar to \cmd{multicitedelim}, but used by certain citation styles when <compressing> multiple citations. The default definition is a comma plus an interword space.

\csitem{textcitedelim}
Similar to \cmd{multicitedelim}, but used by \cmd{textcite} and related commands (\secref{use:cit:cbx}). The default is a comma plus an interword space. The standard styles modify this provisional definition to ensure that the delimiter before the final citation is the localised term <and>, separated by interword spaces. See also \cmd{finalandcomma} and \cmd{finalandsemicolon} in \secref{use:fmt:lng}.

\csitem{nametitledelim}\CSdelimMark
The delimiter printed between the author\slash editor and the title by author-title and some verbose citation styles and in the bibliography. In author-year bibliography styles this delimiter is placed after the author\slash editor and year and before the title. The default definition inside bibliographies is the now deprecated \cmd{labelnamepunct} and is a comma plus an interword space otherwise.

\csitem{nameyeardelim}\CSdelimMark
The delimiter printed between the author\slash editor and the year by author-year citation and bibliography styles. The default definition is an interword space.

\csitem{namelabeldelim}\CSdelimMark
The delimiter printed between the name\slash title and the label by alphabetic and numeric citation styles. The default definition is an interword space.

\csitem{nonameyeardelim}\CSdelimMark
The delimiter printed between the substitute for the labelname when it does not exist (usually the label or title in standard styles) and the year in author-year citation and bibliography styles. This is only used when there is no labelname since when the labelname exists, \cmd{nameyeardelim} is used. The default definition is an interword space.

\csitem{authortypedelim}\CSdelimMark
The delimiter printed between the author and the \texttt{authortype}. The default is a comma followed by a space.

\csitem{editortypedelim}\CSdelimMark
The delimiter printed between the editor and the \texttt{editor} or \texttt{editortype} string. The default is a comma followed by a space.

\csitem{translatortypedelim}\CSdelimMark
The delimiter printed between the translator and the \texttt{translator} string. The default is a comma followed by a space.

\csitem{labelalphaothers}
A string to be appended to the non"=numeric portion of the \bibfield{labelalpha} field (\ie the field holding the citation label used by alphabetic citation styles) if the number of authors\slash editors exceeds the \opt{maxalphanames} threshold or the \bibfield{author}\slash \bibfield{editor} list was truncated in the \file{bib} file with the keyword <\texttt{and others}>. This will typically be a single character such as a plus sign or an asterisk. The default is a plus sign. This command may also be redefined to an empty string to disable this feature. In any case, it must be redefined in the preamble.

\csitem{sortalphaothers}
Similar to \cmd{labelalphaothers} but used in the sorting process. Setting it to a different value is advisable if the latter contains formatting commands, for example:

\begin{ltxexample}
\renewcommand*{\labelalphaothers}{\textbf{+}}
\renewcommand*{\sortalphaothers}{+}
\end{ltxexample}
%
If \cmd{sortalphaothers} is not redefined, it defaults to \cmd{labelalphaothers}.

\csitem{volcitedelim}
The delimiter printed between the volume portion and the page/text portion of \cmd{volcite} and related commands (\secref{use:cit:spc}).

\cmditem{mkvolcitenote}{volume}{pages}

This macro formats the \prm{volume} and \prm{pages} arguments of \cmd{volcite} and related commands (\secref{use:cit:spc}) when they are passed on to the underlying citation command.

\csitem{prenotedelim}\CSdelimMark
The delimiter printed after the \prm{prenote} argument of a citation command. See \secref{use:cit} for details. The default is an interword space.

\csitem{postnotedelim}\CSdelimMark
The delimiter printed before the \prm{postnote} argument of a citation command. See \secref{use:cit} for details. The default is a comma plus an interword space.

\csitem{extpostnotedelim}\CSdelimMark
The delimiter printed between the citation and the parenthetical \prm{postnote} argument of a citation command when the postnote occurs outside of the citation parentheses. In the standard styles, this occurs when the citation uses the shorthand field of the entry. See \secref{use:cit} for details. The default is an interword space.

\csitem{multiprenotedelim}\CSdelimMark
The delimiter printed after the \prm{multiprenote} argument of a citation command. See \secref{use:cit} for details. The default is \cs{prenotedelim}.

\csitem{multipostnotedelim}\CSdelimMark
The delimiter printed before the \prm{multipostnote} argument of a citation command. See \secref{use:cit} for details. The default is \cs{postnotedelim}.

\cmditem{mkbibname<namepart>}{text}
This command, which takes one argument, is used to format the name part <namepart> of name list fields. The default datamodel defines the name parts <family>, <given>, <prefix> and <suffix> and therefore the following macros are automatically defined:

\begin{ltxexample}
\mkbibnamefamily
\mkbibnamegiven
\mkbibnameprefix
\mkbibnamesuffix
\end{ltxexample}
%
For backwards compatibility with the legacy \bibtex name parts, the following are also defined, will generate warnings and will set the correct macro:

\begin{ltxexample}
\mkbibnamelast
\mkbibnamefirst
\mkbibnameaffix
\end{ltxexample}

\cmditem{mkbibcompletenamefamily}{text}
This command, which takes one argument, is used to format the complete name in \texttt{family} format order.

\cmditem{mkbibcompletenamefamilygiven}{text}
This command, which takes one argument, is used to format the complete name in \texttt{family-given} format order.

\cmditem{mkbibcompletenamegivenfamily}{text}
This command, which takes one argument, is used to format the complete name in \texttt{given-family} format order.

\cmditem{mkbibcompletename}{text}
The initial value of all default formatting hooks \cmd{mkbibcompletename<formatorder>}.

\csitem{datecircadelim}\CSdelimMark
When formatting dates with the global option \opt{datecirca} enabled, the delimiter printed after any localised <circa> term. Defaults to interword space.

\csitem{dateeradelim}\CSdelimMark
When formatting dates with the global option \opt{dateera} set, the delimiter printed before the localisation era term. Defaults to interword space.

\csitem{dateuncertainprint}
Prints date uncertainty information when the global option \opt{dateuncertain} is enabled and the \cmd{ifdateuncertain} test is true. By default, prints the language specific \cmd{bibdateuncertain} string (\secref{use:fmt:lng}).

\csitem{enddateuncertainprint}
Prints date uncertainty information when the global option \opt{dateuncertain} is enabled and the \cmd{ifenddateuncertain} test is true. By default, prints the language specific \cmd{bibdateuncertain} string (\secref{use:fmt:lng}).

\csitem{datecircaprint}
Prints date circa information when the global option \opt{datecirca} is enabled and the \cmd{ifdatecirca} test is true. By default, prints the <circa> localised term (\secref{aut:lng:key:dt}) and the \opt{datecircadelim} delimiter.

\csitem{enddatecircaprint}
Prints date circa information when the global option \opt{datecirca} is enabled and the \cmd{ifenddatecirca} test is true. By default, prints the <circa> localised term (\secref{aut:lng:key:dt}) and the \opt{datecircadelim} delimiter.

\csitem{datecircaprintiso}
Prints \acr{ISO8601-2} format date circa information when the global option \opt{datecirca} is enabled and the \cmd{ifdatecirca} test is true. Prints \cmd{textasciitilde}.

\csitem{enddatecircaprintiso}
Prints \acr{ISO8601-2} format date circa information when the global option \opt{datecirca} is enabled and the \cmd{ifenddatecirca} test is true. Prints \cmd{textasciitilde}.

\cmditem{dateeraprint}{yearfield}
Prints date era information when the global option \opt{dateera} is set to <secular> or <christian>. By default, prints the \opt{dateeradelim} delimiter and the appropriate localised era term (\secref{aut:lng:key:dt}). If the \opt{dateeraauto} option is set, then the passed \prm{yearfield} (which is the name of a year field such as <year>, <origyear>, <endeventyear> etc.) is tested to see if its value is earlier than the \opt{dateeraauto} threshold and if so, then the BCE/CE localisation will be output too. The default setting for \opt{dateeraauto} is 0 and so only BCE/BC localisation strings are candidates for output. Detects whether the start or end year era information is to be printed by looking at the \prm{yearfield} name passed to it.

\csitem{dateeraprintpre}
Prints date era information when the global option \opt{dateera} is set to <astronomical>. By default, prints \opt{bibdataeraprefix}. Detects whether the start or end year era information is to be printed by looking at the \prm{yearfield} name passed to it.

\csitem{relatedpunct}
The separator between the \bibfield{relatedtype} bibliography localisation string and the data from the first related entry. Here is an example with \cmd{relatedpunct} set to a dash:

\begin{ltxexample}
A. Smith. Title. 2000, (Orig. pub. as<<->>Origtitle)
\end{ltxexample}

\csitem{relateddelim}
The generic separator between the data of multiple related entries. The default definition is an optional dot plus linebreak. Here is an example where volumes A-E are related entries of the 5 volume main work:

\begin{ltxexample}
Donald E. Knuth. Computers & Typesetting. 5 vols. Reading, Mass.: Addison-Wesley, 1984-1986.
Vol. A: The TEXbook. 1984.
Vol. B: TEX: The Program. 1986.
Vol. C: The METAFONTbook. By. 1986.
Vol. D: METAFONT: The Program. 1986.
Vol. E: Computer Modern Typefaces. 1986.
\end{ltxexample}

\csitem{relateddelim$<$relatedtype$>$}
The separator between the data of multiple related entries inside related entries of type <relatedtype>. There is no default, if such a type-specific delimiter does not exist, \cmd{relateddelim} is used.

\csitem{begrelateddelim}
The generic separator before the block of related entries. The default definition is \cmd{newunitpunct}.

\csitem{begrelateddelim$<$relatedtype$>$}
The separator between the block of related entries of type <relatedtype>. There is no default, if such a type-specific delimiter does not exist, \cmd{relateddelim} is used.

\end{ltxsyntax}

\subsubsection{Context-sensitive Delimiters}
\label{use:fmt:csd}
The delimiters described in \secref{use:fmt:fmt} are globally defined. That is, no matter where you use them, they print the same thing. This is not necessarily desirable for delimiters which you might want to print different things in different contexts. Here <context> means things like <inside a text citation> or <inside a bibliography item>. For this reason, \biblatex\ provides a more sophisticated delimiter specification and user interface alongside the standard one based on normal macros defined with \cmd{newcommand}.

\begin{ltxsyntax}

\cmditem{DeclareDelimFormat}[context, \dots]{name, \dots}{code}
\cmditem*{DeclareDelimFormat}*[context, \dots]{name, \dots}{code}

Declares the delimiter macros in the comma"=separated list \prm{names} with the replacement text \prm{code}. If the optional comma"=separated list of \prm{contexts} is given, declare the \prm{names} only for those contexts. \prm{names} defined without any \prm{contexts} behave just like the global delimiter definitions which \cmd{newcommand} gives---just a plain macro with a replacement which can be used as \cmd{name}. However, you can also call delimiter macros defined in this way by using \cmd{printdelim}, which is context-aware. The starred version clears all \prm{context} specific declarations for all \prm{names} first.

\cmditem{DeclareDelimAlias}[alias context, \dots]{alias}[delim context]{delim}

Declares \prm{alias} to be an alias for the delimiter \prm{delim}. If the optional \prm{alias context, \dots} nor \prm{delim context} are given, the assigment is performed for all existing contexts of the target \prm{delim} separately, so that \prm{alias} becomes an exact copy of \prm{delim} in all contexts. If only the second optional argument \prm{delim context} is given, all existing contexts of \prm{alias} will be cleared and the global/empty context becomes an alias of \prm{delim} in the context \prm{delim context}. The first optional argument \prm{alias context, \dots} may hold a list of contexts for which the alias is assigned. In that case the second optional argument \prm{delim context} specifies the context of the target delimiter. This argument may not be a list, it can only hold one context. If it is missing, the \prm{alias context} is assumed (if \prm{alias context} is a list, the assignment is performed separately for each list item), if it is empty the global context is used.


\begin{ltxexample}[style=latex]
\DeclareDelimAlias[bib,biblist]{finalnamedelim}[]{multinamedelim}
\end{ltxexample}
%
Defines the \opt{bib} and \opt{biblist} contexts of \cmd{finalnamedelim} as aliases of \cmd{multinamedelim} in global context.
On the other hand
\begin{ltxexample}[style=latex]
\DeclareDelimAlias[bib,biblist]{finalnamedelim}{multinamedelim}
\end{ltxexample}
%
defines \cmd{finalnamedelim} in the context \opt{bib} to be an alias of \cmd{multinamedelim} in the \opt{bib} context and defines \cmd{finalnamedelim} in \opt{biblist} context to be an alias of \cmd{multinamedelim} in \opt{biblist}.

\cmditem*{DeclareDelimAlias}*[alias context, \dots]{alias}[delim context]{delim}\DeprecatedMark

The starred version of \cmd{DeclareDelimAlias} is deprecated in favour of using unstarred \cmd{DeclareDelimAlias} with optional arguments.

It assigns the delimiter alias for specific contexts only. The first optional argument \prm{alias context} holds a list of contexts for which the assignment is going to be performed. If it is empty or missing the global/empty context is assumed. The second optional argument \prm{delim context} specifies the context of the target delimiter. This argument may not be a list, it can only hold one context. If it is missing the \prm{alias context} is assumed (if \prm{alias context} is a list, the assignment is performed separately for each list item), if it is empty the global context is used.

\cmditem{printdelim}[context]{name}

Prints a delimiter with name \prm{name}, locally establishing a optional \prm{context} first. Without the optional \prm{context}, \cmd{printdelim} uses the currently active delimiter context.

Delimiter contexts are simply a string, the value of the internal macro \cmd{blx@delimcontext} which can be set manually by the command \cmd{delimcontext}

\cmditem{delimcontext}{context}

Set the delimiter context to \prm{context}. This setting is not global so that delimiter contexts can be nested using the usual \latex group method.

\cmditem{DeclareDelimcontextAlias}{alias}{name}

The context-sensitive delimiter system creates delimiter contexts based on
the name of citation commands (<parencite>, <textcite> etc.) passed to
\cmd{DeclareCiteCommand}. In certain cases where there are nested
definitions of citation commands where \cmd{DeclareCiteCommand} calls
itself (see the definition of \cmd{textcite} in \sty{authoryear-icomp}
for example). The delimiter context is then usually incorrect and the
delimiter specifications do not work. For example, the definition of
\cmd{textcite} in fact defines and uses \cmd{cbx@textcite} and so the
context is automatically set to \opt{cbx@textcite} when printing the
citation. Delimiter definitions expecting to see the context \opt{textcite}
therefore do not work. Therefore this command is provided for style authors
which aliases the context \prm{alias} to the context \prm{name}. For
example:

\begin{ltxexample}[style=latex]
\DeclareDelimcontextAlias{cbx@textcite}{textcite}
\end{ltxexample}
%
This (which is a default setting), makes sure that when inside the
\cmd{cbx@textcite} citation command, the context is in fact \opt{textcite}
as expected.

\end{ltxsyntax}
%
\biblatex\ has several default contexts which are established automatically in various places:

\begin{description}
\item[none] At begin document
\item[bib] Inside a bibliography begun with \cmd{printbibliography} or inside a \cmd{usedriver}
\item[biblist] Inside a bibliography list begun with \cmd{printbiblist}
\item[<citecommand>] Inside a citation command \cmd{citecommand} defined with \cmd{DeclareCiteCommand}
\end{description}

For example, the defaults for \cmd{nametitledelim} are:

\begin{ltxexample}[style=latex]
\DeclareDelimFormat{nametitledelim}{\addcomma\space}
\DeclareDelimFormat[bib,biblist]{nametitledelim}{\labelnamepunct}
\DeclareDelimFormat[textcite]{nametitledelim}{\addspace}
\end{ltxexample}
%
This means that \cmd{nametitledelim} is defined globally as <\cmd{addcomma}\cmd{space}> as per the standard delimiter interface. However, in addition, the delimiter can be printed using \cmd{printdelim} which would print the same as \cmd{nametitledelim} apart from inside a \cmd{textcite}, in which it would print \cmd{addspace} which is more suitable for running text, and in a bibliography (list) in which it takes the value of \cmd{labelnamepunct} for compatibility reasons. If desired, a context can be forced with the optional argument to \cmd{printdelim}, so

\begin{ltxexample}[style=latex]
\printdelim[textcite]{nametitledelim}
\end{ltxexample}
%
Would print \cmd{addspace} regardless of the surrounding context of the \cmd{printdelim}. Contexts are just arbitrary strings and so you can establish them at any time, using \cmd{delimcontext}. If \cmd{printdelim} finds no special value for the delimiter \prm{name} in the current context, it simply prints \cmd{name}. This means that style authors can use \cmd{printdelim} and users expecting to be able to use \cmd{renewcommand} to redefine delimiters can do so with one caveat---such a definition won't change any context-specific delimiters which are defined:

\begin{ltxexample}[style=latex]
\DeclareDelimFormat{delima}{X}
\DeclareDelimFormat[textcite]{delima}{Y}
\renewcommand*{\delima}{Z}
\end{ltxexample}
%
Here, \cmd{delima} always prints <Z>. \verb+\printdelim{delima}+ in any context other than <textcite> also prints \cmd{delima} and hence <Z> but in a <textcite> context prints <Y>. See the \file{04-delimiters.tex} example file that comes with \biblatex\ for more information.

\subsubsection{Language-specific Commands}
\label{use:fmt:lng}

The commands in this section are language specific. When redefining them, you need to wrap the new definition in a \cmd{DeclareBibliographyExtras} command (in an \file{.lbx} file) or a \cmd{DefineBibliographyExtras} command (user documents), see \secref{use:lng} for details. Note that all commands starting with \cmd{mk\dots} take one or more arguments.

\begin{ltxsyntax}

\csitem{bibrangedash}

The language specific dash to be used for ranges of numbers. Defaults to \cmd{textendash}.

\csitem{bibrangessep}

The language specific separator to be used between multiple ranges. Defaults to a comma followed by a space.

\csitem{bibdatesep}

The language specific separator used between date components in terse date formats. Defaults to \cmd{hyphen}.

\csitem{bibdaterangesep}

The language specific separator to be used for date ranges. Defaults to \cmd{textendash} for all date formats apart from \opt{ymd} which defaults to a \cmd{slash}. The date format option \opt{iso} is hard-coded to \cmd{slash} since this is a standards compliant format.

\csitem{mkbibdatelong}

Takes the names of three field as arguments which correspond to three date components (in the order year\slash month\slash day) and uses their values to print the date in the language specific long date format.

\csitem{mkbibdateshort}

Similar to \cmd{mkbibdatelong} but using the language specific short date format.

\csitem{mkbibtimezone}

Modifies a timezone string passed in as the only argument. By default this changes <Z> to the value of \cmd{bibtimezone}.

\csitem{bibdateuncertain}

The language specific marker to be used after uncertain dates when the global option \opt{dateuncertain} is enabled. Defaults to a space followed by a question mark.

\csitem{bibdateeraprefix}

The language specific marker which is printed as a prefix to beginning BCE/BC dates in a date range when the option \opt{dateera} is set to <astronomical>. Defaults to \cmd{textminus}, if defined and \cmd{textendash} otherwise.

\csitem{bibdateeraendprefix}

The language specific marker which is printed as a prefix to end BCE/BC dates in a date range when the option \opt{dateera} is set to <astronomical>. Defaults to a thin space followed by \cmd{bibdateeraprefix} when \cmd{bibdaterangesep} is set to a dash and to \cmd{bibdateeraprefix} otherwise.  This is a separate macro so that you may add extra space before a negative date marker which, for example follows a dash date range marker as this can look a little odd.

\csitem{bibtimesep}

The language specific marker which separates time components. Defaults to a colon.

\csitem{bibutctimezone}

The language specific string printed for the UTC timezone. Defaults to <Z>.

\csitem{bibtimezonesep}

The language specific marker which separates an optional time zone component from a time. Empty by default.

\csitem{bibtzminsep}

The language specific marker which separates hour and minute component of offset timezones. Defaults to a \cmd{bibtimesep}.

\csitem{bibdatetimesep}

The language specific separator printed between date and time components when printing time components along with date components (see the \opt{$<$datetype$>$dateusetime} option in \secref{use:opt:pre:gen}). Defaults to a space for non-\acr{ISO8601-2} output formats, and 'T' for \acr{ISO8601-2} output format.

\csitem{finalandcomma}

Prints the comma to be inserted before the final <and> in a list, if applicable in the respective language. Here is an example:

\begin{ltxexample}
Michel Goossens, Frank Mittelbach<<,>> and Alexander Samarin
\end{ltxexample}
%
\cmd{finalandcomma} is the comma before the word <and>. See also \cmd{multinamedelim}, \cmd{finalnamedelim}, \cmd{textcitedelim}, and \cmd{revsdnamedelim} in \secref{use:fmt:fmt}.

\csitem{finalandsemicolon}

Prints the semicolon to be inserted before the final <and> in a list of lists, if applicable in the respective language. Here is an example:

\begin{ltxexample}
Goossens, Mittelbach, and Samarin; Bertram and Wenworth<<;>> and Knuth
\end{ltxexample}
%
\cmd{finalandsemicolon} is the semicolon before the word <and>. See also \cmd{textcitedelim} in \secref{use:fmt:fmt}.

\cmditem{mkbibordinal}{integer}

This command, which takes an integer as its argument, prints an ordinal number.

\cmditem{mkbibmascord}{integer}

Similar to \cmd{mkbibordinal}, but prints a masculine ordinal, if applicable in the respective language.

\cmditem{mkbibfemord}{integer}

Similar to \cmd{mkbibordinal}, but prints a feminine ordinal, if applicable in the respective language.

\cmditem{mkbibneutord}{integer}

Similar to \cmd{mkbibordinal}, but prints a neuter ordinal, if applicable in the respective language.

\cmditem{mkbibordedition}{integer}

Similar to \cmd{mkbibordinal}, but intended for use with the term <edition>.

\cmditem{mkbibordseries}{integer}

Similar to \cmd{mkbibordinal}, but intended for use with the term <series>.

\end{ltxsyntax}

\subsubsection{Lengths and Counters}
\label{use:fmt:len}

The length registers and counters in this section may be changed in the document preamble with \cmd{setlength} and \cmd{setcounter}, respectively.

\begin{ltxsyntax}

\lenitem{bibhang}

The hanging indentation of the bibliography, if applicable. This length is initialized to \cmd{parindent} at load-time. If \cmd{parindent} is zero length for some reason, \cmd{bibhang} will default to \texttt{1em}.

\lenitem{biblabelsep}

The horizontal space between entries and their corresponding labels in the bibliography. This only applies to bibliography styles which print labels, such as the \texttt{numeric} and \texttt{alphabetic} styles. This length is initialized to twice the value of \cmd{labelsep} at load-time.

\lenitem{bibitemsep}

The vertical space between the individual entries in the bibliography. This length is initialized to \cmd{itemsep} at load-time. Note that \len{bibitemsep}, \len{bibnamesep}, and \len{bibinitsep} obey the rules for \cmd{addvspace}, that is, when vertical space introduced by any of these commands immediately follows on from space introduced by another of them, the resulting total space is equal to the largest of them.

\lenitem{bibnamesep}

Vertical space to be inserted between two entries in the bibliography whenever an entry starts with a name which is different from the initial name of the previous entry. The default value is zero. Setting this length to a positive value greater than \len{bibitemsep} will group the bibliography by author\slash editor name. Note that \len{bibitemsep}, \len{bibnamesep}, and \len{bibinitsep} obey the rules for \cmd{addvspace}, that is, when vertical space introduced by any of these commands immediately follows on from space introduced by another of them, the resulting total space is equal to the largest of them.

\lenitem{bibinitsep}

Vertical space to be inserted between two entries in the bibliography whenever an entry starts with a letter which is different from the initial letter of the previous entry. The default value is zero. Setting this length to a positive value greater than \len{bibitemsep} will group the bibliography alphabetically. Note that \len{bibitemsep}, \len{bibnamesep}, and \len{bibinitsep} obey the rules for \cmd{addvspace}, that is, when vertical space introduced by any of these commands immediately follows on from space introduced by another of them, the resulting total space is equal to the largest of them.

\lenitem{bibparsep}

The vertical space between paragraphs within an entry in the bibliography. The default value is zero.

\cntitem{abbrvpenalty}

This counter, which is used by the localisation modules, holds the penalty used in short or abbreviated localisation strings. For example, a linebreak in expressions such as et al. or ed. by is unfortunate, but should still be possible to prevent overfull boxes. This counter is initialized to \cmd{hyphenpenalty} at load-time. The idea is making \tex treat the whole expression as if it were a single, hyphenatable word as far as line"=breaking is concerned. If you dislike such linebreaks, use a higher value. If you do not mind them at all, set this counter to zero. If you want to suppress them unconditionally, set it to <infinite> (10\,000 or higher).\footnote{The default values assigned to \cnt{abbrvpenalty}, \cnt{lownamepenalty}, and \cnt{highnamepenalty} are deliberately very low to prevent overfull boxes. This implies that you will hardly notice any effect on line-breaking if the text is set justified. If you set these counters to 10\,000 to suppress the respective breakpoints, you will notice their effect but you may also be confronted with overfull boxes. Keep in mind that line-breaking in the bibliography is often more difficult than in the body text and that you can not resort to rephrasing a sentence. In some cases it may be preferable to set the entire bibliography \cmd{raggedright} to prevent suboptimal linebreaks. In this case, even the fairly low default penalties will make a visible difference.}

\cntitem{highnamepenalty}

This counter holds a penalty affecting line"=breaking in names. Please refer to \secref{use:cav:nam,use:fmt:fmt} for explanation. The counter is initialized to \cmd{hyphenpenalty} at load-time. Use a higher value if you dislike the respective linebreaks. If you do not mind them at all, set this counter to zero. If you prefer the traditional \bibtex behavior (no linebreaks at \cnt{highnamepenalty} breakpoints), set it to <infinite> (10\,000 or higher).

\cntitem{lownamepenalty}

Similar to \cnt{highnamepenalty}. Please refer to \secref{use:cav:nam,use:fmt:fmt} for explanation. The counter is initialized to half the \cmd{hyphenpenalty} at load-time. Use a higher value if you dislike the respective linebreaks. If you do not mind them at all, set this counter to zero.

\cntitem{biburlnumpenalty}

If this counter is set to a value greater than zero, \biblatex will permit linebreaks after numbers in all strings formatted with the \cmd{url} command from the \sty{url} package. This will affect \acr{url}s and \acr{doi}s in the bibliography. The breakpoints will be penalized by the value of this counter. If \acr{url}s and/or \acr{doi}s in the bibliography run into the margin, try setting this counter to a value greater than zero but less than 10000 (you normally want to use a high value like 9000). Setting the counter to zero disables this feature. This is the default setting.

\cntitem{biburlucpenalty}

Similar to \cnt{biburlnumpenalty}, except that it will add a breakpoint after all uppercase letters.

\cntitem{biburllcpenalty}

Similar to \cnt{biburlnumpenalty}, except that it will add a breakpoint after all lowercase letters.

\cntitem{biburlbigbreakpenalty}

The \sty{biblatex} version of \sty{url}'s \len{UrlBigBreakPenalty}. The default value is \texttt{100}.

\cntitem{biburlbreakpenalty}

The \sty{biblatex} version of \sty{url}'s \len{UrlBreakPenalty}. The default value is \texttt{200}.

\lenitem{biburlbigskip}

The \sty{biblatex} version of \len{Urlmuskip}. This length holds the additional (stretchable) space inserted around breakable characters in the \cmd{url} command from the \sty{url} package. The default value is \texttt{0mu plus 3mu}.

\lenitem{biburlnumskip}

The additional space inserted after numbers in strings formatted with the \cmd{url} command from the \sty{url} package. This will affect \acr{url}s and \acr{doi}s in the bibliography. If \acr{url}s and/or \acr{doi}s in the bibliography run into the margin, it may help to set this length to add some small stretchable space, for example \texttt{0mu plus 1mu}. The default setting is \texttt{0mu}. This value is only used if \cnt{biburlnumpenalty} is set to a value different from zero.

\lenitem{biburlucskip}

Similar to \cnt{biburlnumskip}, except that it will add space after all uppercase letters.

\lenitem{biburllcskip}

Similar to \cnt{biburlnumskip}, except that it will add space after all uppercase letters.

\end{ltxsyntax}

\subsubsection{All-purpose Commands}
\label{use:fmt:aux}

The commands in this section are all-purpose text commands which are generally available, not only in citations and the bibliography.

\begin{ltxsyntax}

\csitem{bibellipsis}

An ellipsis symbol with brackets: <[\dots\unkern]>.

\csitem{noligature}

Disables ligatures at this position and adds some space. Use this command to break up standard ligatures like <fi> and <fl>. It is similar to the \verb+"|+ shorthand provided by some language modules of the \sty{babel}/\sty{polyglossia} packages.

\csitem{hyphenate}

A conditional hyphen. In contrast to the standard \cmd{-} command, this one allows hyphenation in the rest of the word. It is similar to the \verb|"-| shorthand provided by some language modules of the \sty{babel}/\sty{polyglossia} packages.

\csitem{hyphen}

An explicit, breakable hyphen intended for compound words. In contrast to a literal <\texttt{-}>, this command allows hyphenation in the rest of the word. It is similar to the \verb|"=| shorthand provided by some language modules of the \sty{babel}/\sty{polyglossia} packages.

\csitem{nbhyphen}

An explicit, non-breakable hyphen intended for compound words. In contrast to a literal <\texttt{-}>, this command does not permit line breaks at the hyphen but still allows hyphenation in the rest of the word. It is similar to the \verb|"~| shorthand provided by some language modules of the \sty{babel}/\sty{polyglossia} packages.

\csitem{nohyphenation}

A generic switch which suppresses hyphenation locally. Its scope should normally be confined to a group. The command uses a language without hyphenation patterns to suppress hyphenation. The idea was taken from Peter Wilson's \sty{hyphenat} package. Note that this command should only be used for small portions of text and that its effects are negated if \sty{babel}/\sty{polyglossia} is used to switch the language while it is active.

\cmditem{textnohyphenation}{text}

Similar to \cmd{nohyphenation} but restricted to the \prm{text} argument.

\cmditem{mknumalph}{integer}

Takes an integer in the range 1--702 as its argument and converts it to a string as follows: 1=a, \textellipsis, 26=z, 27=aa, \textellipsis, 702=zz. This is intended for use in formatting directives for the \bibfield{extradate}, \bibfield{extraname} and \bibfield{extraalpha} fields.

\cmditem{mkbibacro}{text}

Generic command which typesets an acronym using the small caps variant of the current font, if available, and as-is otherwise. The acronym should be given in uppercase letters.

\cmditem{autocap}{character}

Automatically converts the \prm{character} to its uppercase form if \biblatex's punctuation tracker would capitalize a localisation string at the current location. This command is robust. It is useful for conditional capitalization of certain strings in an entry. Note that the \prm{character} argument is a single character given in lowercase. For example:

\begin{ltxexample}
\autocap{s}pecial issue
\end{ltxexample}
%
will yield <Special issue> or <special issue>, as appropriate. If the string to be capitalized starts with an inflected character given in Ascii notation, include the accent command in the \prm{character} argument as follows:

\begin{ltxexample}
\autocap{\'e}dition sp\'eciale
\end{ltxexample}
%
This will yield <dition spciale> or <dition spciale>. If the string to be capitalized starts with a command which prints a character, such as \cmd{ae} or \cmd{oe}, simply put the command in the \prm{character} argument:

\begin{ltxexample}
\autocap{\oe}uvres
\end{ltxexample}
%
This will yield <uvres> or <uvres>.

\end{ltxsyntax}

\subsection[Language notes]{Language-specific Notes}
\label{use:loc}

The facilities discussed in this section are specific to certain localisation modules.

\subsubsection{American}
\label{use:loc:us}

The American localisation module uses \cmd{uspunctuation} from \secref{aut:pct:cfg} to enable <American-style> punctuation. If this feature is enabled, all trailing commas and periods after \cmd{mkbibquote} will be moved inside the quotes. If you want to disable this feature, use \cmd{stdpunctuation} as follows:

\begin{ltxexample}
\DefineBibliographyExtras{american}{%
  \stdpunctuation
}
\end{ltxexample}
%
By default, the <American punctuation> feature is enabled by the \texttt{american} localisation module only. The above code is only required if you want American localisation without American punctuation. Since standard punctuation is the package default, it would be redundant with any other language.

It is highly advisable to always specify \texttt{american}, \texttt{british}, \texttt{australian}, etc. rather than \texttt{english} when loading the \sty{babel}/\sty{polyglossia} packages to avoid any possible confusion. Older versions of the \sty{babel} package used to treat \opt{english} as an alias for \opt{british}; more recent ones treat it as an alias for \opt{american}. The \biblatex package essentially treats \texttt{english} as an alias for \opt{american}, except for the above feature which is only enabled if \texttt{american} is requested explicitly.

\subsubsection{Bulgarian}
\label{use:loc:bul}

Like the Greek localisation module, the Bulgarian module also requires \utf support. It will not work with any other encoding.

\subsubsection{Greek}
\label{use:loc:grk}

The Greek localisation module requires \utf support. It will not work with any other encoding. Generally speaking, the \biblatex package is compatible with the \sty{inputenc} package and with the Unicode engines \lualatex and \xelatex. The \sty{ucs} package will not work. Note that you may need to load additional packages which set up Greek fonts. As a rule of thumb, a setup which works for regular Greek documents should also work with \biblatex. However, there is one fundamental limitation. As of this writing, \biblatex has no support for switching scripts. Greek titles in the bibliography should work fine, but English and other titles in the bibliography may be rendered in Greek letters. If you need multi-script bibliographies, using a Unicode engine is the only sensible choice.

\subsubsection{Hungarian}
\label{use:loc:hun}

The Hungarian localisation module needs to redefine certain field formats to obtain the grammatically correct word order. This means that these field formats are overwritten whenever the Hungarian localisation is active, no matter whether they were defined in the preamble or by a custom style. So please be aware that using the Hungarian localisation module may cause the bibliography output to deviate from the format dictated by the loaded style and preamble definitions. Changes to this behaviour need to be made using \cmd{DefineBibliographyExtras}. In particular \cmd{mkpageprefix} is redefined to output the <page> or <pages> string as a suffix after the page number following Hungarian conventions, and all formats of fields involving pages, chapters and volumes were modified so that numbers are printed as ordinals. The Hungarian localisation module will print a warning to remind you of these changes whenever it is loaded in a document. The warning tells you how to silence it.

\subsubsection{Latvian}
\label{use:loc:lav}

The Latvian localisation module, like the Hungarian language module, needs to redefine certain field formats to obtain the grammatically correct word order. This means that these field formats are overwritten whenever the Latvian localisation is active, no matter whether they were defined in the preamble or by a custom style. So please be aware that using the Latvian localisation module may cause the bibliography output to deviate from the format dictated by the loaded style and preamble definitions. Changes to this behaviour need to be made using \cmd{DefineBibliographyExtras}. In particular \cmd{mkpageprefix} is redefined to output the <page> or <pages> string as a suffix after the page number following Latvian conventions, and all formats of fields involving pages, chapters and volumes were modified so that numbers are printed as ordinals. The Latvian localisation module will print a warning to remind you of these changes whenever it is loaded in a document. The warning tells you how to silence it.

\subsubsection{Lithuanian}
\label{use:loc:lit}

The Lithuanian localisation module needs \utf support and will only work with this encoding.

\subsubsection{Russian}
\label{use:loc:rus}

Like the Greek and Lithuanian localisation module, the Russian module also requires \utf support. It will not work with any other encoding.

\subsubsection{Spanish}
\label{use:loc:esp}

Handling the word <and> is more difficult in Spanish than in the other languages supported by this package because it may be <y> or <e>, depending on the initial sound of the following word. Therefore, the Spanish localisation module does not use the localisation string <\texttt{and}> but a special internal <smart and> command. The behavior of this command is controlled by the \cnt{smartand} counter.

\begin{ltxsyntax}

\cntitem{smartand}

This counter controls the behavior of the internal <smart and> command. When set to 1, it prints <y> or <e>, depending on the context. When set to 2, it always prints <y>. When set to 3, it always prints <e>. When set to 0, the <smart and> feature is disabled. This counter is initialized to 1 at load-time and may be changed in the preamble. Note that setting this counter to a positive value implies that the Spanish localisation module ignores \cmd{finalnamedelim} and \cmd{finallistdelim}.

\csitem{forceE}

Use this command in \file{bib} files if \biblatex gets the <and> before a certain name wrong. As its name suggests, it will enforce <e>. This command must be used in a special way to be correct \bibtex datafile format. Here is an example:

\begin{lstlisting}[style=bibtex]{}
author = {Edward Jones and Eoin Maguire},
author = {Edward Jones and <<{\forceE{E}}>>oin Maguire},
\end{lstlisting}
%
Note that the initial letter of the respective name component is given as an argument to \cmd{forceE} and that the entire construct is wrapped in an additional pair of curly braces.

\csitem{forceY}

Similar to \cmd{forceE} but enforces <y>.

\end{ltxsyntax}

\subsubsection{Turkish}
\label{use:loc:tur}

By default \sty{babel}'s Turkish localisation module makes <\texttt{=}> a <shorthand>, which breaks the \keyval parser uses by \biblatex. This problem can be resolved by telling \sty{babel} not to make <\texttt{=}> a shorthand (for example by loading \sty{babel} with the option \texttt{shorthands=:!}) or by loading a \keyval package that can deal with active characters (\sty{kvsetkeys} and \sty{xkeyval})\fnurl{https://tex.stackexchange.com/a/160428/35864}.


\subsection{Usage Notes}
\label{use:use}

The following sections give a basic overview of the \biblatex package and discuss some typical usage scenarios.

\subsubsection{Overview}
\label{use:use:int}

Using the \biblatex package is slightly different from using traditional \bibtex styles and related packages. Before we get to specific usage scenarios, we will therefore have a look at the structure of a typical document first:

\begin{ltxexample}
\documentclass{...}
\usepackage[...]{biblatex}
<<\addbibresource>>{<<bibfile.bib>>}
\begin{document}
<<\cite>>{...}
...
<<\printbibliography>>
\end{document}
\end{ltxexample}
%
With traditional \bibtex, the \cmd{bibliography} command serves two purposes. It marks the location of the bibliography and it also specifies the \file{bib} file(s). The file extension is omitted. With \biblatex, resources are specified in the preamble with  \cmd{addbibresource} using the full name with \file{.bib} suffix. The bibliography is printed using the \cmd{printbibliography} command which may be used multiple times (see \secref{use:bib} for details). The document body may contain any number of citation commands (\secref{use:cit}). Processing this example file requires that a certain procedure be followed. Suppose our example file is called \path{example.tex} and our bibliographic data is in \path{bibfile.bib}. The procedure, then, is as follows:

\begin{enumerate}

\item Run \bin{latex} on \path{example.tex}. If the file contains any citations, \biblatex will request the respective data from \biber by writing commands to the auxiliary file \path{example.bcf}.
\item Run \bin{biber} on \path{example.bcf}. \biber will retrieve the data from \path{bibfile.bib} and write it to the auxiliary file \path{example.bbl} in a format which can be processed by \biblatex.
\item Run \bin{latex} on \path{example.tex}. \biblatex will read the data from \path{example.bbl} and print all citations as well as the bibliography.

\end{enumerate}

\subsubsection{Auxiliary Files}
\label{use:use:aux}

The \biblatex package uses one auxiliary \file{bcf} file only. Even if there are citation commands in a file included via \cmd{include}, you only need to run \biber on the main \file{bcf} file. All information \biber needs is in the \file{bcf} file, including information about all refsections if using multiple \env{refsection} environments (see \secref{use:use:mlt}).

\subsubsection{Multiple Bibliographies}
\label{use:use:mlt}

In a collection of articles by different authors, such as a conference proceedings volume for example, it is very common to have one bibliography for each article rather than a global one for the entire book. In the example below, each article would be presented as a separate \cmd{chapter} with its own bibliography.

\begin{ltxexample}
\documentclass{...}
\usepackage{biblatex}
\addbibresource{...}
\begin{document}
\chapter{...}
<<\begin{refsection}>>
...
<<\printbibliography[heading=subbibliography]>>
<<\end{refsection}>>
\chapter{...}
<<\begin{refsection}>>
...
<<\printbibliography[heading=subbibliography]>>
<<\end{refsection}>>
\end{document}
\end{ltxexample}
%
If \cmd{printbibliography} is used inside a \env{refsection} environment, it automatically restricts the scope of the list of references to the enclosing \env{refsection} environment. For a cumulative bibliography which is subdivided by chapter but printed at the end of the book, use the \opt{section} option of \cmd{printbibliography} to select a reference section, as shown in the next example.

\begin{ltxexample}
\documentclass{...}
\usepackage{biblatex}
<<\defbibheading>>{<<subbibliography>>}{%
  \section*{References for Chapter \ref{<<refsection:\therefsection>>}}}
\addbibresource{...}
\begin{document}
\chapter{...}
<<\begin{refsection}>>
...
<<\end{refsection}>>
\chapter{...}
<<\begin{refsection}>>
...
<<\end{refsection}>>
\printbibheading
<<\printbibliography>>[<<section=1>>,<<heading=subbibliography>>]
<<\printbibliography>>[<<section=2>>,<<heading=subbibliography>>]
\end{document}
\end{ltxexample}
%
Note the definition of the bibliography heading in the above example. This is the definition taking care of the subheadings in the bibliography. The main heading is generated with a plain \cmd{chapter} command in this case. The \biblatex package automatically sets a label at the beginning of every \env{refsection} environment, using the standard \cmd{label} command. The identifier used is the string \texttt{refsection:} followed by the number of the respective \env{refsection} environment. The number of the current section is accessible via the \cnt{refsection} counter. When using the \opt{section} option of \cmd{printbibliography}, this counter is also set locally. This means that you may use the counter in heading definitions to print subheadings like References for Chapter 3, as shown above. You could also use the title of the respective chapter as a subheading by loading the \sty{nameref} package and using \cmd{nameref} instead of \cmd{ref}:

\begin{ltxexample}
\usepackage{<<nameref>>}
\defbibheading{subbibliography}{%
  \section*{<<\nameref{refsection:\therefsection}>>}}
\end{ltxexample}
%
Since giving one \cmd{printbibliography} command for each part of a subdivided bibliography is tedious, \biblatex provides a shorthand. The \cmd{bibbysection} command automatically loops over all reference sections. This is equivalent to giving one \cmd{printbibliography} command for every section but has the additional benefit of automatically skipping sections without references. In the example above, the bibliography would then be generated as follows:

\begin{ltxexample}
\printbibheading
<<\bibbysection[heading=subbibliography]>>
\end{ltxexample}
%
When using a format with one cumulative bibliography subdivided by chapter (or any other document division) it may be more appropriate to use \env{refsegment} rather than \env{refsection} environments. The difference is that the \env{refsection} environment generates labels local to the environment while \env{refsegment} does not affect the generation of labels, hence they will be unique across the entire document. The next example could also be given in \secref{use:use:div} because, visually, it creates one global bibliography subdivided into multiple segments.

\begin{ltxexample}
\documentclass{...}
\usepackage{biblatex}
<<\defbibheading>>{<<subbibliography>>}{%
  \section*{References for Chapter \ref{<<refsegment:\therefsection\therefsegment>>}}}
\addbibresource{...}
\begin{document}
\chapter{...}
<<\begin{refsegment}>>
...
<<\end{refsegment}>>
\chapter{...}
<<\begin{refsegment}>>
...
<<\end{refsegment}>>
\printbibheading
<<\printbibliography>>[<<segment=1>>,<<heading=subbibliography>>]
<<\printbibliography>>[<<segment=2>>,<<heading=subbibliography>>]
\end{document}
\end{ltxexample}
%
The use of \env{refsegment} is similar to \env{refsection} and there is also a corresponding \opt{segment} option for \cmd{printbibliography}. The \biblatex package automatically sets a label at the beginning of every \env{refsegment} environment using the string \texttt{refsegment:} followed by the number of the respective \env{refsegment} environment as an identifier. There is a matching \cnt{refsegment} counter which may be used in heading definitions, as shown above. As with reference sections, there is also a shorthand command which automatically loops over all reference segments:

\begin{ltxexample}
\printbibheading
<<\bibbysegment[heading=subbibliography]>>
\end{ltxexample}
%
This is equivalent to giving one \cmd{printbibliography} command for every segment in the current \env{refsection}.

\subsubsection{Subdivided Bibliographies}
\label{use:use:div}

It is very common to subdivide a bibliography by certain criteria. For example, you may want to list printed and online resources separately or divide a bibliography into primary and secondary sources. The former case is straightforward because you can use the entry type as a criterion for the \opt{type} and \opt{nottype} filters of \cmd{printbibliography}. The next example also demonstrates how to generate matching subheadings for the two parts of the bibliography.

\begin{ltxexample}
\documentclass{...}
\usepackage{biblatex}
\addbibresource{...}
\begin{document}
...
\printbibheading
\printbibliography[<<nottype=online>>,heading=subbibliography,
                   <<title={Printed Sources}>>]
\printbibliography[<<type=online>>,heading=subbibliography,
                   <<title={Online Sources}>>]

\end{document}
\end{ltxexample}
%
You may also use more than two subdivisions:

\begin{ltxexample}
\printbibliography[<<type=article>>,...]
\printbibliography[<<type=book>>,...]
\printbibliography[<<nottype=article>>,<<nottype=book>>,...]
\end{ltxexample}
%
It is even possible to give a chain of different types of filters:

\begin{ltxexample}
\printbibliography[<<section=2>>,<<type=book>>,<<keyword=abc>>,<<notkeyword=xyz>>]
\end{ltxexample}
%
This would print all works cited in reference section~2 whose entry type is \bibtype{book} and whose \bibfield{keywords} field includes the keyword <abc> but not <xyz>. When using bibliography filters in conjunction with a numeric style, see \secref{use:cav:lab}. If you need complex filters with conditional expressions, use the \opt{filter} option in conjunction with a custom filter defined with \cmd{defbibfilter}. See \secref{use:bib:flt} for details on custom filters.

\begin{ltxexample}
\documentclass{...}
\usepackage{biblatex}
\addbibresource{...}
\begin{document}
...
\printbibheading
\printbibliography[<<keyword=primary>>,heading=subbibliography,%
                   <<title={Primary Sources}>>]
\printbibliography[<<keyword=secondary>>,heading=subbibliography,%
                   <<title={Secondary Sources}>>]
\end{document}
\end{ltxexample}
%
Dividing a bibliography into primary and secondary sources is possible with a \bibfield{keyword} filter, as shown in the above example. In this case, with only two subdivisions, it would be sufficient to use one keyword as filter criterion:

\begin{ltxexample}
\printbibliography[<<keyword=primary>>,...]
\printbibliography[<<notkeyword=primary>>,...]
\end{ltxexample}
%
Since \biblatex has no way of knowing if an item in the bibliography is considered to be primary or secondary literature, we need to supply the bibliography filter with the required data by adding a \bibfield{keywords} field to each entry in the \file{bib} file. These keywords may then be used as targets for the \opt{keyword} and \opt{notkeyword} filters, as shown above. It may be a good idea to add such keywords right away while building a \file{bib} file.

\begin{lstlisting}[style=bibtex]{}
@Book{key,
  <<keywords>>      = {<<primary>>,some,other,keywords},
  ...
\end{lstlisting}
%
An alternative way of subdividing the list of references are bibliography categories. They differ from the keywords"=based approach shown in the example above in that they work on the document level and do not require any changes to the \file{bib} file.

\begin{ltxexample}
\documentclass{...}
\usepackage{biblatex}
<<\DeclareBibliographyCategory>>{<<primary>>}
<<\DeclareBibliographyCategory>>{<<secondary>>}
<<\addtocategory>>{<<primary>>}{key1,key3,key6}
<<\addtocategory>>{<<secondary>>}{key2,key4,key5}
\addbibresource{...}
\begin{document}
...
\printbibheading
\printbibliography[<<category=primary>>,heading=subbibliography,%
                   <<title={Primary Sources}>>]
\printbibliography[<<category=secondary>>,heading=subbibliography,%
                   <<title={Secondary Sources}>>]
\end{document}
\end{ltxexample}
%
In this case it would also be sufficient to use one category only:

\begin{ltxexample}
\printbibliography[<<category=primary>>,...]
\printbibliography[<<notcategory=primary>>,...]
\end{ltxexample}
%
It is still a good idea to declare all categories used in the bibliography explicitly because there is a \cmd{bibbycategory} command which automatically loops over all categories. This is equivalent to giving one \cmd{printbibliography} command for every category, in the order in which they were declared.

\begin{ltxexample}
\documentclass{...}
\usepackage{biblatex}
<<\DeclareBibliographyCategory>>{<<primary>>}
<<\DeclareBibliographyCategory>>{<<secondary>>}
\addtocategory{primary}{key1,key3,key6}
\addtocategory{secondary}{key2,key4,key5}
<<\defbibheading>>{<<primary>>}{\section*{Primary Sources}}
<<\defbibheading>>{<<secondary>>}{\section*{Secondary Sources}}
\addbibresource{...}
\begin{document}
...
\printbibheading
<<\bibbycategory>>
\end{document}
\end{ltxexample}
%
The handling of the headings is different from \cmd{bibbysection} and \cmd{bibbysegment} in this case. \cmd{bibbycategory} uses the name of the current category as a heading name. This is equivalent to passing \texttt{heading=\prm{category}} to \cmd{printbibliography} and implies that you need to provide a matching heading for every category.

\subsubsection{Entry Sets}
\label{use:use:set}

An entry set is a group of entries which are cited as a single reference and listed as a single item in the bibliography. The individual entries in the set are separated by \cmd{entrysetpunct} (\secref{aut:fmt:fmt}). The \biblatex package supports two types of entry sets. Static entry sets are defined in the \file{bib} file like any other entry. Dynamic entry sets are defined with \cmd{defbibentryset} (\secref{use:bib:set}) on a per-document\slash per-refsection basis in the document preamble or the document body. This section deals with the definition of entry sets; style authors should also see \secref{aut:cav:set} for further information. Please note that entry sets only make sense for styles which refer to entries by labels such as the provided \texttt{numeric} and \texttt{alphabetic} styles. Styles which refer to entries via names, titles etc. (\texttt{authoryear}, \texttt{authortitle}, \texttt{verbose} etc.) rarely employ sets and do not support them by default when they are cited directly. Custom styles may of course choose to implement some manner of set citation support in any manner they choose.

\paragraph{Static entry sets}

Static entry sets are defined in the \file{bib} file like any other entry. Defining an entry set is as simple as adding an entry of type \bibtype{set}. The entry has an \bibfield{entryset} field defining the members of the set as a separated list of entry keys:

\begin{lstlisting}[style=bibtex]{}
<<@Set>>{<<set1>>,
  <<entryset>> = {<<key1,key2,key3>>},
}
\end{lstlisting}
%
Entries may be part of a set in one document\slash refsection and stand-alone references in another one, depending on the presence of the \bibtype{set} entry. If the \bibtype{set} entry is cited, the set members are grouped automatically. If not, they will work like any regular entry. Note that with \bibtex as the backend, there must also be an \bibfield{entryset} field in the set members which point to the set parent. With \biber, this is not necessary.

\paragraph[Dynamic entry sets]{Dynamic entry sets}

Dynamic entry sets are set up and work much like static ones. The main difference is that they are defined in the document preamble or on the fly in the document body using the \cmd{defbibentryset} command from \secref{use:bib:set}:

\begin{lstlisting}[style=bibtex]{}
\defbibentryset{set1}{key1,key2,key3}
\end{lstlisting}
%
Dynamic entry sets in the document body are local to the enclosing \env{refsection} environment, if any. Otherwise, they are assigned to reference section~0. Those defined in the preamble are assigned to reference section~0.

\subsubsection[Data Containers]{Data Containers}
\label{use:use:xdat}

The \bibtype{xdata} entry type serves as a data container holding one or more fields. Data in \bibtype{xdata} entries may be referenced and used by other entries. \bibtype{xdata} entries may not be cited or added to the bibliography, they only serve as a data source for other entries (including other \bibtype{xdata} entries). This data inheritance mechanism is useful for fixed field combinations such as \bibfield{publisher}\slash \bibfield{location} and for other frequently used data:

\begin{lstlisting}[style=bibtex]{}
<<@XData>>{<<hup>>,
  publisher  = {Harvard University Press},
  location   = {Cambridge, Mass.},
}
@Book{...,
  author     = {...},
  title	     = {...},
  date	     = {...},
  <<xdata>>      = {<<hup>>},
}
\end{lstlisting}
%
Using a separated list of keys in its \bibfield{xdata} field, an entry may inherit data from several \bibtype{xdata} entries. Cascading \bibtype{xdata} entries are supported as well, \ie an \bibtype{xdata} entry may reference one or more other \bibtype{xdata} entries:

\begin{lstlisting}[style=bibtex]{}
@XData{macmillan:name,
  publisher  = {Macmillan},
}
@XData{macmillan:place,
  location   = {New York and London},
}
@XData{macmillan,
  xdata      = {macmillan:name,macmillan:place},
}
@Book{...,
  author     = {...},
  title	     = {...},
  date	     = {...},
  xdata	     = {macmillan},
}
\end{lstlisting}
%
More granular \bibtype{xdata} entry data may be referenced. It is not necessary to reference only entire fields. For example:

\begin{lstlisting}[style=bibtex]{}
@XData{someauthors,
  author     = {John Smith and Brian Brown}
}
@XData{somelocations,
  location   = {Location1 and Location2}
}
@XData{somenotes,
  note   = {A note}
}
@Book{...,
  author     = {Alan Drudge and xdata=someauthors-author-2},
  editor     = {xdata=someauthors-author},
  location   = {xdata=somelocations-location-1 and Location3},
  note       = {xdata=somenotes-note}
}
\end{lstlisting}
%
The format of granular \bibtype{xdata} references are as follows:

\begin{namesample}
\item~\delim{x}{1}data\delim{=}{2}\delim{$<$}{3}key$>$\delim{-}{4}\delim{$<$}{5}field$>$\delim{-}{6}\delim{$<$}{7}index$>$
\end{namesample}

\begin{enumerate}
  \item The value of the \biber option \opt{--xdatamarker} (by default '\texttt{xdata}')
  \item The value of the \biber option \opt{--xnamesep} (by default '\texttt{=}')
  \item A valid entry key of an \bibtype{xdata} entry
  \item The value of the \biber option \opt{--xdatasep} (by default '\texttt{-}')
  \item A valid entry field in the source \bibtype {xdata} entry
  \item (Optional) The value of the \biber option \opt{--xdatasep} (by default '\texttt{-}')
  \item (Optional) A valid 1-based index into a list/name field in the source \bibtype {xdata} entry
\end{enumerate}
%
There are \opt{--output-*} variants of the above options for \biber tool mode output so that these separators and markers can be programatically changed. Taking the example above, this \bibtype{book} would resolve to:

\begin{lstlisting}[style=bibtex]{}
@Book{...,
  author     = {Alan Drudge and Brian Brown},
  editor     = {John Smith},
  location   = {Location1 and Location3},
  note       = {A note}
}
\end{lstlisting}
%
Things to note with granular \bibtype{xdata} references:

\begin{itemize}
  \item References must be made only to \bibtype{xdata} fields. An warning will be generated otherwise and the reference will not be resolved
  \item References must be made only to \bibtype{xdata} fields of the same type (list/name and datatype) as the referencing field. An warning will be generated otherwise and the reference will not be resolved
  \item References to fields of datatype 'date' are not possible. References to legacy \bibfield{year} and \bibfield{month} fields are possible
  \item References to missing entries, fields or list/name indices will generate a warning and the reference will not be resolved
  \item If an index is missing for a reference to a list/name field, 1 is assumed
\end{itemize}
%
See also \secref{bib:typ:blx,bib:fld:spc}.

\subsubsection{Electronic Publishing Information}
\label{use:use:epr}

The \biblatex package provides three fields for electronic publishing information: \bibfield{eprint}, \bibfield{eprinttype}, and \bibfield{eprintclass}. The \bibfield{eprint} field is a verbatim field similar to \bibfield{doi} which holds the identifier of the item. The \bibfield{eprinttype} field holds the resource name, \ie the name of the site or electronic archive. The optional \bibfield{eprintclass} field is intended for additional information specific to the resource indicated by the \bibfield{eprinttype} field. This could be a section, a path, classification information, etc. If the \bibfield{eprinttype} field is available, the standard styles will use it as a literal label. In the following example, they would print Resource: identifier rather than the generic eprint: identifier:

\begin{lstlisting}[style=bibtex]{}
<<eprint>>     = {<<identifier>>},
<<eprinttype>> = {<<Resource>>},
\end{lstlisting}
%
The standard styles feature dedicated support for a few online archives. For arXiv references, put the identifier in the \bibfield{eprint} field and the string \texttt{arxiv} in the \bibfield{eprinttype} field:

\begin{lstlisting}[style=bibtex]{}
<<eprint>>     = {<<math/0307200v3>>},
<<eprinttype>> = {<<arxiv>>},
\end{lstlisting}
%
For papers which use the new identifier scheme (April 2007 and later) add the primary classification in the \bibfield{eprintclass} field:

\begin{lstlisting}[style=bibtex]{}
eprint      = {1008.2849v1},
eprinttype  = {arxiv},
<<eprintclass>> = {<<cs.DS>>},
\end{lstlisting}
%
There are two aliases which ease the integration of arXiv entries. \bibfield{archiveprefix} is treated as an alias for \bibfield{eprinttype}; \bibfield{primaryclass} is an alias for \bibfield{eprintclass}. If hyperlinks are enabled, the eprint identifier will be transformed into a link to \nolinkurl{arxiv.org}. See the package option \opt{arxiv} in \secref{use:opt:pre:gen} for further details.

For \acr{JSTOR} references, put the stable \acr{JSTOR} number in the \bibfield{eprint} field and the string \texttt{jstor} in the \bibfield{eprinttype} field:

\begin{lstlisting}[style=bibtex]{}
<<eprint>>     = {<<number>>},
<<eprinttype>> = {<<jstor>>},
\end{lstlisting}
%
When using \acr{JSTOR}'s export feature to export citations in \bibtex format, \acr{JSTOR} uses the \bibfield{url} field by default (where the \prm{number} is a unique and stable identifier):

\begin{lstlisting}[style=bibtex]{}
url = {http://www.jstor.org/stable/<<number>>},
\end{lstlisting}
%
While this will work as expected, full \acr{URL}s tend to clutter the bibliography. With the \bibfield{eprint} fields, the standard styles will use the more readable \acr{JSTOR}: \prm{number} format which also supports hyperlinks. The \prm{number} becomes a clickable link if \sty{hyperref} support is enabled.

For PubMed references, put the stable PubMed identifier in the \bibfield{eprint} field and the string \texttt{pubmed} in the \bibfield{eprinttype} field. This means that:

\begin{lstlisting}[style=bibtex]{}
url = {http://www.ncbi.nlm.nih.gov/pubmed/<<pmid>>},
\end{lstlisting}
%
becomes:

\begin{lstlisting}[style=bibtex]{}
<<eprint>>     = {<<pmid>>},
<<eprinttype>> = {<<pubmed>>},
\end{lstlisting}
%
and the standard styles will print \acr{PMID}: \prm{pmid} instead of the lengthy \acr{URL}. If hyperref support is enabled, the \prm{pmid} will be a clickable link to PubMed.

For handles (\acr{HDL}s), put the handle in the \bibfield{eprint} field and the string \texttt{hdl} in the \bibfield{eprinttype} field:

\begin{lstlisting}[style=bibtex]{}
<<eprint>>     = {<<handle>>},
<<eprinttype>> = {<<hdl>>},
\end{lstlisting}
%
For Google Books references, put Google's identifier in the \bibfield{eprint} field and the string \texttt{googlebooks} in the \bibfield{eprinttype} field. This means that, for example:

\begin{lstlisting}[style=bibtex]{}
url = {http://books.google.com/books?id=<<XXu4AkRVBBoC>>},
\end{lstlisting}
%
would become:

\begin{lstlisting}[style=bibtex]{}
<<eprint>>     = {<<XXu4AkRVBBoC>>},
<<eprinttype>> = {<<googlebooks>>},
\end{lstlisting}
%
and the standard styles would print Google Books: |XXu4AkRVBBoC| instead of the full \acr{URL}. If hyperref support is enabled, the identifier will be a clickable link to Google Books.\footnote{Note that the Google Books \acr{id} seems to be a bit of an <internal> value. As of this writing, there does not seem to be any way to search for an \acr{id} on Google Books. You may prefer to use the \bibfield{url} in this case.}

Note that \bibfield{eprint} is a verbatim field. Always give the identifier in its unmodified form. For example, there is no need to replace |_| with |\_|. Also see \secref{aut:cav:epr} on how to add dedicated support for other eprint resources.

\subsubsection{External Abstracts and Annotations}
\label{use:use:prf}

Styles which print the fields \bibfield{abstract} and\slash or \bibfield{annotation} may support an alternative way of adding abstracts or annotations to the bibliography. Instead of including the text in the \file{bib} file, it may also be stored in an external \latex file. For example, instead of saying

\begin{ltxexample}[style=bibtex]
@Article{key1,
  ...
  abstract = {This is an abstract of entry `key1'.}
}
\end{ltxexample}
%
in the \file{bib} file, you may create a file named \path{bibabstract-key1.tex} and put the abstract in this file:

\begin{ltxexample}
This is an abstract of entry `key1'.
\endinput
\end{ltxexample}
%
The name of the external file must be the entry key prefixed with \texttt{bibabstract-} or \texttt{bibannotation-}, respectively. You can change these prefixes by redefining \cmd{bibabstractprefix} and \cmd{bibannotationprefix}. Note that this feature needs to be enabled explicitly by setting the package option \opt{loadfiles} from \secref{use:opt:pre:gen}. The option is disabled by default for performance reasons. Also note that any \bibfield{abstract} and \bibfield{annotation} fields in the \file{bib} file take precedence over the external files. Using external files is strongly recommended if you have long abstracts or a lot of annotations since this may increase memory requirements significantly. It is also more convenient to edit the text in a dedicated \latex file. Style authors should see \secref{aut:cav:prf} for further information.

\subsection{Hints and Caveats}
\label{use:cav}

This section provides additional usage hints and addresses some common problems and potential misconceptions.

\subsubsection{Usage with \acr{KOMA}-Script Classes}
\label{use:cav:scr}

When used in conjunction with a recent version\footnote{At least \acr{KOMA}-Script 3.25 (2018/03/30).} one of the \sty{scrbook}, \sty{scrreprt}, or \sty{scrartcl} classes, \biblatex passes control over the (default) headings \texttt{bibliography} and \texttt{biblist} from \secref{use:bib:hdg} to the class. Hence, bibliography-heading-related class options can be used as usual. You can override the default headings by using the \opt{heading} option of \cmd{printbibliography}, \cmd{printbibheading} and \cmd{printbiblist}. See \secref{use:bib:bib, use:bib:biblist, use:bib:hdg} for details.

\biblatex also tries to detect bibliography-related class options and settings itself.\footnote{This applies to the traditional syntax of these options (\opt{bibtotoc} and \opt{bibtotocnumbered}) as well as to the \keyval syntax introduced in \acr{KOMA}-Script 3.x, \ie to \kvopt{bibliography}{nottotoc}, \kvopt{bibliography}{totoc}, and \kvopt{bibliography}{totocnumbered}. The global \kvopt{toc}{bibliography} and \kvopt{toc}{bibliographynumbered} options as well as their aliases are detected as well. In any case, the options must be set globally in the optional argument to \cmd{documentclass}.} This was required to be able to adapt the bibliography headings to the class settings in older versions of \acr{KOMA}-Script. If one of the above classes is detected, \biblatex will provide the following additional tests which may be useful in custom heading definitions. Since these tests rely on the error-prone external detection of \acr{KOMA}-Script settings and are no longer used with newer \acr{KOMA}-Script versions, these tests are deprecated and should no longer be used.

\begin{ltxsyntax}

\cmditem{ifkomabibtotoc}{true}{false}\DeprecatedMark

Expands to \prm{true} if the class would add the bibliography to the table of contents, and to \prm{false} otherwise. This test is deprecated.

\cmditem{ifkomabibtotocnumbered}{true}{false}\DeprecatedMark

Expands to \prm{true} if the class would add the bibliography to the table of contents as a numbered section, and to \prm{false} otherwise. If this test yields \prm{true}, \cmd{ifkomabibtotoc} will always yield \prm{true} as well, but not vice versa. This test is deprecated.

\end{ltxsyntax}

\subsubsection{Usage with the Memoir Class}
\label{use:cav:mem}

When using \biblatex with the \sty{memoir} class, most class facilities for adapting the bibliography have no effect. Use the corresponding facilities of this package instead (\secref{use:bib:bib, use:bib:hdg, use:bib:nts}). Instead of redefining \sty{memoir}'s \cmd{bibsection}, use the \opt{heading} option of \cmd{printbibliography} and \cmd{defbibheading} (\secref{use:bib:bib, use:bib:hdg}). Instead of \cmd{prebibhook} and \cmd{postbibhook}, use the \opt{prenote} and \opt{postnote} options of \cmd{printbibliography} and \cmd{defbibnote} (\secref{use:bib:bib, use:bib:nts}). All default headings are adapted at load-time such that they blend well with the default layout of this class. The default headings \texttt{bibliography} and \texttt{biblist} (\secref{use:bib:hdg}) are also responsive to \sty{memoir}'s \cmd{bibintoc} and \cmd{nobibintoc} switches. The length register \len{bibitemsep} is used by \biblatex in a way similar to \sty{memoir} (\secref{use:fmt:len}). This section also introduces some additional length registers which correspond to \sty{memoir}'s \cmd{biblistextra}. Lastly, \cmd{setbiblabel} does not map to a single facility of the \biblatex package since the style of all labels in the bibliography is controlled by the bibliography style. See \secref{aut:bbx:env} in the author section of this manual for details. If the \sty{memoir} class is detected, \biblatex will also provide the following additional test which may be useful in custom heading definitions:

\begin{ltxsyntax}

\cmditem{ifmemoirbibintoc}{true}{false}

Expands to \prm{true} or \prm{false}, depending on \sty{memoir}'s \cmd{bibintoc} and \cmd{nobibintoc} switches. This is a \latex frontend to \sty{memoir}'s \cmd{ifnobibintoc} test. Note that the logic of the test is reversed.

\end{ltxsyntax}

\subsubsection{Page Numbers in Citations}
\label{use:cav:pag}

If the \prm{postnote} argument to a citation command is a page number or page range, \biblatex will automatically prefix it with <p.> or <pp.> by default. This works reliably in typical cases, but sometimes manual intervention may be required. In this case, it is important to understand how this argument is handled in detail. First, \biblatex checks if the postnote is an Arabic or Roman numeral (case insensitive). If this test succeeds, the postnote is considered as a single page or other number which will be prefixed with <p.> or some other string which depends on the \sty{pagination} field (see \secref{bib:use:pag}). If it fails, a second test is performed to find out if the postnote is a range or a list of Arabic or Roman numerals. If this test succeeds, the postnote will be prefixed with <pp.> or some other string in the plural form. If it fails as well, the postnote is printed as is. Note that both tests expand the \prm{postnote}. All commands used in this argument must therefore be robust or prefixed with \cmd{protect}. Here are a few examples of \prm{postnote} arguments which will be correctly recognized as a single number, a range of numbers, or a list of numbers, respectively:

\begin{ltxexample}
\cite[25]{key}
\cite[vii]{key}
\cite[XIV]{key}
\cite[34--38]{key}
\cite[iv--x]{key}
\cite[185/86]{key}
\cite[XI \& XV]{key}
\cite[3, 5, 7]{key}
\cite[vii--x; 5, 7]{key}
\end{ltxexample}
%
In some other cases, however, the tests may get it wrong and you need to resort to the auxiliary commands \cmd{pnfmt}, \cmd{pno}, \cmd{ppno}, and \cmd{nopp} from \secref{use:cit:msc}. For example, suppose a work is cited by a special pagination scheme consisting of numbers and letters. In this scheme, the string <|27a|> would mean <page~27, part~a>. Since this string does not look like a number or a range to \biblatex, you need to force the prefix for a single number manually:

\begin{ltxexample}
\cite[\pno~27a]{key}
\end{ltxexample}
%
There is also a \cmd{ppno} command which forces a range prefix as well as a \cmd{nopp} command which suppresses all prefixes:

\begin{ltxexample}
\cite[\ppno~27a--28c]{key}
\cite[\nopp 25]{key}
\end{ltxexample}
%
These commands may be used anywhere in the \prm{postnote} argument. They may also be used multiple times. For example, when citing by volume and page number, you may want to suppress the prefix at the beginning of the postnote and add it in the middle of the string:

\begin{ltxexample}
\cite[VII, \pno~5]{key}
\cite[VII, \pno~3, \ppno~40--45]{key}
\cite[see][\ppno~37--46, in particular \pno~40]{key}
\end{ltxexample}
%
The command \cmd{pnfmt} can be used for \prm{postnote}s consisting of a page range and some additional text. \cmd{pnfmt} prints its argument in the format specified for the postnote and can be used to format the page range automatically without the need for \cmd{pno} and \cmd{ppno}.

\begin{ltxexample}
\cite[\pnfmt{37-46}, in particular \pnfmt{40}]{key}
\end{ltxexample}
%
There are also two auxiliary command for suffixes like <the following page(s)>. Instead of inserting such suffixes literally (which would require \cmd{ppno} to force a prefix):

\begin{ltxexample}
\cite[\ppno~27~sq.]{key}
\cite[\ppno~55~sqq.]{key}
\end{ltxexample}
%
use the auxiliary commands \cmd{psq} and \cmd{psqq}. Note that there is no space between the number and the command. This space will be inserted automatically and may be modified by redefining the macro \cmd{sqspace}.

\begin{ltxexample}
\cite[27\psq]{key}
\cite[55\psqq]{key}
\end{ltxexample}
%
Since the postnote is printed without any prefix if it includes any character which is not an Arabic or Roman numeral, you may also type the prefix manually---though this is discouraged:

\begin{ltxexample}
\cite[p.~5]{key}
\end{ltxexample}
%
It is possible to suppress the prefix on a per"=entry basis by setting the \bibfield{pagination} field of an entry to <\texttt{none}>, see \secref{bib:use:pag} for details. If you do not want any prefixes at all or prefer to type them manually, you can also disable the entire mechanism in the document preamble or the configuration file as follows:

\begin{ltxexample}
\DeclareFieldFormat{postnote}{#1}
\end{ltxexample}
%
The \prm{postnote} argument is handled as a field and the formatting of this field is controlled by a field formatting directive which may be freely redefined. The above definition will simply print the postnote as is. See \secref{aut:cbx:fld, aut:bib:fmt} in the author guide for further details.

\subsubsection{Name Parts and Name Spacing}
\label{use:cav:nam}

The \biblatex package gives users and style authors very fine"=grained control of name spacing and the line"=breaking behavior of names. The commands discussed in the following are documented in \secref{use:fmt:fmt,aut:fmt:fmt}. This section is meant to give an overview of how they are put together. A note on terminology: a name \emph{part} is a basic part of the name, for example the given or the family name. Each part of a name may be a single name or it may be composed of multiple names. For example, the name part \enquote*{given name} may be composed of a given and a middle name. The latter are referred to as name \emph{elements} in this section. Let's consider a simple name first: \enquote{John Edward Doe}. This name is composed of the following parts:

\begin{nameparts}
Given	& John Edward \\
Prefix	& --- \\
Family	& Doe \\
Suffix	& --- \\
\end{nameparts}
%
The spacing, punctuation and line"=breaking behavior of names is controlled by six macros:

\begin{namedelims}
a & \cmd{bibnamedelima} & Inserted by the backend after the first element of every name part if that element is less than three characters long and before the last element of every name part. \\
b & \cmd{bibnamedelimb} & Inserted by the backend between all elements of a name part where \cmd{bibnamedelima} does not apply. \\
c & \cmd{bibnamedelimc} & Inserted by a formatting directive between the name prefix and the family name if \kvopt{useprefix}{true}. If \kvopt{useprefix}{false}, \cmd{bibnamedelimd} is used instead. \\
d & \cmd{bibnamedelimd} & Inserted by a formatting directive between name parts where \cmd{bibnamedelimc} does not apply. \\
i & \cmd{bibnamedelimi} & Replaces \cmd{bibnamedelima/b} after initials \\
p & \cmd{revsdnamepunct} & Inserted by a formatting directive after the family name when the name parts are reversed.
\end{namedelims}
%
This is how the delimiters are employed:

\begin{namesample}
\item John\delim{~}{a}Edward\delim{ }{d}Doe
\item Doe\delim{,}{p}\delim{ }{d}John\delim{~}{a}Edward
\end{namesample}
%
Initials in the \file{bib} file get a special delimiter:

\begin{namesample}
\item J.\delim{~}{i}Edward\delim{ }{d}Doe
\end{namesample}
%
Let's consider a more complex name: \enquote{Charles-Jean tienne Gustave Nicolas de La Valle Poussin}. This name is composed of the following parts:

\begin{nameparts}
Given	& Charles-Jean tienne Gustave Nicolas \\
Prefix	& de \\
Family	& La Valle Poussin \\
Suffix	& --- \\
\end{nameparts}
%
The delimiters:

\begin{namesample}
\item Charles-Jean\delim{~}{b}tienne\delim{~}{b}Gustave\delim{~}{a}Nicolas\delim{ }{d}%
      de\delim{ }{c}%
      La\delim{~}{a}Valle\delim{~}{a}Poussin
\end{namesample}
%
Note that \cmd{bibnamedelima/b/i} are inserted by the backend. The backend processes the name parts and takes care of the delimiters between the elements that make up a name part, processing each part individually. In contrast to that, the delimiters between the parts of the complete name (\cmd{bibnamedelimc/d}) are added by name formatting directives at a later point in the processing chain. The spacing and punctuation of initials is also handled by the backend and may be customized by redefining the following three macros:

\begin{namedelims}
a & \cmd{bibinitperiod} & Inserted by the backend after initials.\\
b & \cmd{bibinitdelim} & Inserted by the backend between multiple initials.\\
c & \cmd{bibinithyphendelim} & Inserted by the backend between the initials of hyphenated name parts, replacing \cmd{bibinitperiod} and \cmd{bibinitdelim}.\\
\end{namedelims}
%
This is how they are employed:

\begin{namesample}
\item J\delim{.}{a}\delim{~}{b}E\delim{.}{a} Doe
\item K\delim{.-}{c}H\delim{.}{a} Mustermann
\end{namesample}

\subsubsection{Split Bibliographies and Citation Labels}
\label{use:cav:lab}

The citation labels generated by this package are assigned to the full list of references before it is split up by any bibliography filters. They are guaranteed to be unique across the entire document (or a \env{refsection} environment), no matter how many bibliography filters you are using. When using a numeric citation scheme, however, this will most likely lead to discontinuous numbering in split bibliographies. Use the \opt{defernumbers} package option to avoid this problem. If this option is enabled, numeric labels are assigned the first time an entry is printed in any bibliography.

Compare the output of the following example with \texttt{defernumbers=true} and \texttt{defernumbers=false}.
%
\begin{ltxexample}
\documentclass{article}
\usepackage[defernumbers=true]{biblatex}

\addbibresource{biblatex-examples.bib}

\begin{document}
Lorem \autocite{worman} ipsum \autocite{sigfridsson}
dolor \autocite{nussbaum} sit \autocite{aksin}

\printbibheading[title={Bibliography}]
\printbibliography[heading=subbibliography,
  type=book, title={Books}]
\printbibliography[heading=subbibliography,
  type=article, title={Articles}]
\end{document}
\end{ltxexample}

There are pathological cases where neither \texttt{defernumbers=true} nor \texttt{defernumbers=false} produce fully desirable output. This may be the case when entries can end up in several of the split bibliographies---or if there is an additional global bibliography. But in most cases with non-overlapping split bibliographies \texttt{defernumbers=true} produces better results. \sty{biblatex} will therefore suggest setting \opt{defernumbers} to \opt{true} in a warning when a split bibliography setup is detected. That warning can be suppressed in case setting \opt{defernumbers} to \opt{true} is not desirable.

\subsubsection{Active Characters in Bibliography Headings}
\label{use:cav:act}

Packages using active characters, such as \sty{babel}, \sty{polyglossia}, \sty{csquotes}, or \sty{underscore}, usually do not make them active until the beginning of the document body to avoid interference with other packages. A typical example of such an active character is the Ascii quote |"|, which is used by various language modules of the \sty{babel}/\sty{polyglossia} packages. If shorthands such as |"<| and |"a| are used in the argument to \cmd{defbibheading} and the headings are defined in the document preamble, the non-active form of the characters is saved in the heading definition. When the heading is typeset, they do not function as a command but are simply printed literally. The most straightforward solution consists in moving \cmd{defbibheading} after |\begin{document}|. Alternatively, you may use \sty{babel}'s \cmd{shorthandon} and \cmd{shorthandoff} commands to temporarily make the shorthands active in the preamble. The above also applies to bibliography notes and the \cmd{defbibnote} command.

\subsubsection{Grouping in Reference Sections and Segments}
\label{use:cav:grp}

All \latex environments enclosed in \cmd{begin} and \cmd{end} form a group. This may have undesirable side effects if the environment contains anything that does not expect to be used within a group. This issue is not specific to \env{refsection} and \env{refsegment} environments, but it obviously applies to them as well. Since these environments will usually enclose much larger portions of the document than a typical \env{itemize} or similar environment, they are simply more likely to trigger problems related to grouping. If you observe any malfunctions after adding \env{refsection} environments to a document (for example, if anything seems to be <trapped> inside the environment), try the following syntax instead:

\begin{ltxexample}
\chapter{...}
<<\refsection>>
...
<<\endrefsection>>
\end{ltxexample}
%
This will not from a group, but otherwise works as usual. As far as \biblatex is concerned, it does not matter which syntax you use. The alternative syntax is also supported by the \env{refsegment} environment. Note that the commands \cmd{newrefsection} and \cmd{newrefsegment} do not form a group. See \secref{use:bib:sec, use:bib:seg} for details.

\subsection{Using the fallback \bibtex\ backend}
\label{use:bibtex}

To utilise all of the features described here, \biblatex must be used with the
\biber program as a backend. Indeed, the documentation in general assumes this. However, for a \emph{limited} subset of use cases it is possible to use the
long-established \bibtex program (either the 7-bit \texttt{bibtex} or
8-bit \texttt{bibtex8}) as the supporting backend. This works in much the
same way as for \biber with the only proviso being that \bibtex is much more
limited as a backend.

Using \bibtex as the backend requires that the option \opt{backend=bibtex}
or \opt{backend=bibtex8} is given at load time. The \biblatex package will
then write appropriate data for use by \bibtex into the auxiliary file(s)
and a special data file (automatically included in those to be read by
\bibtex).  The \bibtex(8) program should then be run on each auxiliary file:
\biblatex will list all of the required files in the log.

Key limitations of the \bibtex backend are:
\begin{itemize}
\item Sorting is global and is limited to Ascii ordering
\item No re-encoding is possible and thus database entries must be in
  LICR form to work reliably
\item The data model is fixed
\item Cross-referencing is more limited and entry sets must be written
  into the \path{.bib} file
\item Fixed memory capacity (using the \verb|--wolfgang| option with
  \verb|bibtex8| is strongly recommended to minimize the likelihood of
  an issue here)
\end{itemize}

